---
title: 'Streammetabolizer output for BWU'
author: "Kelly Loria"
date: "2025-02-04"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 12px;}
code.r{font-size: 8px;}
pre {font-size: 10px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWU_k600/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F}
### Packages
setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWU_k600/")

library(tidyverse)
library(lubridate)
library(dataRetrieval)
library(viridis)
library(cowplot)
library(zoo)
library(readxl)
library(ggpubr)
library(streamMetabolizer)
library(knitr)
library(kableExtra)
library(plotly)
```
### Background
This is the **upper** reach of the larger turbulent mountain stream, Blackwood Creek in CA, USA.

#### Identifying reasonable modeled estimates of K600
We then ran the normal stream metabolizer model: `b_Kb_oipi_tr_plrckm.stan` to get modeled K600 to see if we could resolve the negative correlation between ER and K600.
Priors on `K600_lnQ_nodes_meanlog` were set as 5 bins based on `rnorm(1000, mean = logQ_mean, sd = logQ_sd)` centered around the mean and logQ values 1-2 sd away from the mean. 

##### Load different model segments with normal prior for gas exchange. 
We chose segments of time where we believe GPP occurred and was greater than 0. 
These chunks of time are from a previous model where we binned flow and incorporated measured and estimated K600 priors from gas exchange measurements a the reach. 

```{r, warning=F, size = 'small', echo=FALSE}
input_dat <- readRDS("./25_BWU_mod_dat.rds") 

measured_K <- read.csv("./ArN_NEON_K_BWL_notpooled_241204.csv")%>%
  mutate(date= as.Date(date, format=c("%m/%d/%y")))  %>%
  filter(date>as.Date("2024-06-26"))
measured_K$source <- ("measured")


met.full <- read_csv("./BWU_Full_2025-02-04_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Full")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  

fits.full <- read_csv("./BWU_Full_2025-02-04_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Full")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

```

### Metabolism with CIs for the full time series 
This the raw model output. It looks okay aside from some small parts of 2023.
Where GPP is in blue and ER is in orange, and the black points represent NEP.

**Here is the run configuration for full model:**

`mm_name(type = 'bayes', pool_K600 = "binned", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`

**Fitting priors:**

`K600_lnQ_nodes_meanlog = log(16)` Where 16 was the mean value from observed measurements and normal pooled modeled, 
`K600_lnQ_nodes_sdlog = 1.31` 
`bayes_specs_new$K600_lnQ_nodes_centers <- log_bins` was from `prior_samples <- rnorm(1000, mean = logQ_mean, sd = logQ_sd)`




```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=12}
met.full %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~year, scales="free", nrow = 4)+
  theme_bw()
```

### Checking full model convergence
Make sure the chains converged; all r-hat values were well below 1.05 (the red line)
for GPP, ER, and K600. 
The blue lines are the mean for each parameter. 

** Some poor convergence in 2023 for K600 


Looks like that weird 2023 time period corresponds to bad rhats for all parameters. 

```{r, echo=T, warning=F, size = 'small', fig.width=9, fig.height = 2.5}
k600.rhat <- mean(na.omit(met.full$K600_daily_Rhat))
k600.rhat
GPP.rhat <- mean(na.omit(met.full$GPP_Rhat))
GPP.rhat
ER.rhat <- mean(na.omit(met.full$ER_Rhat))
ER.rhat
```

```{r, echo=F, warning=F, size = 'small', fig.width=9, fig.height = 2.5}
# Create a dataframe for the vertical lines
rhat_lines <- data.frame(
  Param = c("K600_daily_Rhat", "GPP_daily_Rhat", "ER_daily_Rhat"),
  yintercept = c(k600.rhat, GPP.rhat, ER.rhat)
)

met.full %>%
  select(date, GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  pivot_longer(2:4, values_to = "Rhat", names_to = "Param") %>%
  ggplot(aes(x = date, y = Rhat), col=lab) +
  geom_hline(yintercept = 1.05, col = "red", lty = 2) + 
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(data = rhat_lines, aes(yintercept = yintercept), col = "blue", lty = 2) + 
  theme_classic() +
  facet_wrap(~Param, ncol = 3)

```


```{r, warning=F, , echo=FALSE, size = 'small', fig.width=9, fig.height = 18}
fits.day <- fits.full%>% 
group_by(date, lab) %>%
 dplyr:: summarize(rmse = sqrt(mean((DO.mod-DO.obs)^2)),
            sd = sd(DO.obs),
            min = min(DO.obs, na.rm = T),
            max = max(DO.obs, na.rm = T),
            range = range(max-min, na.rm = T),
            nrmse = rmse/range,
            minT = min(temp.water, na.rm = T),
            maxT = max(temp.water, na.rm = T),
            rangeT= range(maxT-minT, na.rm = T))

summary(fits.day)
```

### Compared modeled and measured K600

Plots for (1) measured v modeled K600 and flow and (2) logK600 and log(flow+1).

Could be one poor measurement at the highest flow for measured gas exchange.
But in general the modeled K600 does seem similar to the measured, which is kind of nice to see how robust the `pool_K600 = "normal"` is getting at K600.

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=4, fig.height = 2.5}
met.clean <- met.full %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  filter(K600_daily_mean<45)

mean_k_mod <- mean(met.clean$K600_daily_mean)
mean_k_mod
mean_k_measure <- mean(measured_K$K600)
mean_k_measure

KER_cor <- round(cor(met.clean$ER_daily_mean, met.clean$K600_daily_mean, use = "complete.obs"),3)
print(KER_cor)

KGPP_cor <-round(cor(met.clean$GPP_daily_mean, met.clean$K600_daily_mean, use = "complete.obs"),3)
print(KGPP_cor)
```

```{r, warning=F, size = 'small', echo=FALSE, fig.align = "center", fig.width=4, fig.height = 2.5}

met.clean$source <- "modeled"

mod.env.ag <- input_dat%>%
  mutate(date= as.Date(solar.time))%>%
  group_by(date) %>%
  dplyr:: summarise(
    DO.obs = mean(DO.obs, na.rm=T),
    DO.sat = mean(DO.sat, na.rm=T),
    depth = mean(depth, na.rm=T),
    temp.water = mean(temp.water, na.rm=T),
    light = mean(light, na.rm=T),
    discharge =  mean(discharge, na.rm=T))

k.q <- met.clean %>%
  left_join(mod.env.ag, by = "date") %>%
  full_join(measured_K, by = c("date", "source",
                                 "K600_daily_mean"= "K600",
                                "discharge"= "Q_cms")) %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = source, shape=source))+
  geom_point( size = 2) +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "viridis")

k.q
```


### Full run with binned K600 informed by both measured and modeled priors

**Here is the run configuration for full model:**

`mm_name(type = 'bayes', pool_K600 = "binned", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`


#### Quick plot of flow bins given flow range:
Where dashed vertical lines correspond to the prior locations for flow bins in `bayes_specs_new$K600_lnQ_nodes_centers <- log_bins`
```{r, warning=F, size = 'small', echo=FALSE, fig.align = "center", fig.width=9, fig.height = 3}
# Compute log-transformed discharge
logQ <- log(na.omit(input_dat$discharge))
sd_logQ <- round(sd(na.omit(logQ)),2)

## Compute log-transformed K600 for prior estimation
logQ_mean <- round(mean(logQ, na.rm = TRUE),2)  # Center the prior on observed mean
logQ_sd <- round(c(sd(logQ, na.rm = TRUE)),1)  # Double the SD for a wider prior


logK600_mean <- round(log(16),1)  # Center the prior on observed mean
logK600_sd <- round(c(log((29.2/15.8)*2)),2)  # Double the SD for a wider prior

prior_samples <- rnorm(1000, mean = logQ_mean, sd = logQ_sd)
# Convert back from log-scale to natural scale
# Define bin edges based on log-scale mean and SD
log_bins <- c(logQ_mean - c(2 * logQ_sd), 
              logQ_mean - c(1 * logQ_sd), 
              logQ_mean, 
              logQ_mean + c(1 * logQ_sd), 
              logQ_mean + c(2 * logQ_sd))

# Convert bins to log scale
bins <- exp(log_bins)
prior_df <- data.frame(Q_cms = prior_samples)
prior_bins_plot <- ggplot(prior_df, aes(x = Q_cms)) +
  geom_density(fill = "lightblue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = log_bins, color = "black", lty="dashed") +  # Bin edges
  labs(title = "Binned log(flow) for K600 Priors", y = "Density", x = "log(Q) (cms)") +
  theme_bw()
prior_bins_plot
```

Plots made on filtered data: 
**`met.clean`** filtered for days with `GPP_daily_Rhat<1.05`,`ER_daily_Rhat<1.05`,
`K600_daily_Rhat <1.05`, as well as  `(GPP_97.5pct>0)` and  `(ER_2.5pct<0)`.

```{r, warning=F, size = 'small', echo=F, fig.align = "center", fig.width=8, fig.height = 5}

ker <- ggplot(met.clean, aes(x = K600_daily_mean, y = ER_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+
  theme_bw()+ geom_smooth(method="lm", se=F, color="grey50") +
  scale_color_viridis_d(option = "inferno") +  theme(legend.position = "null")

kgpp <- ggplot(met.clean, aes(y = K600_daily_mean, x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")

kyear <- met.clean %>%
ggplot(aes(x = as.factor(year), y = K600_daily_mean, fill = as.factor(lab), col = as.factor(lab)))+
    geom_hline(yintercept = mean_k_mod, lty = 2)+
  geom_boxplot(alpha = 0.8)+
  theme_bw()+
  scale_fill_viridis_d(option ="inferno") +  # Use _d for discrete
  scale_color_viridis_d(option = "inferno") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

k.q <- met.clean %>%
  left_join(mod.env.ag, by = "date") %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = lab))+
  geom_point( size = 2)+ geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "inferno") 

p1 <- plot_grid(ker, kgpp, k.q, rel_widths =c(0.3, 0.3, 0.4), ncol = 3)
plot_grid(p1, kyear, ncol = 1)

```


The vertical dashed is the overall mean modeled K600 in the box plot. 


##### GPP and ER ?
```{r, warning=F, size = 'small', echo=F, fig.align = "center", fig.width=8, fig.height = 5}
GPP_E_cor <-round(cor(met.clean$GPP_daily_mean, met.clean$ER_daily_mean, use = "complete.obs"),3)
print(GPP_E_cor)


kgpp <- ggplot(met.clean, aes(y = (ER_daily_mean*-1), x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50",lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")
kgpp

```

The direction of the K600 ~ flow relationship looks more logical, where K600 increases with flow.
ER and K600 are positively correlated (0.686), GPP and K600 are negatively correlated (-0.587). but, less strongly relative to the lower reach (GBL). The relationship between K600 and flow appears to be positive but still inflected in a strange way. 

However I'm still think we should be cautious in over interpreting ER trends. 

The mean modeled K600 is lower than expected, 6.50 when it should be closer to 16. 

### Final cleaned model output:
Where GPP is in blue and ER is in orange, and the black points represent NEP.

```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=6}
p <- met.clean %>%
  ggplot(aes(x = date)) +
  geom_abline(slope = 0, intercept = 0, linewidth = 1, alpha = 0.6) +
  geom_point(aes(y = GPP_daily_mean), col = "cyan4", alpha = 0.8) +
  geom_errorbar(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
                width = 0.2, color = "cyan4") +
  geom_point(aes(y = ER_daily_mean), col = "chocolate3", alpha = 0.8) +
  geom_errorbar(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
                width = 0.2, color = "chocolate3") +
  geom_point(aes(y = NEP_daily_mean), alpha = 0.5, size = 0.2) +
  theme_bw()

# Convert to Plotly
ggplotly(p)
```


### Table summarizing excluded data from cleaned model output: 
Of the 449 days with DO observations 159 days were removed. 

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=12}

total_days <- length(unique(na.omit(mod.env.ag$date)))

unrealistic_gpp <- met.full %>% 
  filter(GPP_97.5pct < 0 ) 

negative_GPP_day <- length(unique(unrealistic_gpp$date))

unrealistic_ER <- met.full %>% 
  filter(ER_2.5pct > 0 ) 
pos_ER_day <- length(unique(unrealistic_ER$date))

###
df_length <- mod.env.ag %>% 
  left_join(met.full)
num_na_GPP <- sum(is.na(df_length$GPP_mean))

df_length <- mod.env.ag %>% 
  left_join(met.clean)
num_na_clean <- sum(is.na(df_length$GPP_mean))

poor_gpp <- met.full %>% 
  filter(GPP_daily_Rhat > 1.05) 
poor_gpp_day <- length(unique(poor_gpp$date))

poor_ER <- met.full %>% 
  filter(ER_daily_Rhat > 1.05) 
poor_ER_day <- length(unique(poor_ER$date))

poor_K600 <- met.full %>% 
  filter(K600_daily_Rhat > 1.05) 
poor_K600_day <- length(unique(poor_K600$date))


pertot <- round((total_days/total_days)*100,3)
perna <- round((num_na_clean/total_days)*100,1)
pernaGPP <- round((num_na_GPP/total_days)*100,1)
perPoorGPP <- round((poor_gpp_day/total_days)*100,1)
perPoorER <- round((poor_ER_day/total_days)*100,1)
perPoorK600 <- round((poor_K600_day/total_days)*100,1)
perNegGPP <- round((negative_GPP_day/total_days)*100,1)
perPosER <- round((pos_ER_day/total_days)*100,1)


# Create table
excluded_summary <- tibble(
  `Number of Days` = c(total_days, num_na_clean, num_na_GPP,
                       poor_gpp_day, poor_ER_day, poor_K600_day,
                       negative_GPP_day, pos_ER_day), 
  `Explaination` = c("Total days of DO observations", "Total days removed", "Days model was unable to fit", "days where GPP rhat > 1.05", "days where ER rhat > 1.05", "days where K600 rhat > 1.05",
                     "days where modeled GPP was negative", "days where modeled ER was positive"),
    `Percent of Days` = c(pertot,
                          perna, 
                          pernaGPP,
                          perPoorGPP,
                          perPoorER, 
                          perPoorK600, 
                          perNegGPP,
                          perPosER))


# Print formatted table
excluded_summary %>%
  kable() %>%
  kable_styling(font_size = 15) %>%  
  row_spec(0:2, bold = TRUE, font_size = 16) %>% 
  row_spec(2, hline_after = TRUE)  
```



#### Session info

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=12}
pander::pander(sessionInfo())
```