---
title: 'Streammetabolizer output for GBL'
author: "Kelly Loria"
date: "2025-02-04"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 12px;}
code.r{font-size: 8px;}
pre {font-size: 10px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F}
### Packages
setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/")

library(tidyverse)
library(lubridate)
library(dataRetrieval)
library(viridis)
library(cowplot)
library(zoo)
library(readxl)
library(ggpubr)
library(streamMetabolizer)
library(knitr)
library(kableExtra)
library(plotly)

```

### Background
This is the model for the **lower** reach of a small, turbulent, shaded stream with a fair amount of fine particulate organic matter, Glenbrook Creek NV, USA.

##### Identifying reasonable modeled estimates of K600
We then ran normal pooling on k600 in stream metabolizer model: `b_Kb_oipi_tr_plrckm.stan` to get modeled K600 to see if we could resolve the negative correlation between ER and K600.
Priors on `K600_lnQ_nodes_meanlog` were set as 5 bins based on `rnorm(1000, mean = logQ_mean, sd = logQ_sd)` centered around the mean and logQ values 1-2 sd away from the mean. 

##### Load metab data from different model segments with normal prior for gas exchange. 
We chose segments of time where we believe GPP occurred and was greater than 0. 
These chunks of time are from a previous model where we binned flow and incorporated measured and estimated K600 priors from gas exchange measurements a the reach. 

```{r, warning=F, size = 'small', echo=FALSE}
met.May23 <- read_csv("./GBL_May23_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
  mutate(lab = "May23")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  filter(K600_daily_Rhat<1.1)

fits.May23 <- read_csv("./GBL_May23_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
  mutate(lab = "May23")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.May21 <- read_csv("./GBL_May21_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
  mutate(lab = "May21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.May21 <- read_csv("./GBL_May21_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
    mutate(lab = "May21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100)) 

met.Mar24 <- read_csv("./GBL_Mar24_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
    mutate(lab = "Mar24")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar24 <- read_csv("./GBL_Mar24_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Mar24")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.Mar22 <- read_csv("./GBL_Mar22_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
     mutate(lab = "Mar22")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar22 <- read_csv("./GBL_Mar22_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
       mutate(lab = "Mar22")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100)) 


met.Mar21 <- read_csv("./GBL_Mar21_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
       mutate(lab = "Mar21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar21 <- read_csv("./GBL_Mar21_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
       mutate(lab = "Mar21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))


met.Feb24 <- read_csv("./GBL_Feb24_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
       mutate(lab = "Feb24")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Feb24 <- read_csv("./GBL_Feb24_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
   mutate(lab = "Feb24")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))  


met.Apr22 <- read_csv("./GBL_Apr22_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Apr22")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Apr22 <- read_csv("./GBL_Apr22_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Apr22")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.full21 <- read_csv("./GBL_21_newbin_2025-02-02_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Full_21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.full21 <- read_csv("./GBL_21_newbin_2025-02-02_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Full_21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))


met.full <- read_csv("./GBL_Full_2025-02-02_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Full")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  

fits.full <- read_csv("./GBL_Full_2025-02-02_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Full")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.df <- rbind(met.May23, 
                met.May21, 
                met.Mar24,
                met.Mar21,
                met.Apr22,
                met.Feb24)
met.df$source <- ("normal_mod")
fits.df <- rbind(fits.May23, fits.May21, fits.Mar21,
                 fits.Mar24, fits.Apr22, met.Feb24)


input_dat <- readRDS("./25_GBL_mod_dat.rds") 

measured_K <- read.csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/ArN_NEON_K_GBL_notpooled_241204.csv")%>%
  mutate(date= as.Date(date, format=c("%m/%d/%y"))) 
measured_K$source <- ("measured")

```


```{r, warning=F, size = 'small', include=FALSE, echo=FALSE, fig.width=10, fig.height=12}
met.df %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~lab, scales="free", nrow = 4)+
  theme_bw()
```


### Metabolism with CIs for the full time series 
This the raw model output. It looks like there are many days with fitting issues.
Where GPP is in blue and ER is in orange.

**Here is the run configuration for full model:**

`mm_name(type = 'bayes', pool_K600 = "binned", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`

**Fitting priors:**

`K600_lnQ_nodes_meanlog = log(25)` Where 25 was the mean value from observed measurements and normal pooled modeled, 
`K600_lnQ_nodes_sdlog = 1.1` was 1.1, 
`bayes_specs_new$K600_lnQ_nodes_centers <- log_bins` was from `prior_samples <- rnorm(1000, mean = logQ_mean, sd = logQ_sd)`
based on mean discharge for all flows > 0.99 cms to cut off some extreme values.


```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=12}
met.full %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~year, scales="free", nrow = 4)+
  theme_bw()
```

### Checking full model convergence
Make sure the chains converged; all r-hat values were well below 1.05 (the red line)
for GPP, ER, and K600. 
The blue lines are the mean for each parameter. 

```{r, , echo=FALSE, warning=F, include=FALSE, size = 'small', fig.width=9, fig.height = 2.5}
k600.rhat <- mean(na.omit(met.df$K600_daily_Rhat))
GPP.rhat <- mean(na.omit(met.df$GPP_Rhat))
ER.rhat <- mean(na.omit(met.df$ER_Rhat))

# Create a dataframe for the vertical lines
rhat_lines <- data.frame(
  Param = c("K600_daily_Rhat", "GPP_daily_Rhat", "ER_daily_Rhat"),
  yintercept = c(k600.rhat, GPP.rhat, ER.rhat)
)

met.df %>%
  select(date, GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  pivot_longer(2:4, values_to = "Rhat", names_to = "Param") %>%
  ggplot(aes(x = date, y = Rhat), col=lab) +
  geom_hline(yintercept = 1.05, col = "red", lty = 2) + 
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(data = rhat_lines, aes(yintercept = yintercept), col = "blue", lty = 2) + 
  theme_classic() +
  facet_wrap(~Param, ncol = 3)
```



```{r, , echo=T, warning=F, size = 'small', fig.width=9, fig.height = 2.5}
k600.rhat <- mean(na.omit(met.full$K600_daily_Rhat))
k600.rhat
GPP.rhat <- mean(na.omit(met.full$GPP_Rhat))
GPP.rhat
ER.rhat <- mean(na.omit(met.full$ER_Rhat))
ER.rhat

# Create a dataframe for the vertical lines
rhat_lines <- data.frame(
  Param = c("K600_daily_Rhat", "GPP_daily_Rhat", "ER_daily_Rhat"),
  yintercept = c(k600.rhat, GPP.rhat, ER.rhat)
)

met.full %>%
  select(date, GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  pivot_longer(2:4, values_to = "Rhat", names_to = "Param") %>%
  ggplot(aes(x = date, y = Rhat), col=lab) +
  geom_hline(yintercept = 1.05, col = "red", lty = 2) + 
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(data = rhat_lines, aes(yintercept = yintercept), col = "blue", lty = 2) + 
  theme_classic() +
  facet_wrap(~Param, ncol = 3)
```

In general convergence looks alright, but there are still many days with unrealistic GPP or ER.

```{r, warning=F, , echo=FALSE, size = 'small', fig.width=9, fig.height = 18}
fits.day <- fits.full%>% 
group_by(date, lab) %>%
 dplyr:: summarize(rmse = sqrt(mean((DO.mod-DO.obs)^2)),
            sd = sd(DO.obs),
            min = min(DO.obs, na.rm = T),
            max = max(DO.obs, na.rm = T),
            range = range(max-min, na.rm = T),
            nrmse = rmse/range,
            minT = min(temp.water, na.rm = T),
            maxT = max(temp.water, na.rm = T),
            rangeT= range(maxT-minT, na.rm = T),
            minQ= min(discharge, na.rm = T), 
            maxQ= max(discharge, na.rm = T),
            rangeQ= range(maxQ-minQ, na.rm = T)
            )

summary(fits.day)
```

### Looking at the estimated K600 and ER relationships from normal pooling

**Here is the run configuration to get modeled K600:**

`mm_name(type = 'bayes', pool_K600 = "normal", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`

 Where each "lab" segment was run as individual model with the `K600_lnQ_nodes_meanlog` adjusted to match streamline in cms during that time. 

#### K-ER, K-Q, general K600 trends from segements with GPP > 0
There is a strong negative correlation between ER and K600 (-0.933). 
GPP and K600 are slightly less correlated (-0.074).
Still seeing a strange/paradoxical negative correlation between flow and K600.

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=8, fig.height = 5}

mean_k_mod <- mean(met.df$K600_daily_mean)
mean_k_mod
mean_k_measure <- mean(measured_K$K600)
mean_k_measure

possibleK <- (mean_k_mod + mean_k_measure)/2
possibleK

KER_cor <- round(cor(met.df$ER_daily_mean, met.df$K600_daily_mean, use = "complete.obs"),3)
print(KER_cor)

KGPP_cor <-round(cor(met.df$GPP_daily_mean, met.df$K600_daily_mean, use = "complete.obs"),3)
print(KGPP_cor)

```

```{r, warning=F, size = 'small', echo=F, fig.align = "center", fig.width=8, fig.height = 5}
ker <- ggplot(met.df, aes(x = K600_daily_mean, y = ER_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+
  theme_bw()+ geom_smooth(method="lm", se=F, color="grey50") +
  scale_color_viridis_d(option = "inferno") +  theme(legend.position = "null")

kgpp <- ggplot(met.df, aes(y = K600_daily_mean, x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")

kyear <- met.df %>%
ggplot(aes(x = as.factor(year), y = K600_daily_mean, fill = as.factor(lab), col = as.factor(lab)))+
    geom_hline(yintercept = mean_k_mod, lty = 2)+
  geom_boxplot(alpha = 0.8)+
  theme_bw()+
  scale_fill_viridis_d(option ="inferno") +  # Use _d for discrete
  scale_color_viridis_d(option = "inferno") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mod.env.ag <- input_dat%>%
  mutate(date= as.Date(solar.time))%>%
  group_by(date) %>%
  dplyr:: summarise(
    DO.obs = mean(DO.obs, na.rm=T),
    DO.sat = mean(DO.sat, na.rm=T),
    depth = mean(depth, na.rm=T),
    temp.water = mean(temp.water, na.rm=T),
    light = mean(light, na.rm=T),
    discharge =  mean(discharge, na.rm=T))


k.q <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = lab))+
  geom_point( size = 2)+ geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "inferno") 

p1 <- plot_grid(ker, kgpp, k.q, rel_widths =c(0.3, 0.3, 0.4), ncol = 3)
plot_grid(p1, kyear, ncol = 1)

```

The vertical dashed is the overall mean modeled K600 in the box plot. 


### Compared modeled and measured K600

Plots for (1) measured v modeled K600 and flow and (2) logK600 and log(flow+1).

Could be one poor measurement at the highest flow for measured gas exchange.
But in general the modeled K600 does seem similar to the measured, which is kind of nice to see how robust the `pool_K600 = "normal"` is getting at K600.

```{r, warning=F, size = 'small', echo=FALSE, fig.align = "center", fig.width=9, fig.height = 2.5}
k.q <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  full_join(measured_K, by = c("date", "source",
                                 "K600_daily_mean"= "K600",
                                "discharge"= "Q_cms")) %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = source, shape=source))+
  geom_point( size = 2) +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "viridis") 


k.q_log <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  full_join(measured_K, by = c("date", "source",
                                 "K600_daily_mean"= "K600",
                                "discharge"= "Q_cms")) %>%
  ggplot(aes(x = log(discharge+1), y = log(K600_daily_mean), col = source, shape=source))+
  geom_point( size = 2) +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "viridis") 

plot_grid(k.q, k.q_log, ncol = 2)

```



### Full run with binned K600 informed by both measured and modeled priors


**Here is the run configuration for full model:**

`mm_name(type = 'bayes', pool_K600 = "binned", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`

#### Quick plot of flow bins given flow range:

Where dashed vertical lines correspond to the prior locations for flow bins in `bayes_specs_new$K600_lnQ_nodes_centers <- log_bins`

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=9, fig.height = 3}
dat1 <- input_dat%>% filter(discharge<0.99) 
## Set bayes specs
bayes_name_new <- mm_name(type = 'bayes', pool_K600 = "binned",
                          err_obs_iid = TRUE, err_proc_iid = TRUE,
                          ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')
bayes_specs_new <- specs(bayes_name_new)

# Compute log-transformed discharge
logQ <- log(na.omit(dat1$discharge))
sd_logQ <- round(sd(na.omit(logQ)),2)


## Compute log-transformed K600 for prior estimation
logQ_mean <- round(mean(logQ, na.rm = TRUE),2)  # Center the prior flow mean
logQ_mean
logQ_sd <- round(c(sd(logQ, na.rm = TRUE)),1)  # Double the SD for a wider prior
logQ_sd


logK600_mean <- round(log(25),1)  # Center the prior on mean k600
logK600_mean
logK600_sd <- round(c(log(75/25)),2)  # Double the SD for a wider prior
logK600_sd

range_prior_U <- exp(logK600_mean +logK600_sd)
range_prior_U

range_prior_L <- exp(logK600_mean -logK600_sd)
range_prior_L


prior_samples <- rnorm(1000, mean = logQ_mean, sd = logQ_sd)
# Define bins based on log-scale mean and SD
log_bins <- c(logQ_mean - c(2 * logQ_sd), 
              logQ_mean - c(1 * logQ_sd), 
              logQ_mean, 
              logQ_mean + c(1 * logQ_sd), 
              logQ_mean + c(2 * logQ_sd))

# Plot:
#bins <- exp(log_bins)
prior_df <- data.frame(Q_cms = prior_samples)
prior_bins_plot <- ggplot(prior_df, aes(x = Q_cms)) +
  geom_density(fill = "lightblue", alpha = 0.5) +  # Density plot
  geom_vline(xintercept = log_bins, color = "black", lty="dashed") +  # Bin edges
  labs(title = "Binned log(flow) for K600 Priors", y = "Density", x = "log(Q) (cms)") +
  theme_bw()
prior_bins_plot

# Assign priors for binned K600
bayes_specs_new$K600_lnQ_nodes_meanlog <- rep(logK600_mean, length(log_bins))  # Centered at 25
bayes_specs_new$K600_lnQ_nodes_meanlog
bayes_specs_new$K600_lnQ_nodes_sdlog <- rep(logK600_sd, length(log_bins))  # Wide enough to allow 40
bayes_specs_new$K600_lnQ_nodes_sdlog 

bayes_specs_new$K600_lnQ_nodes_centers <- log_bins
bayes_specs_new$K600_lnQ_nodes_centers

# Keep other parameters
bayes_specs_new$K600_daily_sigma_sigma <- 0.05 

bayes_specs_new$n_chains <- c(3)
bayes_specs_new$n_cores <- c(3)
bayes_specs_new$burnin_steps <- c(2500)
bayes_specs_new$saved_steps <- c(2500)

bayes_specs_new
```


Plots made on filtered data: 
**`met.clean`** filtered for days with `GPP_daily_Rhat<1.05`,`ER_daily_Rhat<1.05`,
`K600_daily_Rhat <1.05`, as well as  `(GPP_97.5pct>0)` and  `(ER_2.5pct<0)`.

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=8, fig.height = 5}

met.clean <- met.full %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05)

mean_k_mod <- mean(met.clean$K600_daily_mean)
mean_k_mod
mean_k_measure <- mean(measured_K$K600)
mean_k_measure

KER_cor <- round(cor(met.clean$ER_daily_mean, met.clean$K600_daily_mean, use = "complete.obs"),3)
print(KER_cor)

KGPP_cor <-round(cor(met.clean$GPP_daily_mean, met.clean$K600_daily_mean, use = "complete.obs"),3)
print(KGPP_cor)
```



```{r, warning=F, size = 'small', echo=F, fig.align = "center", fig.width=8, fig.height = 5}
ker <- ggplot(met.clean, aes(x = K600_daily_mean, y = ER_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+
  theme_bw()+ geom_smooth(method="lm", se=F, color="grey50") +
  scale_color_viridis_d(option = "inferno") +  theme(legend.position = "null")


kgpp <- ggplot(met.clean, aes(y = K600_daily_mean, x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")

kyear <- met.clean %>%
ggplot(aes(x = as.factor(year), y = K600_daily_mean, fill = as.factor(lab), col = as.factor(lab)))+
    geom_hline(yintercept = mean_k_mod, lty = 2)+
  geom_boxplot(alpha = 0.8)+
  theme_bw()+
  scale_fill_viridis_d(option ="inferno") +  # Use _d for discrete
  scale_color_viridis_d(option = "inferno") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mod.env.ag <- input_dat%>%
  mutate(date= as.Date(solar.time))%>%
  group_by(date) %>%
  dplyr:: summarise(
    DO.obs = mean(DO.obs, na.rm=T),
    DO.sat = mean(DO.sat, na.rm=T),
    depth = mean(depth, na.rm=T),
    temp.water = mean(temp.water, na.rm=T),
    light = mean(light, na.rm=T),
    discharge =  mean(discharge, na.rm=T))


k.q <- met.clean %>%
  left_join(mod.env.ag, by = "date") %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = lab))+
  geom_point( size = 2)+ geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "inferno") 

p1 <- plot_grid(ker, kgpp, k.q, rel_widths =c(0.3, 0.3, 0.4), ncol = 3)
plot_grid(p1, kyear, ncol = 1)
```

The vertical dashed is the overall mean modeled K600 in the box plot. 



##### GPP and ER
```{r, warning=F, size = 'small', echo=F, fig.align = "center", fig.width=8, fig.height = 5}
GPP_E_cor <-round(cor(met.clean$GPP_daily_mean, met.clean$ER_daily_mean, use = "complete.obs"),3)
print(GPP_E_cor)


kgpp <- ggplot(met.clean, aes(y = (ER_daily_mean*-1), x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50",lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")
kgpp

```

## Final thoughts.
The direction of the K600 ~ flow relationship looks more logical, where K600 increases with flow.
ER and K600 are still very negatively correlated (-0.759) but it might just mean that ER is not an identifiable parameter in this reach. 

The mean modeled K600 is lower than expected, 16.9 when it should be closer to 25. Overall this could just be a challenging reach to fit the model to. 

### Final cleaned model output:
Where GPP is in blue and ER is in orange, and the black points represent NEP.


```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=6}
p <- met.clean %>%
  ggplot(aes(x = date)) +
  geom_abline(slope = 0, intercept = 0, linewidth = 1, alpha = 0.6) +
  geom_point(aes(y = GPP_daily_mean), col = "cyan4", alpha = 0.8) +
  geom_errorbar(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
                width = 0.2, color = "cyan4") +
  geom_point(aes(y = ER_daily_mean), col = "chocolate3", alpha = 0.8) +
  geom_errorbar(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
                width = 0.2, color = "chocolate3") +
  geom_point(aes(y = NEP_daily_mean), alpha = 0.5, size = 0.2) +
  theme_bw()

# Convert to Plotly
ggplotly(p)
```



### Table summarizing excluded data from cleaned model output: 
Of the 1037 days with DO observations 699 days were removed.

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=12}

total_days <- length(unique(na.omit(mod.env.ag$date)))

unrealistic_gpp <- met.full %>% 
  filter(GPP_97.5pct < 0 ) 

negative_GPP_day <- length(unique(unrealistic_gpp$date))

unrealistic_ER <- met.full %>% 
  filter(ER_2.5pct > 0 ) 
pos_ER_day <- length(unique(unrealistic_ER$date))

###
df_length <- mod.env.ag %>% 
  left_join(met.full)
num_na_GPP <- sum(is.na(df_length$GPP_mean))

df_length <- mod.env.ag %>% 
  left_join(met.clean)
num_na_clean <- sum(is.na(df_length$GPP_mean))

poor_gpp <- met.full %>% 
  filter(GPP_daily_Rhat > 1.05) 
poor_gpp_day <- length(unique(poor_gpp$date))

poor_ER <- met.full %>% 
  filter(ER_daily_Rhat > 1.05) 
poor_ER_day <- length(unique(poor_ER$date))

poor_K600 <- met.full %>% 
  filter(K600_daily_Rhat > 1.05) 
poor_K600_day <- length(unique(poor_K600$date))


pertot <- round((total_days/total_days)*100,3)
perna <- round((num_na_clean/total_days)*100,1)
pernaGPP <- round((num_na_GPP/total_days)*100,1)
perPoorGPP <- round((poor_gpp_day/total_days)*100,1)
perPoorER <- round((poor_ER_day/total_days)*100,1)
perPoorK600 <- round((poor_K600_day/total_days)*100,1)
perNegGPP <- round((negative_GPP_day/total_days)*100,1)
perPosER <- round((pos_ER_day/total_days)*100,1)


# Create table
excluded_summary <- tibble(
  `Number of Days` = c(total_days, num_na_clean, num_na_GPP,
                       poor_gpp_day, poor_ER_day, poor_K600_day,
                       negative_GPP_day, pos_ER_day), 
  `Explaination` = c("Total days of DO observations", "Total days removed", "Days model was unable to fit", "days where GPP rhat > 1.05", "days where ER rhat > 1.05", "days where K600 rhat > 1.05",
                     "days where modeled GPP was negative", "days where modeled ER was positive"),
    `Percent of Days` = c(pertot,
                          perna, 
                          pernaGPP,
                          perPoorGPP,
                          perPoorER, 
                          perPoorK600, 
                          perNegGPP,
                          perPosER))


# Print formatted table
excluded_summary %>%
  kable() %>%
  kable_styling(font_size = 15) %>%  
  row_spec(0:2, bold = TRUE, font_size = 16) %>% 
  row_spec(2, hline_after = TRUE)  
```


#### Session info

```{r, warning=F, size = 'small', echo=FALSE, fig.width=5, fig.height=12}
pander::pander(sessionInfo())
```

