---
title: 'Hydro-climate characteristics'
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```


```{r, include=FALSE}
library(snotelr)
library(tidyverse)
library(lubridate)
library(readr)
library(ggpubr)
library(ggplot2)
library(ggdist)
library(dataRetrieval)
library(lme4)
library(lmerTest)
library(MuMIn)

site_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#97b5c2",
  "GBL" = "#DD6E42",
  "GBU" = "#d9cb7c"
)


catch_colors <- c(
  "BW" = "#054fb9",
  "GB" = "#DD6E42"
)
```

## Calculating hydrology characteristics for stream flow data

### 1. Flow Duration Analysis

Bring in stream flow to look at snow to flow dynamics from USGS stream gage network ([streamstats](https://streamstats.usgs.gov/ss/))

```{r eval=TRUE, echo = TRUE, error=FALSE, warning=FALSE, message=FALSE, include=FALSE}
# identify and scrape data from USGS gages for target sites (BW and GB)

# Define the site numbers for "GB" and "BW"
siteNo_GB <- "10336730"
siteNo_BW <- "10336660"

# Define the parameter codes for flow and stage
pCode_flow <- "00060"
pCode_stage <- "00065"

# Set the start date to "1980-01-01" and end date to today (current date)
start.date <- "2003-09-30"
end.date <- "2024-10-30" #Sys.Date()  # Use the current date as the end date

# Ensure that dplyr functions are correctly used
flow_data_GB <- readNWISdata(siteNumbers = siteNo_GB, parameterCd = pCode_flow, startDate = start.date, endDate = end.date) %>%
  mutate(site = "GBL", catch = "GB") %>%
  dplyr::rename(date = dateTime, dischargeCFS = X_00060_00003) %>%
  mutate(
    dischargeCMS = dischargeCFS * 0.0283168,
    scale_Q = ((dischargeCFS * 0.0283168) / 10.64485)) %>%
    dplyr::select(date, dischargeCFS, site, catch, dischargeCMS, scale_Q)

# View the processed data
print(head(flow_data_GB))

flow_data_BW <- readNWISdata(siteNumbers = siteNo_BW, parameterCd = pCode_flow, startDate = start.date, endDate = end.date) %>%
  mutate(site = "BWL", catch = "BW") %>%
  rename(date = dateTime, dischargeCFS = X_00060_00003) %>%
  mutate(
    dischargeCMS = dischargeCFS * 0.0283168,
    scale_Q = ((dischargeCFS * 0.0283168) / 29.00787)
  ) %>%
  dplyr::select(date, dischargeCFS, site, catch, dischargeCMS, scale_Q)


# Combine data for both sites
flow_data <- rbind(flow_data_GB, flow_data_BW)
```

### A. Flow Duration Analysis: Estimating flow exceedance probabilities

```{r eval=TRUE, echo = TRUE, error=FALSE, warning=FALSE, message=FALSE, include=FALSE}
# Processing for both sites using pipes
result <- flow_data %>%
  filter(date>as.Date("2020-09-30")) %>%
  group_by(site) %>%
  dplyr::summarize(
    total_intervals = n(),
    ranked_scale_discharge = list(sort(scale_Q, decreasing = TRUE)),
    ranked_discharge = list(sort(dischargeCMS, decreasing = TRUE))
  ) %>%
  mutate(
    exceedence_prob = purrr::map(ranked_discharge, ~ seq(1, length(.), length.out = length(.)) / length(.) * 100)
  ) %>%
  unnest(cols = c(ranked_scale_discharge, ranked_discharge, exceedence_prob))
```


```{r, echo=FALSE, fig.width=8, fig.height=4, fig.align='center', inlcude=T}

# Plotting for each site
plot <- result %>%
  ggplot(aes(y = ranked_scale_discharge, x = exceedence_prob, color = site)) +
  geom_line(linewidth=1) + theme_classic() +
    scale_color_manual(values=alpha(c(site_colors), 0.75)) +
  labs(y = "Discharge (cms/km)", x = "Exceedence probability",
       title = "Exceedence Probability vs. Discharge by site") #+ facet_grid(.~site)
plot

```


```{r eval=TRUE, echo = TRUE, error=FALSE, warning=FALSE, message=FALSE, include=FALSE}
# Save the plot as a file (e.g., PNG)
# ggsave("exceedence_probability_vs_discharge.png", plot, width = 4, height = 3, dpi = 300)
```

#### Recap:

Blackwood appears to be flashier relative to glenbrook based on daily streamflow from 2020-2024.

### B. Baseflow separation based on EcoHydRology baseflow function:

<http://cran.nexr.com/web/packages/EcoHydRology/EcoHydRology.pdf>

```{r, BaseflowSeparation fxn, include=FALSE}
BaseflowSeparation <-
function(streamflow, filter_parameter=0.925, passes=3){
suppressWarnings(Ends<-c(1,length(streamflow))*rep(1,(passes+1))) # Start and end values for the filter function
suppressWarnings(AddToStart<-c(1,-1)*rep(1,passes))
btP<-streamflow##Previous pass's baseflow approximation
qft<-vector(length=length(streamflow))
bt<-vector(length=length(streamflow))
bt[1]<-if(streamflow[1]<quantile(streamflow,0.25)) streamflow[1] else mean(streamflow)/1.5
##Guess baseflow value in first time step.  
for(j in 1:passes){
for (i in (Ends[j]+AddToStart[j]):Ends[j+1]){
if ((filter_parameter*bt[i-AddToStart[j]]+((1-filter_parameter)/2)*(btP[i]+btP[i-AddToStart[j]]))>btP[i]){
bt[i]<-btP[i]
} else bt[i]<-filter_parameter*bt[i-AddToStart[j]]+((1-filter_parameter)/2)*(btP[i]+btP[i-AddToStart[j]])
qft[i]<-streamflow[i]-bt[i]
}
if (j<passes){
btP<-bt
bt[Ends[j+1]]<-if(streamflow[Ends[j+1]]<mean(btP))streamflow[Ends[j+1]]/1.2 else mean(btP)
## Refines the approximation of end values after the first pass
}
}
f <- data.frame(bt,qft)
return(f)
}
```

#### Estimate baseflow:

```{r, echo=F, include=FALSE}
# Blackwood #
flow_data_BW <- na.omit(flow_data_BW)

bfs<-BaseflowSeparation(flow_data_BW$scale_Q, passes=3)
bfs[,1] <- as.numeric(bfs[,1])

flow_data_BW <- cbind(flow_data_BW, bfs)

# Glenbrook #
bfs_GB<-BaseflowSeparation(flow_data_GB$scale_Q , passes=3)
bfs_GB[,1] <- as.numeric(bfs_GB[,1])

flow_data_GB <- cbind(flow_data_GB, bfs_GB)
```

### Visualize hydrograph and baseflow components:

##### Plots for Blackwood flow
```{r, echo=FALSE, fig.width=9, fig.height=4, fig.align='center', inlcude=F}
# Plotting for each site
bt_plot_BW <- ggplot(data = flow_data_BW) + 
  geom_line(aes(y = scale_Q, x = date, color = "Streamflow")) +
  geom_line(aes(y = bt, x = date, color = "Baseflow")) +
  geom_line(aes(y = qft, x = date, color = "Quickflow")) +
  labs(color = "Components", x = "Date (month-year)", y = expression(Flow~(m^3~s^-1~km^-1))) +
  scale_color_manual(values = c("Streamflow" = "black", "Baseflow" = "#B51963", "Quickflow" = "#0073E6")) +
  ggtitle("Blackwood Creek") +
  scale_x_datetime(date_breaks = "124 week", date_labels = "%b-%y") +
  theme_classic() +
  theme(legend.position = NULL)

# Remove legend
bt_plot_BW <- bt_plot_BW + guides(color = FALSE)


# get correlations between stream flow and bt as well as qft
BW_flow_bt <- lm(scale_Q~bt, data=flow_data_BW)
summary(BW_flow_bt)
r_gsquared_bt <- summary(BW_flow_bt)$r.squared

BW_flow_qft <- lm(scale_Q~qft, data=flow_data_BW)
summary(BW_flow_qft)
r_gsquared_qt <- summary(BW_flow_qft)$r.squared

flow_plot_BW <- ggplot(data=flow_data_BW) + 
   geom_point(data = flow_data_BW, aes(y = scale_Q, x = bt, color = "Baseflow, R² = 0.79"), alpha = 0.5) +
  geom_point(data = flow_data_BW, aes(y = scale_Q, x = qft, color = "Quickflow, R² = 0.89"), alpha = 0.5) +
  stat_smooth(data = flow_data_BW, aes(y = scale_Q, x = bt, color = "Baseflow, R² = 0.79"), method = "lm", se = FALSE) +
  stat_smooth(data = flow_data_BW, aes(y = scale_Q, x = qft, color = "Quickflow, R² = 0.89"), method = "lm", se = FALSE) +
    labs(color = "Components", x = expression("Component flow" ~(m^3~s^-1~km^-1)), y = expression("Streamflow" ~(m^3~s^-1~km^-1))) +
  scale_color_manual(values = c("Baseflow, R² = 0.79" = "#B51963", "Quickflow, R² = 0.89" = "#0073E6"))  + ggtitle(" ") + 
  theme_classic()+
  theme(legend.position = "bottom") 

```


```{r, echo=FALSE, fig.width=9, fig.height=6, fig.align='center', inlcude=F}
flow_plot_BW <- ggplot(data=flow_data_BW%>%mutate(year = year(date), 
                                                  DOY=yday(date)) %>% filter(year>2020)) + 
  geom_line(aes(y = scale_Q, x = DOY, color = "Streamflow")) +
  geom_line(aes(y = bt, x = DOY, color = "Baseflow")) +
  geom_line(aes(y = qft, x = DOY, color = "Quickflow")) +
  labs(color = "Components", x = "Date (month-year)", y = expression(Flow~(m^3~s^-1~km^-1))) +
  scale_color_manual(values = c("Streamflow" = "black", "Baseflow" = "#B51963", "Quickflow" = "#0073E6")) +
  ggtitle("Blackwood Creek") +
  #scale_x_datetime(date_breaks = "124 week", date_labels = "%b-%y") +
  theme_classic() + facet_grid(year~.)+
  theme(legend.position = NULL)

```


```{r, echo=FALSE, include=TRUE, fig.width=9, fig.height=4, fig.align='center'}

baseflow_plt <- ggarrange(bt_plot_BW, 
                     flow_plot_BW,
                     ncol = 2, nrow = 1,
                     common.legend = T, 
                     legend = "bottom",
                     widths = c(.75, 0.35))

```

```{r, echo=FALSE, fig.width=9, fig.height=4, fig.align='center', inlcude=F}
###
# ggsave(plot = baseflow_plt, filename = paste("./24_BW_flow_separation_plot.png",sep=""),width=10,height=4,dpi=300)
baseflow_plt

```

##### Plots for Glenbrook flow

```{r, echo=FALSE, fig.width=9, fig.height=4, fig.align='center', inlcude=F}

bt_plot_GB <- ggplot(data = flow_data_GB) + 
  geom_line(aes(y = scale_Q, x = date, color = "Streamflow")) +
  geom_line(aes(y = bt, x = date, color = "Baseflow")) +
  geom_line(aes(y = qft, x = date, color = "Quickflow")) +
  labs(color = "Components", x = "Date (month-year)", y = expression(Flow~(m^3~s^-1~km^-1))) +
  scale_color_manual(values = c("Streamflow" = "black", "Baseflow" = "#B51963", "Quickflow" = "#0073E6")) +
  ggtitle("Glenbrook Creek") +
  scale_x_datetime(date_breaks = "124 week", date_labels = "%b-%y") +
  theme_classic() +
  theme(legend.position = NULL)


# Remove legend
bt_plot_GB <- bt_plot_GB + guides(color = FALSE)

# get correlations between stream flow and bt as well as qft
GB_flow_bt <- lm(scale_Q~bt, data=flow_data_GB)
summary(GB_flow_bt)
GB_flow_qft <- lm(scale_Q~qft, data=flow_data_GB)
summary(GB_flow_qft)

flow_plot_GB <- ggplot(data=flow_data_GB) + 
   geom_point(data = flow_data_GB, aes(y = scale_Q, x = bt, color = "Baseflow, R² = 0.88"), alpha = 0.5) +
  geom_point(data = flow_data_GB, aes(y = scale_Q, x = qft, color = "Quickflow, R² = 0.95"), alpha = 0.5) +
  stat_smooth(data = flow_data_GB, aes(y = scale_Q, x = bt, color = "Baseflow, R² = 0.88"), method = "lm", se = FALSE) +
  stat_smooth(data = flow_data_GB, aes(y = scale_Q, x = qft, color = "Quickflow, R² = 0.95"), method = "lm", se = FALSE) +
      labs(color = "Components", x = expression("Component flow" ~(m^3~s^-1~km^-1)), y = expression("Streamflow" ~(m^3~s^-1~km^-1))) +
  scale_color_manual(values = c("Baseflow, R² = 0.88" = "#B51963", "Quickflow, R² = 0.95" = "#0073E6"))  + ggtitle(" ") + theme_classic()+
  theme(legend.position = "bottom") 



```

```{r, echo=FALSE, include=TRUE, fig.width=9, fig.height=4, fig.align='center'}
baseflow_plt <- ggarrange(bt_plot_GB, 
                     flow_plot_GB,
                     ncol = 2, nrow = 1,
                     common.legend = T, 
                     legend = "bottom",
                     widths = c(.75, 0.35))
baseflow_plt

```

```{r, echo=FALSE, fig.width=9, fig.height=4, fig.align='center', inlcude=F}

###
# ggsave(plot = baseflow_plt, filename = paste("./24_GB_flow_separation_plot.png",sep=""),width=10,height=4,dpi=300)
```

## 2. Hydro-climate characterization

Major question to address:

-   How do the watersheds filter surface water inputs to streams based on differences in local climate (precipitation events, snow-accumulation and snow-melt)?

##### Function for water year:

```{r eval=TRUE, echo = T, error=FALSE, warning=FALSE, message=FALSE}
# fxn for water year
water_year <- function(data) {
  data %>%
    mutate(date = ymd(date)) %>%
    mutate(WaterYear = if_else(month(date) >= 10, year(date) + 1, year(date)))
}
```

##### Read in data from SNOTEL

```{r eval=TRUE, echo = F, error=FALSE, warning=FALSE, message=FALSE}
# read in the sites data for all SNOTEL sites
target_snow_data <- snotelr::snotel_download(network = "sntl",
                                             site_id = c(848, 615), 
                                             metric = TRUE, internal = TRUE)
target_snow_data$date<- as.Date(target_snow_data$date)

# transform for water year:
target_snow_dat<- water_year(target_snow_data)

#Trim for columns 
target_snow_df <- target_snow_dat %>%
  filter(WaterYear>2003) %>%
  filter(date < as.Date("2024-10-10")) %>%
  mutate(site = case_when(
    site_id == 615 ~ "GBL",
    site_id %in% c(848) ~ "BWL")) %>%
  dplyr::select(site_name,site_id,date,snow_water_equivalent,precipitation_cumulative,
         temperature_max,temperature_min,temperature_mean,precipitation,WaterYear,
         site)

summary(target_snow_df)

```


## B. Melt and stream flow.

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}
snowflow_dat <- left_join(target_snow_df, 
                          flow_data, by = c("date", "site")) %>%
  mutate(dischargeCMD = scale_Q * 86400) %>%
  group_by(site, WaterYear)%>%
  arrange(date) %>%  
  mutate(accum_SWE = cumsum(snow_water_equivalent),
         accum_Qm3D = cumsum(dischargeCMD))
  
  




plot_SWE <- snowflow_dat %>%
  ggplot(aes(y = accum_SWE*0.001, x = date, color = site)) +
  geom_line(linewidth=1) + theme_classic() +
    scale_color_manual(values=alpha(c(site_colors), 0.75)) +
    ylab(expression(Accumulated~SWE~(m))) + xlab(NULL) 

plot_SWE

plot_flow <- snowflow_dat %>%
  ggplot(aes(y = accum_Qm3D, x = date, color = site)) +
  geom_line(linewidth=1) + theme_classic() +
    scale_color_manual(values=alpha(c(site_colors), 0.75)) +
    ylab(expression(Accumulated~flow~(m^3~km^-1~d^-1))) + xlab(NULL) 

plot_flow


plot_ppt <- snowflow_dat %>%
  ggplot(aes(y = precipitation_cumulative, x = date, color = site)) +
  geom_line(linewidth=1) + theme_classic() +
    scale_color_manual(values=alpha(c(site_colors), 0.75)) +
    ylab(expression(Accumulated~precipt~(mm~d^-1))) + xlab(NULL) 

plot_ppt
```


```{r, warning=F, echo=F,include=T, fig.width=8, fig.height=10}
grid <- ggarrange(plot_SWE,
                  plot_flow, 
                  plot_ppt,
                       ncol = 1, 
                       nrow = 3, 
                       common.legend = TRUE,
                       legend = "bottom")

grid
```

### Estimate values for: 

- Total accumulated SWE per water year in mm/yr

- Water yield (cms/km yr)

- Max SWE measurement mm

- Max precipitation accumulation (ppt) in mm 


```{r, echo=T, error=FALSE, warning=FALSE, include=T, message=FALSE}

snowflow_dat_sum <- snowflow_dat%>%
  filter(WaterYear<2025) %>%
  group_by(site, WaterYear) %>%
  summarise(Total_SWE = max(accum_SWE, na.rm=T),
            Water_yield = max(accum_Qm3D, na.rm = T),
            Max_SWE = max(snow_water_equivalent, na.rm=T),
            Total_ppt = max(precipitation_cumulative, na.rm=T),
            scale_pk_Q = max(scale_Q, na.rm = T)
            )

summary(snowflow_dat_sum)

### get 20 year water year averages
snowflow_HA <- snowflow_dat_sum%>%
  filter(WaterYear<2025) %>%
  group_by(site) %>%
  summarise(mean_SWE_acc = mean(Total_SWE, na.rm=T),
            mean_Water_yield = mean(Water_yield, na.rm = T),
            mean_SWE = mean(Max_SWE, na.rm = T),
            mean_ppt_acc = mean(Total_ppt, na.rm = T)
            )

snowflow_HA


snowflow_HA1 <- snowflow_dat_sum%>%
  filter(WaterYear>2020 & WaterYear<2024) %>%
  mutate(WY_lab = case_when(
    WaterYear %in% c(2021, 2022) ~ "dry",
    WaterYear %in% c(2023) ~ "wet",
    TRUE ~ NA_character_  )) %>%
    group_by(site,WY_lab) %>%
  summarise(mean_SWE_acc = mean(Total_SWE, na.rm=T),
            mean_Water_yield = mean(Water_yield, na.rm = T),
            mean_SWE = mean(Max_SWE, na.rm = T),
            mean_ppt_acc = mean(Total_ppt, na.rm = T),
            m_pk_Q= mean(scale_pk_Q, na.rm=T)
            )

snowflow_HA1

```


```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}

SWP_acc_plots <- ggplot(snowflow_dat_sum, aes(x = WaterYear, y = Total_SWE*0.001, color = site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.1), shape=15) +
  geom_line(position = position_dodge(width = 0.1))+
  geom_hline(yintercept = 98089.47*0.001, col = "#054fb9", lty = 2) + 
  geom_hline(yintercept = 65774.26*0.001, col = "#DD6E42", lty = 2) + 
  theme_classic() + ylab(expression(Annual~accumulated~SWE~(m~y^-1))) +
  xlab(NULL) + scale_color_manual(values = site_colors) +
  theme(legend.position =  "none") +
      scale_x_continuous(breaks = seq(floor(min(snowflow_dat_sum$WaterYear)), ceiling(max(snowflow_dat_sum$WaterYear)), by = 1), labels = function(x) as.character(round(x)))


WYeild_acc_plots <- ggplot(snowflow_dat_sum, aes(x = WaterYear, y = log(Water_yield), color = site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.1), shape=17) +
  geom_line(position = position_dodge(width = 0.1))+
    geom_hline(yintercept = log(1052085.5), col = "#054fb9", lty = 2) + 
  geom_hline(yintercept = log(149169.5), col = "#DD6E42", lty = 2) + 
  theme_classic() + 
    ylab(expression(atop("log(Water yield)", ~(m^3~km^-1~y^-1)))) +
  xlab(NULL) + scale_color_manual(values = site_colors) +
  theme(legend.position =  "none") +
        scale_x_continuous(breaks = seq(floor(min(snowflow_dat_sum$WaterYear)), ceiling(max(snowflow_dat_sum$WaterYear)), by = 1), labels = function(x) as.character(round(x)))


SWE_m_plots <- ggplot(snowflow_dat_sum, aes(x = WaterYear, y = Max_SWE*0.001, color = site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.1), shape=15) +
  geom_line(position = position_dodge(width = 0.1))+
    geom_hline(yintercept = 974.1571*0.001, col = "#054fb9", lty = 2) + 
  geom_hline(yintercept = 644.1905*0.001, col = "#DD6E42", lty = 2) + 
  theme_classic() + ylab(expression(Annual~max~SWE~(m~y^-1))) +
  xlab(NULL) + scale_color_manual(values = site_colors) +
  theme(legend.position =  "none") +
        scale_x_continuous(breaks = seq(floor(min(snowflow_dat_sum$WaterYear)), ceiling(max(snowflow_dat_sum$WaterYear)), by = 1), labels = function(x) as.character(round(x)))



ppt_accum_plots <- ggplot(snowflow_dat_sum, aes(x = WaterYear, y = Total_ppt*0.001, color = site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.1), shape=15) +
  geom_line(position = position_dodge(width = 0.1))+
    geom_hline(yintercept = 1718.2429*0.001, col = "#054fb9", lty = 2) + 
  geom_hline(yintercept = 843.7667*0.001, col = "#DD6E42", lty = 2) + 
  theme_classic() + 
  ylab(expression(atop("Accumulated PPT",~(m~y^-1)))) +
  xlab(NULL) + scale_color_manual(values = site_colors) +
  theme(legend.position =  "none") +
        scale_x_continuous(breaks = seq(floor(min(snowflow_dat_sum$WaterYear)), ceiling(max(snowflow_dat_sum$WaterYear)), by = 1), labels = function(x) as.character(round(x)))
  
  
  
```




```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}

ppt_accum_plots <- ggplot(snowflow_dat_sum, aes(x = WaterYear, y = Total_ppt*0.001, color = site, shape=site)) +
  annotate("rect", xmin = -Inf, xmax = 2020.5, ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.3) +
annotate("rect", xmin = 2024.5, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "grey80", alpha = 0.3) +
  geom_point(size = 1.5, alpha = 0.95, position = position_dodge(width = 0.1)) +
  geom_line(position = position_dodge(width = 0.1))+
    geom_hline(yintercept = 1718.*0.001, col = "#054fb9", lty = 2) + 
  geom_hline(yintercept = 844*0.001, col = "#DD6E42", lty = 2) + 
  theme_bw() + 
 # geom_vline(xintercept = 2021.5, linetype = "dotted", color = "grey30") +
#    geom_vline(xintercept = 2023.5, linetype = "dotted", color = "grey30") +
  ylab(expression("PPT"~(m~y^-1))) +
  xlab(NULL) + scale_color_manual(values = site_colors) +
  theme(legend.position =  "none") +
        scale_shape_manual(name = "Shore",
                     values = c("BWL" = 19, 
                                "GBL" = 17)) +
        scale_x_continuous(breaks = seq(floor(min(snowflow_dat_sum$WaterYear)), ceiling(max(snowflow_dat_sum$WaterYear)), by = 1), labels = function(x) as.character(round(x)))
  
  
  
```


```{r, warning=F, echo=F,include=T, fig.width=9.5, fig.height=6}
grid <- ggarrange(ppt_accum_plots,
                  WYeild_acc_plots, 
                       ncol = 1, 
                       nrow = 2, 
                        align = "v",
                       common.legend = TRUE,
                       legend = "right")

grid
```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("./Figs/Hydro_panel_long.png", plot = grid, width = 12, height = 3.25, units = "in")

```




## A. Calculate delta SWE

```{r eval=TRUE, echo = F, error=FALSE, warning=FALSE, message=FALSE, include=FALSE}
result <- snowflow_dat %>%
  filter(date > as.Date("2020-09-29"))%>%
  arrange(site, date) %>%
  group_by(site) %>%
  mutate(
    daily_delta_SWE = snow_water_equivalent - lag(snow_water_equivalent, default = 0),
    snow_water_equivalent_cumulative = cumsum(daily_delta_SWE),
    precipitation_quality = ifelse(daily_delta_SWE > 0, "Accumulate", "Melt"), # not really rain
    rain = ifelse(daily_delta_SWE < 0, 1, 0)  # again just a really simple proxy
  ) %>%
  ungroup()
```

Functions to identify max SWE per site for each water year and last day of snow pack persistence. 

```{r eval=TRUE, echo = F, error=FALSE, warning=FALSE, message=FALSE, include=FALSE}
summarize_peak_SWE_date <- function(data) {
  data %>%
    filter(lubridate::month(date) >= 1 & lubridate::month(date) <= 6) %>%
    arrange(site, date) %>%
    group_by(site, WaterYear = lubridate::year(date) + if_else(lubridate::month(date) < 10, 0, 1)) %>%
    summarize(
      peak_SWE_date = date[which.max(snow_water_equivalent)],
      peak_SWE_day_of_year = lubridate::yday(date[which.max(snow_water_equivalent)]),
      max_SWE = max(snow_water_equivalent)
    ) %>%
    ungroup()
}

summarize_transition_to_zero <- function(data) {
  data %>%
    filter(lubridate::month(date) >= 4 & lubridate::month(date) <= 7) %>%
    arrange(site, date) %>%
    group_by(site, WaterYear = lubridate::year(date) + if_else(lubridate::month(date) < 10, 0, 1)) %>%
    summarize(
      transition_to_zero = date[which.max(
        (daily_delta_SWE == 0 & lead(daily_delta_SWE == 0, default = FALSE) &
         lead(daily_delta_SWE == 0, n = 8, default = FALSE) &
         lag(daily_delta_SWE, default = 0) == 0) &
         (lead(daily_delta_SWE == 0, n = 8) & lag(daily_delta_SWE == 0, n = 5))
      )],
      .groups = 'drop',
      trans_day_of_year = lubridate::yday(transition_to_zero),
    ) %>%
    filter(!is.na(transition_to_zero))
}

summarize_peak_SPflow_date <- function(data) {
  data %>%
    filter(lubridate::month(date) >= 2 & lubridate::month(date) <= 8) %>%
    dplyr::group_by(site, WaterYear = lubridate::year(date) + if_else(lubridate::month(date) < 10, 0, 1)) %>%
    dplyr::summarize(
      peak_SPFlow_date = date[which.max(scale_Q)],  # Date of the highest scale_Q
      peak_flow_day_of_year = lubridate::yday(date[which.max(scale_Q)]),  # DOY of the highest scale_Q
      max_SPFlow = max(scale_Q)  # Maximum scale_Q for the given year
    ) %>%
    ungroup()
}
```

```{r eval=TRUE, echo = F, error=FALSE, warning=FALSE, message=FALSE, include=T}

result1 <- summarize_transition_to_zero(result)
result1
```


End of script. 
