---
title: 'Analysis for Mountain stream ecosystem metabolism and nitrogen cycling responses to hydroclimatic volatility '
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
#setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWL_k600/")
lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB", "plotly", "MASS",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
```

### MS: Spatiotemporal variation in mountain stream metabolism and nitrogen cycling across contrasting flow regimes

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
daoc_date<- Sys.Date()

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
detach("package:plyr", unload = TRUE)
site_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#97b5c2",
  "GBL" = "#DD6E42",
  "GBU" = "#d9cb7c"
)

siteC_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#054fb9",
  "GBL" = "#DD6E42",
  "GBU" = "#DD6E42"
)

catch_colors <- c(
  "BW" = "#054fb9",
  "GB" = "#DD6E42"
)



### fxns:
## Fxn for water year:
water_year <- function(data) {
  # Ensure the `date` column exists
  if (!"date" %in% names(data)) {
    stop("The dataframe must contain a column named 'date'.")
  }
  
  # Convert the `date` column to Date format (if not already)
  data$date <- as.Date(data$date)
  
  # Add the water_year column
  data$water_year <- with(data, {
    year <- as.numeric(format(date, "%Y"))
    month <- as.numeric(format(date, "%m"))
    ifelse(month >= 10, year + 1, year)
  })
  
  return(data)
}
## Function to perform ANOVA and post-hoc test within each site
comparisons <- list(
  c("2021", "2022"),
  c("2021", "2023"),
  c("2022", "2023"),
  c("2023", "2024")
)
```


```{r, warning=F, size = 'small', echo=F, include=F}

covariat_datq <- readRDS("./complied_data_JGR_MS_metab_26.RDS")
```

```{r, warning=F, size = 'small', echo=F, include=F}

nitrogen_data <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/analysis/Mtn-stream-biogeochem-productivity/EGRET_infill_WC.RDS")

```


```{r, warning=F, size = 'small', echo=F, fig.width=10, fig.height=12}
# Separate out BW 
covariat_datq_BW <- covariat_datq%>%
  dplyr::filter(site=="BWL" | site=="BWU")

# Separate out GB 
covariat_datq_GB <- covariat_datq%>%
  dplyr::filter(site=="GBL" | site=="GBU")
```

##### Kernal density plots of GPP and ER

Colored by years

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=7, fig.height=8}
library(ggplot2)

BWL_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWL" & water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5, shape = 15)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
   scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94", "#fad311")) +
#   scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
          annotate("text", x = 7, y = -28, label = "Lower reach\nlarge creek",size =4, fontface = "bold", hjust = 0.5) 
# 

BWU_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWU" & water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5, shape=0)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94", "#fad311")) +
#   scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
            annotate("text", x = 7, y = -28, label = "Upper reach\nlarge creek", size = 4, fontface = "bold", hjust = 0.5) 



GBU_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBU") %>%
filter(water_year<2025)%>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# Compute 2D density
density_data <- with(GBU_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))

# Convert density data to dataframe
density_df <- data.frame(
  expand.grid(x = density_data$x, y = density_data$y),
  z = as.vector(density_data$z)
)

# Plot with contour lines
GBU_KD_plot <- ggplot() +
  geom_point(data = GBU_DF_clean, aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5, shape=2) +
    geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 5, size=1) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   # ylab(NULL) +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
    scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
    scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94","#fad311" )) +

 # scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 7, y = -28, label = "Upper reach\nsmall creek", size = 4, fontface = "bold", hjust = 0.5) 

GBU_KD_plot


GBL_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBL") %>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)


# Plot with contour lines
GBL_KD_plot <- ggplot() +
#  geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 10) +
  geom_point(data = GBL_DF_clean%>%filter(water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5, shape=17) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
  # xlab(NULL) +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
 #scale_color_viridis_d(option = "viridis",  name = "Water year") +
    scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94","#fad311")) +
  theme_classic() +
      annotate("text", x = 7, y = -28, label = "Lower reach\nsmall creek", size = 4, fontface = "bold", hjust = 0.5) 


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4, fig.height=10}
# Modify each plot to rename legend title & adjust legend formatting
BWL_KD_plot <- BWL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

BWU_KD_plot <- BWU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBL_KD_plot <- GBL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBU_KD_plot <- GBU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))
```

###  *Figure 2*

GPP and ER regimes at each reach. 

```{r, warning=F, size = 'small', echo=F, fig.width=9, fig.height=7}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  BWL_KD_plot,
  BWU_KD_plot,
  GBL_KD_plot,
  GBU_KD_plot,
  labels = c("a", "b", "c", "d"),
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom"
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("./Fig3_KD_CH1_grid_rev.png", plot = KD_coeff_grid, width = 6.25, height = 6.25, units = "in")

```



### *Figure 3*

Setting: Showing the delayed onset of baseflow in wetter years.
Color as water year. 

DOY - April 1 = 91
DOY - May 1 = 121
DOY -June 1 = 152
DOY - Sept 1 = 244
DOY - Oct 1 = 274

```{r, warning=F, size = 'small', echo=F, include=FALSE}

covariat_datq<-covariat_datq%>%
  mutate(yday=yday(date))


overall_pf<- covariat_datq %>%
    filter(yday >= 92, yday <= 153) %>%
  group_by(site, water_year) %>%
  summarise(
    max_pf = max(Q_m, na.rm = TRUE),
    min_pf = min(Q_m, na.rm = TRUE),
    max_pf_yday = yday[which.max(Q_m)],
    min_pf_yday = yday[which.min(Q_m)],
    mean_pf = mean(Q_m, na.rm = TRUE),
    mean_pf_yday = yday[which.min(abs(Q_m - mean_pf))]) %>%
  ungroup()


overall_bf<- covariat_datq %>%
    filter(yday >= 170, yday <= 274) %>%
  group_by(site) %>%
  summarise(
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q = mean(Q_m, na.rm = TRUE),
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]) %>%
  ungroup()


overall_bf_df <- covariat_datq %>%
  left_join(overall_bf, by = c("site"))


steepest_slopes <- overall_bf_df %>%
    filter(yday >= 153, yday <= 274) %>%
  #  arrange(site, water_year, yday) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2023 ~ mean_q + 0.031,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2024 ~ mean_q + 0.01,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
    mutate(
    Q_m = case_when(
      site == "GBL" & water_year == 2021 ~ Q_m + 0.001,  # Adjust ave_q for GBL in 2021
      TRUE ~ Q_m)) %>%
  group_by(site, water_year) %>%
  summarise(
    ave_q = mean(mean_q, na.rm = TRUE),
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]
  ) %>%
  ungroup() 

```

```{r, warning=F, size = 'small', echo=F, include=F}
# Merge covariat_datq with steepest_slopes to get mean_q_yday
Baseflowplots <- covariat_datq %>%
  dplyr::select(date, site, catch, water_year, Q_m, yday) %>%
  left_join(steepest_slopes %>% 
              dplyr::select(site, water_year, mean_q_yday, ave_q), by = c("site", "water_year","yday"= "mean_q_yday")) %>%
  left_join(overall_pf %>% 
              dplyr::select(site, water_year, max_pf, max_pf_yday), by = c("site", "water_year", "yday"= "max_pf_yday"))  %>%filter(site=="BWL" | site=="GBL")


Baseflowplot_df <- Baseflowplots%>%
  mutate(
    ave_q = case_when(
      site == "GBL" & water_year == 2021 ~ ave_q - 0.006,  # Adjust ave_q for GBL in 2021
      TRUE ~ ave_q))


Baseflowplot <- ggplot(Baseflowplot_df %>% 
                         filter(water_year < 2025), 
                       aes(x = yday, y = Q_m, color = as.factor(water_year))) +
  geom_line() +  
  scale_color_viridis_d(name = "Year") +  
  xlim(91, 234) + 
  facet_wrap(.~site, scales = "free") + 
  theme_classic() + 
  ylab(expression(Streamflow~(m^3~s^-1))) +
  xlab("Day of year") +

  # Vertical lines at ave_q day
  geom_vline(data = Baseflowplot_df %>% filter(water_year < 2025),
             aes(xintercept = yday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) +

  # Points for max_pf and ave_q
  geom_point(aes(x = yday, y = ave_q, color = as.factor(water_year)), 
             shape = 19, size = 1.5, alpha = 0.9)

BF_sum <- Baseflowplot_df%>%
  filter(!is.na(Baseflowplot_df$ave_q))

BF_sum

```


Read in data from SNOTEL to get SWE timesieries 

```{r eval=TRUE, echo = F, error=FALSE, warning=FALSE, message=FALSE}
# read in the sites data for all SNOTEL sites
target_snow_data <- snotelr::snotel_download(network = "sntl",
                                             site_id = c(848, 615), 
                                             metric = TRUE, internal = TRUE)
target_snow_data$date<- as.Date(target_snow_data$date)

# transform for water year:
target_snow_dat<- water_year(target_snow_data)

#Trim for columns 
target_snow_df <- target_snow_dat %>%
  filter(water_year>2019) %>%
  filter(date < as.Date("2024-10-10")) %>%
  mutate(site = case_when(
    site_id == 615 ~ "GBL",
    site_id %in% c(848) ~ "BWL")) %>%
  dplyr::select(site_id,
                site,
                date,
                water_year,
                precipitation,
                snow_water_equivalent,
                precipitation_cumulative) %>%
  mutate(SWE = 0.001 *snow_water_equivalent)

summary(target_snow_df)

```

```{r, warning=F, size = 'small', echo=F}

Baseflowplot_df$date <- as.Date(Baseflowplot_df$date)
target_snow_df$date  <- as.Date(target_snow_df$date)

Baseflowplot_df <- Baseflowplot_df %>%
  left_join(
    target_snow_df %>% dplyr::select(site, date, SWE),
    by = c("site", "date")
  )
```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=9, fig.height=4.25}

library(dplyr)
library(ggplot2)

df_BWL<- Baseflowplot_df %>%
  filter(site=="BWL")
  
draw_order <- c(2023, 2024, 2022, 2021)

Baseflowplot_BWL <- local({

  df_BWL <- Baseflowplot_df %>%
    filter(water_year < 2025, site == "BWL") %>%
    mutate(
      label = "Large creek",
      year_level = factor(water_year, levels = draw_order)
    ) %>%
    arrange(factor(water_year, levels = draw_order))

  df <- df_BWL %>%
    filter(between(yday, 91, 234))

  # Use only the plotted window to get sensible scaling
  Q_top <- max(df$Q_m, na.rm = TRUE) * 1.35
  SWE_max <- max(df$SWE, na.rm = TRUE)
  k <- Q_top / SWE_max

  df <- df %>%
    mutate(
      SWE_ymin = Q_top - k * SWE * 0.5,
      SWE_ymax = Q_top
    )

  ggplot(df, aes(x = yday)) +

    geom_rect(
      aes(
        xmin = yday - 0.5, xmax = yday + 0.5,
        ymin = SWE_ymin, ymax = SWE_ymax,
        fill = year_level
      ),
      alpha = 0.75,
      inherit.aes = FALSE
    ) +

    geom_line(aes(y = Q_m, color = year_level), linewidth = 0.9) +

    geom_vline(
      data = df_BWL,
      aes(xintercept = yday[ave_q == ave_q], color = year_level),
      linetype = "dashed", linewidth = 0.5, alpha = 0.9
    ) +

    facet_grid(label ~ ., scales = "free") +

    scale_color_manual(
      values = c("2021" = "#440154", "2022" = "#3B528B", "2023" = "#4bcc94", "2024" = "#fad311"),
      name = "Water year"
    ) +
    scale_fill_manual(
      values = c("2021" = "#440154", "2022" = "#3B528B", "2023" = "#4bcc94", "2024" = "#fad311"),
      name = "Water year"
    ) +

    scale_y_continuous(
      limits = c(0, Q_top),
      name = expression(Streamflow~(m^3~s^-1)),
      sec.axis = sec_axis(
        trans = ~ (Q_top - .) / k,
        name  = "Snow water equivalent (m)"
      )
    ) +

    xlab("Day of year") +
    theme_classic() +
    theme(legend.position = "bottom") +

    geom_text(aes(x = 187, y = 11.5, label = "51 days"),
              color = "#440154", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
    geom_text(aes(x = 207, y = 10.25, label = "30 days"),
              color = "#3B528B", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
    geom_text(aes(x = 210, y = 8.5, label = "27 days"),
              color = "#fad311", size = 4, hjust = 1, vjust = 1, fontface = "bold")
})

Baseflowplot_BWL

```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=9, fig.height=4.25}


library(dplyr)
library(ggplot2)

df_GBL<- Baseflowplot_df %>%
  filter(site=="GBL")

draw_order <- c(2023, 2024, 2022, 2021)

Baseflowplot_GBL <- local({

  df_GBL <- Baseflowplot_df %>%
    filter(water_year < 2025, site == "GBL") %>%
    mutate(
      label = "Small creek",
      year_level = factor(water_year, levels = draw_order)
    ) %>%
    arrange(factor(water_year, levels = draw_order))

  df <- df_GBL %>%
    filter(between(yday, 91, 234))

  # Use only the plotted window to get sensible scaling
  Q_top <- max(df$Q_m, na.rm = TRUE) * 1.35
  SWE_max <- max(df$SWE, na.rm = TRUE)
  k <- Q_top / SWE_max

  df <- df %>%
    mutate(
      SWE_ymin = Q_top - k * SWE * 0.5,
      SWE_ymax = Q_top
    )

  ggplot(df, aes(x = yday)) +

    geom_rect(
      aes(
        xmin = yday - 0.5, xmax = yday + 0.5,
        ymin = SWE_ymin, ymax = SWE_ymax,
        fill = year_level
      ),
      alpha = 0.75,
      inherit.aes = FALSE
    ) +

    geom_line(aes(y = Q_m, color = year_level), linewidth = 0.9) +

    geom_vline(
      data = df_GBL,
      aes(xintercept = yday[ave_q == ave_q], color = year_level),
      linetype = "dashed", linewidth = 0.5, alpha = 0.9
    ) +

    facet_grid(label ~ ., scales = "free") +

    scale_color_manual(
      values = c("2021" = "#440154", "2022" = "#3B528B", "2023" = "#4bcc94", "2024" = "#fad311"),
      name = "Water year"
    ) +
    scale_fill_manual(
      values = c("2021" = "#440154", "2022" = "#3B528B", "2023" = "#4bcc94", "2024" = "#fad311"),
      name = "Water year"
    ) +

    scale_y_continuous(
      limits = c(0, Q_top),
      name = expression(Streamflow~(m^3~s^-1)),
      sec.axis = sec_axis(
        trans = ~ (Q_top - .) / k,
        name  = "Snow water equivalent (m)"
      )
    ) +

    xlab("Day of year") +
    theme_classic() +
    theme(legend.position = "bottom") +
  geom_text(
            aes(x = 179, y = 0.88, label = "42 days"),  # Adjusted to max Q_m and max yday for top-right corner
            color = "#440154", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
  geom_text(
            aes(x = 187, y = 0.78, label = "38 days"),  # Adjust the position and text as needed
            color = "#3B528B", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
  geom_text(
            aes(x = 198, y = 0.65, label = "29 days"),  # Adjust the position and text as needed
            color = "#fad311", size = 4, hjust = 1, vjust = 1, fontface = "bold")
})

Baseflowplot_GBL


```


```{r, warning=F, size = 'small', echo=F, fig.width=9, fig.height=6.25}

baseflow_grid <- ggarrange(
  Baseflowplot_BWL,
  Baseflowplot_GBL,
  labels = c("a", "b"),
  ncol = 1, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom"
)


baseflow_grid
```



```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig2_baseflow_combined.png", plot = baseflow_grid, width =5.5, height = 6.5, units = "in")

```

#### Calculate nitrogen demand:

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
# Constants
ra <- 0.5  # Autotrophic respiration coefficient (Hall & Tank 2003)
C_Nauto <- 16  # Autotrophic C:N ratio (Stelzer & Lamberti 2001)
C_Nhetero <- 20  # Heterotrophic C:N ratio (Hall & Tank 2003)
HGE <- 0.05  # Heterotrophic Growth Efficiency (Hall & Tank 2003)

# Calculate components of nitrogen demand
covariat_ndemand <- covariat_datq %>%
  group_by(site) %>% 
  mutate(
    # Autotrophic respiration (raGPP)
    raGPP = GPP_mean * ra,
    # Autotrophic assimilation of N
    Auto_N_assim = GPP_mean / C_Nauto,
    # Heterotrophic respiration (Rh)
    Rh = ER_mean - raGPP,
    # Heterotrophic assimilation of N
    Hetero_N_assim = (Rh * HGE) / C_Nhetero,
    # Total nitrogen demand
    Ndemand = Auto_N_assim + Hetero_N_assim, # unit should be g N m-2 d-1
    Ndemand = if_else(Ndemand < 0, 0.0001, Ndemand)) %>%
  dplyr::select(site, date, Ndemand)
```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

### Quality check on supply calculations:
# (1) How many data points do we have for velocity (v) and width (w)?

morph_counts <- covariat_datq %>%
  group_by(site) %>%
  summarise(
    w_m_count = sum(!is.na(w_m)), 
    v_m_count = sum(!is.na(v_m))
  )

morph_counts
```


```{r, warning=F, size = 'small', echo=F, include=F,fig.width=6, fig.height=4}
##### better way to infill width and velocites to get at reach length:
# na.approx, and using Q as "Variables to be used for interpolation as in"

covariat_datqI <- covariat_datq %>%
  group_by(site) %>%
  arrange(date) %>%
  mutate(
    # Fit a linear model for w_m ~ Q_m using non-missing values
    w_m_apr = ifelse(is.na(w_m), predict(lm(w_m ~ depth_m, data = cur_data(), na.action = na.omit), newdata = cur_data()), w_m),
    v_m_apr = ifelse(is.na(v_m), predict(lm(v_m ~ Q_m, data = cur_data(), na.action = na.omit), newdata = cur_data()), v_m)
  ) %>%
  ungroup()

covariat_datqI <- covariat_datqI %>%
  group_by(site) %>%
  arrange(date) %>%  # Ensure chronological order
   mutate( ## 17 NAs
    w_m_apr = ifelse(is.na(w_m_apr),
                  rollapplyr(w_m_apr, width = 15, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE),
                  w_m_apr),
    v_m_apr = ifelse(is.na(v_m_apr),
                  rollapplyr(v_m_apr, width = 15, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE),
                  v_m_apr))
```



#### Calculate nitrogen supply:

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
covariat_nsupply <- covariat_datqI %>%
  left_join(nitrogen_data, by=c("date", "site")) %>%
  mutate(
    K600_daily_mean = case_when(
      is.na(K600_daily_mean) & site == "BWL" ~ 22, 
      is.na(K600_daily_mean) & site == "BWU" ~ 16,
      is.na(K600_daily_mean) & site == "GBU" ~ 32,
      is.na(K600_daily_mean) & site == "GBL" ~ 25,
      TRUE ~ K600_daily_mean)) %>%
  mutate(Q_Ls = Q_m * 1000, # flow from csm to Ls
  # calculate reach length in m
  reachL = c((v_m_apr*w_m_apr*8640)/K600_daily_mean), # seconds to days
   # Calculate no3 supply 
  NO3_supply = c(((86400*Q_Ls*(NO3_mgL_dl_sw/1000))/(w_m_apr*reachL))),
  NO3_supply_new = c(((86400*Q_Ls*(NO3_mgL_i/1000))/(w_m_apr*reachL))),

   # Calculate nh3 supply 
  NH4_supply = c(((86400*Q_Ls*(NH4_mgL_dl_sw/1000))/(w_m_apr*reachL))), ## unit should be g N m-2 d-1
 NH4_supply_new = c(((86400*Q_Ls*(NH4_mgL_i/1000))/(w_m_apr*reachL))),
  PO4_supply = c(((86400*Q_Ls*(PO4_ugL_dl_sw/1e-6))/(w_m_apr*reachL))) ## unit should be g N m-2 d-1
   )  %>%
  dplyr::select(site, date, reachL, Q_Ls, NO3_supply, 
                NO3_supply_new, NH4_supply, PO4_supply, NH4_supply_new, NO3_mgL_i, NH4_mgL_i, PO4_ugL_dl_sw)

```



```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
ndyn <- covariat_datq %>%
  left_join(covariat_nsupply, by=c("site", "date"))%>%
  left_join(covariat_ndemand, by=c ("site", "date")) 

summary(ndyn)

```

```{r, warning=F, size = 'small', echo=F,include=F, fig.width=9, fig.height=10}

TS_plot_w <- ggplot(covariat_datqI, aes(x = date, y = w_m_apr, color=site, shape =site)) +
  geom_point(size = 2, alpha = 0.5, position = position_dodge(width = 0.3)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_point(aes(x = date, y = w_m), color="black") +
  scale_color_manual(values = siteC_colors) +
  scale_shape_manual(values = c(15, 0, 17, 2)) + theme_classic() +
  facet_grid(site~.)

TS_plot_v <- ggplot(covariat_datqI, aes(x = date, y = v_m_apr, color=site, shape =site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_point(size = 2, alpha = 0.7) +
    geom_point(aes(x = date, y = v_m), color="black") +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
    scale_color_manual(values = siteC_colors) +
  theme_classic() +facet_grid(site~.)

n_s_grid <- ggarrange(
  TS_plot_w,
  TS_plot_v,
  labels = c("a", "b"),
  ncol = 1, nrow = 2,
  label.x = 0.95,  # Move label to the right
  label.y = 1,     # Keep label at the top
  hjust = 1,       # Right-align the label
  vjust = 1        # Top-align the label
)

n_s_grid

# Where the black points are the survey measures of width or velocity.

```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
ndyn <- covariat_datq %>%
  left_join(covariat_nsupply, by=c("site", "date"))%>%
  left_join(covariat_ndemand, by=c ("site", "date")) 

summary(ndyn)

```




```{r, warning=F, size = 'small', echo=F, include=FALSE}

names(covariat_datq)

covariat_datq<-covariat_datq%>%
  mutate(yday=yday(date))


overall_pf<- covariat_datq %>%
    filter(yday >= 92, yday <= 153) %>%
  group_by(site, water_year) %>%
  summarise(
    max_pf = max(Q_m, na.rm = TRUE),
    min_pf = min(Q_m, na.rm = TRUE),
    max_pf_yday = yday[which.max(Q_m)],
    min_pf_yday = yday[which.min(Q_m)],
    mean_pf = mean(Q_m, na.rm = TRUE),
    mean_pf_yday = yday[which.min(abs(Q_m - mean_pf))]) %>%
  ungroup()


overall_bf<- covariat_datq %>%
    filter(yday >= 170, yday <= 274) %>%
  group_by(site) %>%
  summarise(
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q = mean(Q_m, na.rm = TRUE),
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]) %>%
  ungroup()


overall_bf_df <- covariat_datq %>%
  left_join(overall_bf, by = c("site"))


steepest_slopes <- overall_bf_df %>%
    filter(yday >= 153, yday <= 274) %>%
  #  arrange(site, water_year, yday) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2023 ~ mean_q + 0.031,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2024 ~ mean_q + 0.01,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
    mutate(
    Q_m = case_when(
      site == "GBL" & water_year == 2021 ~ Q_m + 0.001,  # Adjust ave_q for GBL in 2021
      TRUE ~ Q_m)) %>%
  group_by(site, water_year) %>%
  summarise(
    ave_q = mean(mean_q, na.rm = TRUE),
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]
  ) %>%
  ungroup() 

```




##### Plots of Ratios of NH4+- N supply to demand, ratios NO3-- N supply to demand

Color is now also by water year 2021-2024, and shape is reach location.

```{r, warning=F, size = 'small', echo=F, include=F}
covariat_NS_ND <- ndyn %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(date, jday, Q_m, w_m, site, catch, water_year, 
                Ndemand, NO3_supply_new, NO3_supply, NH4_supply, NH4_supply_new,
                NH4_mgL_i, NO3_mgL_i) %>%
  mutate(
    NO3_SupDem = NO3_supply / Ndemand,
    NH4_SupDem = NH4_supply / Ndemand,
    N_SupDem = (NH4_supply + NO3_supply) / Ndemand
  ) %>%
  distinct(date, site, .keep_all = TRUE)  %>%
    mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",
    TRUE ~ NA_character_  
  )) %>%
    group_by(site, date, catch) %>%
    summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

covariat_NS_ND1 <- covariat_NS_ND%>%
  left_join(Baseflowplot_df, by=c("site", "date", "catch",
                                "water_year", "Q_m"))
  


NO3_supp_plots <- ggplot(covariat_NS_ND1%>%filter(water_year<2025 & jday<274), aes(y = (NO3_supply_new), x = jday, color = as.factor(water_year), shape=site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_point( aes(y = log(NH4_supply_new+1), x = jday, color = as.factor(water_year), shape=site), 
                size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
    scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
  scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
  ylim(0,1.50) +
   xlab(NULL) +
       scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94", "#fad311")) +
      scale_fill_manual(values=c("#440154", "#3B528B", "#4bcc94", "#fad311")) +
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10)) +
  geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
     ylab(expression(Supply~(g~N~m^-2~d^-1))) +
  theme_classic() +
  scale_shape_manual(values = c(15,0,17,2)) +
    theme(legend.position = "none") +
    geom_vline(data = covariat_NS_ND1 %>% filter(water_year < 2025),
             aes(xintercept = jday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) +
   facet_grid(catch~.) 

# Remove legend

# Define threshold for high NO3 demand (above 3rd quartile)
high_Ndemand_threshold <- 0.075

N_demand_plot1 <- ggplot(covariat_NS_ND1%>%filter(water_year<2025), aes(y = (Ndemand), x = jday, color = as.factor(water_year), shape = site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.8)) +
 # scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
      scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94",  "#fad311")) +
      scale_fill_manual(values=c("#440154", "#3B528B", "#4bcc94",  "#fad311")) +
  #  scale_fill_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
  xlab("Day of year") + 
    geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
    scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                         labels = scales::label_number()) +
 # geom_hline(yintercept = high_Ndemand_threshold, linetype = "dashed", color = "grey25") +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  ylab(expression(Demand~(g~N~m^-2~d^-1))) +
  theme_classic() +
  theme(legend.position = "none")  +
    facet_grid(catch~.) +
    geom_vline(data = covariat_NS_ND1 %>% filter(water_year < 2025),
             aes(xintercept = yday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) 


N_demand_plot1

```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=9, fig.height=4}
# ggsave("./Nsupply_grid_CH1_SFS_ppt_21.png", plot = n_s_grid, width = 8.5, height = 7.75, units = "in")

```



```{r, warning=F, size = 'small', echo=F, include=F}
covariat_NS_ND <- ndyn %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(date, jday, Q_m, w_m, site, catch, water_year, 
                Ndemand, NO3_supply_new, NO3_supply, NH4_supply, NH4_supply_new,
                NH4_mgL_i, NO3_mgL_i) %>%
  mutate(
    NO3_SupDem = NO3_supply / Ndemand,
    NH4_SupDem = NH4_supply / Ndemand,
    N_SupDem = (NH4_supply + NO3_supply) / Ndemand
  ) %>%
  distinct(date, site, .keep_all = TRUE)  %>%
    mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",
    TRUE ~ NA_character_  
  )) %>%
    group_by(site, date, catch) %>%
    summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

N_ratio_plots <- ggplot(covariat_NS_ND%>%filter(water_year<2025), aes(x = (NO3_supply), y = (Ndemand), color = as.factor(water_year), shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    geom_point( aes(x = (NH4_supply), y = Ndemand, color = as.factor(water_year)),
                size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
    scale_color_viridis_d(option = "viridis",  name = "Water year") +
      scale_shape_manual(values = c(15,0,17,2), name = "Site") +
    scale_x_continuous(trans = "log", breaks = scales::log_breaks(base = 10)) + 
  scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                       labels = scales::label_number()) +
   ylab(expression(log(Nitrogen~demand)~(g~N~m^-2~d^-1))) +
     xlab(expression(log(Nitrogen~supply)~(g~N~m^-2~d^-1))) +
  theme_classic() 


NO3_supp_plots <- ggplot(covariat_NS_ND%>%filter(water_year<2025), aes(y = (NO3_supply_new), x = jday, color = as.factor(water_year), shape=site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_point( aes(y = log(NH4_supply_new+1), x = jday, color = as.factor(water_year), shape=site), 
                size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
    scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
#  ylim(0,1.50) +
   xlab(NULL) +
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10)) +
  geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
     ylab(expression(log(Supply)~(g~N~m^-2~d^-1))) +
  theme_classic() +
      scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
      scale_fill_viridis_d(option = "viridis", name = "Water year") + 
  scale_shape_manual(values = c(15,0,17,2)) +
    theme(legend.position = "none") +
  facet_grid(catch~.)


# Define threshold for high NO3 demand (above 3rd quartile)
high_Ndemand_threshold <- 0.075

N_demand_plot1 <- ggplot(covariat_NS_ND%>%filter(water_year<2025), aes(y = (Ndemand), x = jday, color = as.factor(water_year), shape = site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.8)) +
  scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
    scale_fill_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
  xlab("Day of year") + 
    geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                         labels = scales::label_number()) +
 # geom_hline(yintercept = high_Ndemand_threshold, linetype = "dashed", color = "grey25") +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  ylab(expression(log(Demand)~(g~N~m^-2~d^-1))) +
  theme_classic() +
      scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
  theme(legend.position = "none")  +
    facet_grid(catch~.)

N_demand_plot1

```

### *Figure 4*


```{r, warning=F, size = 'small', echo=F,include=T, fig.width=12, fig.height=6}

n_s_grid <- ggarrange(
  NO3_supp_plots,
  N_demand_plot1,
  labels = c("b", "c"),
    align = c("v"),
  ncol = 1, nrow = 2,
  label.x = 0.91,  
  label.y = 1,     
  hjust = 1,      
  vjust = 1       
)

N_ratio_plots <- N_ratio_plots + 
  theme(
    aspect.ratio = 1, 
    plot.margin = margin(10, 20, 10, 20),
    legend.position = "bottom"
  ) +
  guides(
    shape = guide_legend(nrow = 2), 
    color = guide_legend(nrow = 2)   
  )

Ndyn_grid <- ggarrange(
  N_ratio_plots,
  n_s_grid,
  ncol = 2, nrow = 1,
  labels = c("a", ""),
  vjust = 7,
  label.x = 0.23,  
  #common.legend = TRUE, 
  #legend = "bottom",
  widths = c(1.3, 1.6)
)

Ndyn_grid


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=9, fig.height=4}
# ggsave("./Fig4_n_dyn.png", plot = Ndyn_grid, width = 10.75, height = 6.75, units = "in")

```


## Dry to wet year comparisons 
###  ANOVA's for differences in at each reach


##### GPP:

```{r, warning=F, size = 'small', echo=T, include=T}
# Define the water years you want on the x-axis, including 2021
desired_years <- c("2021", "2022", "2023")

anova_n03_sup_bw <- aov(NO3_supply_new ~ as.factor(water_year), data = covariat_NS_ND%>%filter(site == "BWL", water_year %in% 2021:2023))
summary(anova_n03_sup_bw)

posthoc_n03 <- TukeyHSD(anova_n03_sup_bw)
posthoc_n03

#  there are no statistical diffs


anova_n03_sup_bw <- aov(NO3_supply_new ~ as.factor(water_year), data = covariat_NS_ND%>%filter(site == "BWU", water_year %in% 2021:2023))
summary(anova_n03_sup_bw)

posthoc_n03 <- TukeyHSD(anova_n03_sup_bw)
posthoc_n03

#  there are no statistically diffs

anova_n03_sup_gb <- aov(NO3_supply_new ~ as.factor(water_year), data = covariat_NS_ND%>%filter(site == "GBL", water_year %in% 2021:2023))
summary(anova_n03_sup_gb)

posthoc_n03 <- TukeyHSD(anova_n03_sup_gb)
posthoc_n03

#  there are no statistically diffs
```


##### NO3 supply 


```{r, warning=F, size = 'small', echo=F, fig.width=5, include=T,fig.height=4}

desired_years <- c("2021","2022","2023")

NO3_mean_diff_wet_dry <- covariat_NS_ND %>%
  filter(water_year %in% 2021:2023) %>%
  mutate(
    water_year = as.integer(water_year),
    year_group = case_when(
      water_year %in% c(2021, 2022) ~ "2021_2022",
      water_year == 2023            ~ "2023",
      TRUE                          ~ NA_character_
    )
  ) %>%
  filter(!is.na(year_group)) %>%
  group_by(site, year_group) %>%
  summarise(
    mean_NO3 = mean(NO3_supply_new, na.rm = TRUE),
    n = sum(!is.na(NO3_supply_new)),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = year_group,
    values_from = c(mean_NO3, n)
  ) %>%
  mutate(
    # mean difference relative to 2023 (same sign convention as your statement: "rose" => positive)
    delta_wet_dry = mean_NO3_2023- mean_NO3_2021_2022,
    # percent difference relative to 2023
    pct_change = round(100 * (mean_NO3_2023- mean_NO3_2021_2022) / mean_NO3_2021_2022)
  ) %>%
  arrange(site)

NO3_mean_diff_wet_dry
```


```{r, warning=F, size = 'small', echo=F, fig.width=5, include=FALSE,fig.height=4}

library(dplyr)
library(ggplot2)
library(multcompView)

desired_years <- c("2021","2022","2023")



make_site_plot_yearonly <- function(site_id, y_max = 2.0, label_pad = 0.05, alpha_letters = 0.10) {

  df <- covariat_NS_ND %>%
    filter(site == site_id, water_year %in% 2021:2023) %>%
    mutate(water_year = factor(as.character(water_year), levels = desired_years)) %>%
    filter(is.finite(NO3_supply_new))  # drops NA/NaN/Inf

  # guard: if no data, return an empty plot instead of erroring
  if (nrow(df) == 0) {
    p <- ggplot() + theme_classic() + labs(title = paste("No data for", site_id))
    return(list(plot = p, anova = NULL, tukey = NULL))
  }

  fit <- aov(NO3_supply_new ~ water_year, data = df)
  tuk <- TukeyHSD(fit)

  letters_df <- make_letters_from_tukey(
    tukey = tuk,
    term  = "water_year",
    alpha = alpha_letters
  ) %>%
    mutate(water_year = factor(water_year, levels = desired_years))

  y_lab <- min(y_max, max(df$NO3_supply_new, na.rm = TRUE) + label_pad)

  p <- ggplot(df, aes(x = water_year, y = NO3_supply_new)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA,
                 aes(fill = catch, color = catch)) +
    scale_color_manual(values = catch_colors) +
    scale_fill_manual(values = catch_colors) +
    labs(x = NULL, y = NULL, fill = "Catchment", color = "Catchment") +
    coord_cartesian(ylim = c(0, y_max)) +
    theme_classic() +
    geom_text(
      data = letters_df,
      aes(x = water_year, y = y_lab, label = letters),
      inherit.aes = FALSE,
      size = 3
    )

  list(plot = p, anova = summary(fit), tukey = tuk)
}


library(multcompView)
library(dplyr)

make_letters_from_tukey <- function(tukey, term, alpha = 0.10) {

  pvals <- tukey[[term]][, "p adj"]
  names(pvals) <- rownames(tukey[[term]])

  # TRUE = significantly different
  sig <- pvals < alpha

  cld <- multcompLetters(sig)

  data.frame(
    water_year = names(cld$Letters),
    letters = cld$Letters,
    stringsAsFactors = FALSE
  )
}




# Build all site results (letters computed consistently at alpha=0.10)
res_BWL <- make_site_plot_yearonly("BWL", alpha_letters = 0.10)
plot_Nsupply_BWL <- res_BWL$plot

res_BWU <- make_site_plot_yearonly("BWU", alpha_letters = 0.10)
plot_Nsupply_BWU <- res_BWU$plot

res_GBL <- make_site_plot_yearonly("GBL", alpha_letters = 0.10)
plot_Nsupply_GBL <- res_GBL$plot

res_GBU <- make_site_plot_yearonly("GBU", alpha_letters = 0.10)
plot_Nsupply_GBU <- res_GBU$plot

```

##### NH4 supply 


```{r, warning=F, size = 'small', echo=T, fig.width=5, include=T,fig.height=4}

desired_years <- c("2021","2022","2023")

nh4_mean_diff_wet_dry <- covariat_NS_ND %>%
  filter(water_year %in% 2021:2023) %>%
  mutate(
    water_year = as.integer(water_year),
    year_group = case_when(
      water_year %in% c(2021, 2022) ~ "2021_2022",
      water_year == 2023            ~ "2023",
      TRUE                          ~ NA_character_
    )
  ) %>%
  filter(!is.na(year_group)) %>%
  group_by(site, year_group) %>%
  summarise(
    mean_NH4 = mean(NH4_supply_new, na.rm = TRUE),
    n = sum(!is.na(NH4_supply_new)),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = year_group,
    values_from = c(mean_NH4, n)
  ) %>%
  mutate(
    # mean difference relative to 2023 (same sign convention as your statement: "rose" => positive)
    delta_wet_dry = mean_NH4_2023- mean_NH4_2021_2022,
    # percent difference relative to 2023
    pct_change = round(100 * (mean_NH4_2023- mean_NH4_2021_2022) / mean_NH4_2021_2022)
  ) %>%
  arrange(site)

nh4_mean_diff_wet_dry
```


```{r, warning=F, size = 'small', echo=F, fig.width=5, include=FALSE,fig.height=4}


library(dplyr)
library(ggplot2)
library(multcompView)

desired_years <- c("2021","2022","2023")

make_letters_from_tukey <- function(tukey, term, alpha = 0.10) {
  pvals <- tukey[[term]][, "p adj"]
  names(pvals) <- rownames(tukey[[term]])

  sig <- pvals < alpha
  cld <- multcompLetters(sig)

  data.frame(
    water_year = names(cld$Letters),
    letters = cld$Letters,
    stringsAsFactors = FALSE
  )
}

make_site_plot_yearonly <- function(site_id, y_max = 0.75, label_pad = 0.05, alpha_letters = 0.10) {

  df <- covariat_NS_ND %>%
    filter(site == site_id, water_year %in% 2021:2023) %>%
    mutate(water_year = factor(as.character(water_year), levels = desired_years))

  # Safety check (helps avoid confusing Tukey errors if a year is missing)
  if (nrow(df) == 0) stop(paste("No data for site:", site_id))
  if (nlevels(droplevels(df$water_year)) < 2) stop(paste("Need >= 2 water years for site:", site_id))

  # ANOVA (year only) -- within this site only
  fit <- aov(NH4_supply_new ~ water_year, data = df)

  # Tukey for year factor
  tuk <- TukeyHSD(fit, which = "water_year")

  letters_df <- make_letters_from_tukey(tuk, term = "water_year", alpha = alpha_letters) %>%
    mutate(water_year = factor(water_year, levels = desired_years))

  y_lab <- min(y_max, max(df$NH4_supply_new, na.rm = TRUE) + label_pad)

  p <- ggplot(df, aes(x = water_year, y = NH4_supply_new)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA,
                 aes(fill = catch, color = catch)) +
    scale_color_manual(values = catch_colors) +
    scale_fill_manual(values = catch_colors) +
    labs(x = NULL, y = NULL, fill = "Catchment", color = "Catchment") +
    coord_cartesian(ylim = c(0, y_max)) +
    ylab(expression(NH[4]~"supply"~(g~N~m^-2~d^-1))) +
    theme_classic() +
    geom_text(
      data = letters_df,
      aes(x = water_year, y = y_lab, label = letters),
      inherit.aes = FALSE,
      size = 3
    )

  list(plot = p, anova = summary(fit), tukey = tuk, letters = letters_df, data = df)
}


# Build all site results (letters computed consistently at alpha=0.10)
res_BWL <- make_site_plot_yearonly("BWL", alpha_letters = 0.10)
plot_NH4supply_BWL <- res_BWL$plot

res_BWU <- make_site_plot_yearonly("BWU", alpha_letters = 0.05)
plot_NH4supply_BWU <- res_BWU$plot

res_GBL <- make_site_plot_yearonly("GBL", alpha_letters = 0.10)
plot_NH4supply_GBL <- res_GBL$plot

res_GBU <- make_site_plot_yearonly("GBU", alpha_letters = 0.10)
plot_NH4supply_GBU <- res_GBU$plot

```



#### N demand

```{r, warning=F, size = 'small', echo=F, fig.width=5, include=T,fig.height=4}

desired_years <- c("2021","2022","2023")

Ndemand_mean_diff_wet_dry <- covariat_NS_ND %>%
  filter(water_year %in% 2021:2023) %>%
  mutate(
    water_year = as.integer(water_year),
    year_group = case_when(
      water_year %in% c(2021, 2022) ~ "2021_2022",
      water_year == 2023            ~ "2023",
      TRUE                          ~ NA_character_
    )
  ) %>%
  filter(!is.na(year_group)) %>%
  group_by(site, year_group) %>%
  summarise(
    mean_Ndemand = mean(Ndemand, na.rm = TRUE),
    n = sum(!is.na(Ndemand)),
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(
    names_from = year_group,
    values_from = c(mean_Ndemand, n)
  ) %>%
  mutate(
    # mean difference relative to 2023 (same sign convention as your statement: "rose" => positive)
    delta_wet_dry = mean_Ndemand_2023- mean_Ndemand_2021_2022,
    # percent difference relative to 2023
    pct_change = round(100 * (mean_Ndemand_2023- mean_Ndemand_2021_2022) / mean_Ndemand_2021_2022)
  ) %>%
  arrange(site)

Ndemand_mean_diff_wet_dry

```

```{r, warning=F, size = 'small', echo=F, fig.width=5, include=FALSE,fig.height=4}

library(dplyr)
library(ggplot2)
library(multcompView)


desired_years <- c("2021","2022","2023")

make_letters_from_tukey <- function(tukey, term, alpha = 0.10) {
  pvals <- tukey[[term]][, "p adj"]
  names(pvals) <- rownames(tukey[[term]])

  sig <- pvals < alpha
  cld <- multcompLetters(sig)

  data.frame(
    water_year = names(cld$Letters),
    letters = cld$Letters,
    stringsAsFactors = FALSE
  )
}

make_site_plot_yearonly <- function(site_id, y_max = 0.75, label_pad = 0.05, alpha_letters = 0.10) {

  df <- covariat_NS_ND %>%
    filter(site == site_id, water_year %in% 2021:2023) %>%
    mutate(water_year = factor(as.character(water_year), levels = desired_years))

  # Safety check (helps avoid confusing Tukey errors if a year is missing)
  if (nrow(df) == 0) stop(paste("No data for site:", site_id))
  if (nlevels(droplevels(df$water_year)) < 2) stop(paste("Need >= 2 water years for site:", site_id))

  # ANOVA (year only) -- within this site only
  fit <- aov(Ndemand ~ water_year, data = df)

  # Tukey for year factor
  tuk <- TukeyHSD(fit, which = "water_year")

  letters_df <- make_letters_from_tukey(tuk, term = "water_year", alpha = alpha_letters) %>%
    mutate(water_year = factor(water_year, levels = desired_years))

  y_lab <- min(y_max, max(df$Ndemand, na.rm = TRUE) + label_pad)

  p <- ggplot(df, aes(x = water_year, y = Ndemand)) +
    geom_boxplot(alpha = 0.5, outlier.shape = NA,
                 aes(fill = catch, color = catch)) +
    scale_color_manual(values = catch_colors) +
    scale_fill_manual(values = catch_colors) +
    labs(x = NULL, y = NULL, fill = "Catchment", color = "Catchment") +
    coord_cartesian(ylim = c(0, y_max)) +
    ylab(expression(NH[4]~"supply"~(g~N~m^-2~d^-1))) +
    theme_classic() +
    geom_text(
      data = letters_df,
      aes(x = water_year, y = y_lab, label = letters),
      inherit.aes = FALSE,
      size = 3
    )

  list(plot = p, anova = summary(fit), tukey = tuk, letters = letters_df, data = df)
}



# Build all site results (letters computed consistently at alpha=0.10)
res_BWL <- make_site_plot_yearonly("BWL", alpha_letters = 0.10)
plot_Ndemand_BWL <- res_BWL$plot

res_BWU <- make_site_plot_yearonly("BWU", alpha_letters = 0.10)
plot_Ndemand_BWU <- res_BWU$plot

res_GBL <- make_site_plot_yearonly("GBL", alpha_letters = 0.10)
plot_Ndemand_GBL <- res_GBL$plot

res_GBU <- make_site_plot_yearonly("GBU", alpha_letters = 0.10)
plot_Ndemand_GBU <- res_GBU$plot

```

### *Figure 5*

```{r, warning=F, size = 'small', echo=F, fig.width=8, include=T,fig.height=6}

Ndyn_annual_grid <- ggarrange(
  plot_Nsupply_BWL,
  plot_Nsupply_BWU,
  plot_Nsupply_GBL,
  plot_Nsupply_GBU,
  plot_NH4supply_BWL,
  plot_NH4supply_BWU,
  plot_NH4supply_GBL,
  plot_NH4supply_GBU,
  plot_Ndemand_BWL,
  plot_Ndemand_BWU,
  plot_Ndemand_GBL,
  plot_Ndemand_GBU,
  align = "v",
    ncol = 4, nrow = 3,
    common.legend = TRUE, 
   legend = "right")


 Ndyn_annual_grid 
  
  
```


```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("./Figure6_annualNtrend.png", plot = Ndyn_annual_grid, width = 8, height = 6.5, units = "in")

```

#### SI figures ###

### Comparable cumulative metabolism
Where days that the model can't identify GPP (in green) or ER (orange) but there is DO, we then assume the metabolism (GPP/ER) to be 0 


```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
  dplyr::filter(site=="BWL", !is.na(do.obs_m)) %>%
  dplyr::select(date, do.obs_m, GPP_mean) %>%
  dplyr::mutate(
    has_do = !is.na(do.obs_m),
    has_GPP = !is.na(GPP_mean),
    plot_GPP = dplyr::if_else(is.na(GPP_mean), 0, GPP_mean),
    yday = lubridate::yday(date),
    year = lubridate::year(date)
  )


# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(GPP_cumsum = cumsum(plot_GPP)) %>%
  dplyr::ungroup()



# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt_1 <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_GPP), aes(y = 0.090), alpha= 0.8, color = "#579467", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Lower - GPP") +facet_grid(year~.)
templt_1
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_seq <- seq(from = as.Date("2021-03-20"), to = as.Date("2024-10-10"), by = "day")

# Convert the sequence to a dataframe
date_df <- data.frame(date = date_seq)
date_df$site <- "BWL"
date_df$catch <- "BW"
date_df1 <- data.frame(date = date_seq)
date_df1$site <- "GBL"
date_df1$catch <- "GB"
date_df2 <- data.frame(date = date_seq)
date_df2$site <- "GBU"
date_df2$catch <- "GB"
date_df3 <- data.frame(date = date_seq)
date_df3$site <- "BWU"
date_df3$catch <- "BW"

date_datf <- rbind(date_df,date_df1,date_df2, date_df3)


date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_bw <- date_datf_plt %>%
 dplyr::filter(site=="BWL" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

BWL_plot <- ggplot(date_datf_plt_bw, aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  labs(x = "Day of Water Year", color = "Year", 
       title = "BW Lower") +
  ylab(expression(GPP~(g~O[2]~m^-2~yr^-1))) +
  xlab(NULL) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() +
   xlim(91, 335)
#  xlim(115, 290)

BWL_plot

# Compute the maximum cumulative GPP for each year
date_datf_plt_sum <- date_datf_plt_bw %>%
  group_by(year) %>%
  dplyr::summarise(T_gpp = max(GPP_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_gpp -T_gpp[year == 2023]) / T_gpp * 100)


dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_gpp)) %>%
  pull(dry_mean)


wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_gpp)


tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)


```

## Wet to dry 102.0758


```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="BWL") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, ER_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_ER = !is.na(ER_mean),
         ER_abs = (ER_mean*-1),
         plot_ER = ifelse(is.na(ER_abs), 0, ER_abs),
         yday=yday(date),
         year=year(date))



# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(ER_cumsum = cumsum(plot_ER)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_ER), aes(y = 0.090), alpha= 0.8, color = "chocolate", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Lower - ER") +facet_grid(year~.)
templt
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_bw <- date_datf_plt %>%
 dplyr::filter(site=="BWL" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

BWL_plot1 <- ggplot(date_datf_plt_bw, aes(x = yday, y = ER_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
      ylim(0,1500)+
  labs(x = "Day of Water Year", color = "Year") +
  ylab(expression("|"~ER~"|"~(g~O[2]~m^-2~yr^-1))) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() +
    xlab(NULL) +
     xlim(91, 335)
#  xlim(115, 290)


# Compute the maximum cumulative GPP for each year
date_datf_plt_sum <- date_datf_plt_bw %>%
  group_by(year) %>%
  dplyr::summarise(T_cm = max(ER_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)

dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)


wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)


tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)

```




```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="BWU") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_GPP = !is.na(GPP_mean),
         plot_GPP = ifelse(is.na(GPP_mean), 0, GPP_mean),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(GPP_cumsum = cumsum(plot_GPP)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_GPP), aes(y = 0.090), alpha= 0.8, color = "#579467", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Upper - GPP") +facet_grid(year~.)
templt
```


```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_bw <- date_datf_plt %>%
 dplyr::filter(site=="BWU" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

BWU_plot <- ggplot(date_datf_plt_bw, aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  labs(x = "Day of Water Year", color = "Year", 
       title = "BW Upper") +
  ylab(expression(GPP~(g~O[2]~m^-2~yr^-1))) +
  ylim(0,400) +
  scale_color_manual(values=c("#440154", 
                              "#3B528B", 
                              "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() +
    xlab(NULL) +
       xlim(91, 335)



# Compute the maximum cumulative GPP for each year
date_datf_plt_sum <- date_datf_plt_bw %>%
  group_by(year) %>%
  dplyr::summarise(T_cm = max(GPP_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)

dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)


wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)


tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)


```




```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="BWU") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, ER_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_ER = !is.na(ER_mean),
         ER_abs = (ER_mean*-1),
         plot_ER = ifelse(is.na(ER_abs), 0, ER_abs),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(ER_cumsum = cumsum(plot_ER)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_ER), aes(y = 0.090), alpha= 0.8, color = "chocolate", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Upper - ER") +facet_grid(year~.)
templt
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_bw <- date_datf_plt %>%
 dplyr::filter(site=="BWU" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

BWU_plot1 <- ggplot(date_datf_plt_bw, aes(x = yday, y = ER_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  labs(x = "Day of Water Year", color = "Year") +
  ylab(expression("|"~ER~"|"~(g~O[2]~m^-2~yr^-1))) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() +
        ylim(0,1500)+
    xlab(NULL) +
       xlim(91, 335)

  #xlim(185, 250)


# Compute the maximum cumulative GPP for each year
date_datf_plt_sum <- date_datf_plt_bw %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(ER_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)


dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)


wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)


tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)


```


```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="GBL") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_GPP = !is.na(GPP_mean),
         plot_GPP = ifelse(is.na(GPP_mean), 0, GPP_mean),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, 
                                      yday_2022, 
                                      yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 
                                             2022, 
                                             2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(GPP_cumsum = cumsum(plot_GPP)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_GPP), aes(y = 0.090), alpha= 0.8, color = "#579467", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Lower - GPP") +facet_grid(year~.)
templt
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_gb <- date_datf_plt %>%
 dplyr::filter(site=="GBL" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

GBL_plot <- ggplot(date_datf_plt_gb, aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) +
  ylim(0,30) +
  labs(x = "Day of Water Year", color = "Year", 
       title = "GB Lower") +
  ylab(expression(GPP~(g~O[2]~m^-2~yr^-1))) +
    xlab(NULL) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() + xlim(91, 335)


# Compute the maximum cumulative GPP for each year
date_datf_plt_sum <- date_datf_plt_gb %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(GPP_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)


date_datf_plt_sum <- date_datf_plt_gb %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(GPP_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm[year == 2023]-T_cm) / T_cm[year == 2023] * 100)

dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)

wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)

tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)

```


```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="GBL") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, ER_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_ER = !is.na(ER_mean),
         ER_abs = (ER_mean*-1),
         plot_ER = ifelse(is.na(ER_abs), 0, ER_abs),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))


filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(ER_cumsum = cumsum(plot_ER)) %>%
  dplyr::ungroup()


# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_ER), aes(y = 0.090), alpha= 0.8, color = "chocolate", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Lower - ER") +facet_grid(year~.)
templt
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_gb <- date_datf_plt %>%
 dplyr::filter(site=="GBL" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

GBL_plot1 <- ggplot(date_datf_plt_gb, aes(x = yday, y = ER_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  labs(x = "Day of Water Year", color = "Year") +  xlab(NULL) +
  ylab(expression("|"~ER~"|"~(g~O[2]~m^-2~yr^-1))) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() + xlim(91, 335)
#xlim(115, 332)

GBL_plot1


date_datf_plt_sum <- date_datf_plt_gb %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(ER_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm[year == 2023]-T_cm) / T_cm[year == 2023] * 100)


dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)

wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)

tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)
```




```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="GBU") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_GPP = !is.na(GPP_mean),
         plot_GPP = ifelse(is.na(GPP_mean), 0, GPP_mean),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, 
                                      yday_2022, 
                                      yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 
                                             2022, 
                                             2023))

filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(GPP_cumsum = cumsum(plot_GPP)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_GPP), aes(y = 0.090), alpha= 0.8, color = "#579467", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Upper - GPP") +facet_grid(year~.)
templt
```

```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_gb <- date_datf_plt %>%
 dplyr::filter(site=="GBU" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

GBU_plot <- ggplot(date_datf_plt_gb, aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  labs(x = "Day of Water Year", color = "Year", 
       title = "GB Upper") +
  ylab(expression(GPP~(g~O[2]~m^-2~yr^-1))) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() + xlim(91, 335)
# xlim(115, 340)


date_datf_plt_sum <- date_datf_plt_gb %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(GPP_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)

dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)

wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)

tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (wet_2023 - dry_mean) / dry_mean * 100
)

```



```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=2}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
 dplyr::filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
   dplyr::filter(site=="GBU") %>%
 dplyr::filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, ER_mean) %>%
  mutate(has_do = !is.na(do.obs_m), 
         has_ER = !is.na(ER_mean),
         ER_abs = (ER_mean*-1),
         plot_ER = ifelse(is.na(ER_abs), 0, ER_abs),
         yday=yday(date),
         year=year(date))

# Identify the unique yday values for each year
yday_2021 <- unique(plot_data_DO$yday[plot_data_DO$year == 2021])
yday_2022 <- unique(plot_data_DO$yday[plot_data_DO$year == 2022])
yday_2023 <- unique(plot_data_DO$yday[plot_data_DO$year == 2023])

# Find the common yday values across all three years
common_yday <- Reduce(intersect, list(yday_2021, yday_2022, yday_2023))

# Filter the dataframe for only those common yday values
filtered_data <- plot_data_DO %>%
 dplyr::filter(yday %in% common_yday & year %in% c(2021, 2022, 2023))


filtered_data_cs <- filtered_data %>%
  dplyr::group_by(year) %>%
  dplyr::arrange(yday, .by_group = TRUE) %>%
  dplyr::mutate(ER_cumsum = cumsum(plot_ER)) %>%
  dplyr::ungroup()

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(filtered_data, aes(x = yday)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data =dplyr::filter(filtered_data, has_do), aes(y = 0.095), alpha=0.8, color = "black", size = 1) + # DO presence tick marks
  geom_point(data =dplyr::filter(filtered_data, has_ER), aes(y = 0.090), alpha= 0.8, color = "chocolate", size = 1) + # GPP presence tick marks
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Upper - ER") +facet_grid(year~.)
templt
```


```{r, warning=F, size = 'small', echo=F, include=F}
#####
date_datf_plt <- date_datf %>%
  mutate(yday=yday(date),
         year=year(date))

date_datf_plt_gb <- date_datf_plt %>%
 dplyr::filter(site=="GBU" & year <2024) %>%
  left_join(filtered_data_cs, by=c("yday", "year", "date"))

GBU_plot1 <- ggplot(date_datf_plt_gb, aes(x = yday, y = ER_cumsum, color = factor(year))) +
  geom_line(size = 1, alpha=0.95) + 
  ylim(0,600) +
  labs(x = "Day of Water Year", color = "Year") +
  ylab(expression("|"~ER~"|"~(g~O[2]~m^-2~yr^-1))) +
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
 #   scale_color_viridis_d(option = "viridis",  name = "Year") +
  theme_classic() + xlim(91, 335)
#xlim(115, 332)


date_datf_plt_sum <- date_datf_plt_gb %>%
  group_by(year) %>%
 dplyr::summarise(T_cm = max(ER_cumsum, na.rm = TRUE)) %>%
  mutate(percent_change = (T_cm -T_cm[year == 2023]) / T_cm * 100)


dry_mean <- date_datf_plt_sum %>%
  dplyr::filter(year %in% c(2021, 2022)) %>%
  dplyr::summarise(dry_mean = mean(T_cm)) %>%
  pull(dry_mean)

wet_2023 <- date_datf_plt_sum %>%
  dplyr::filter(year == 2023) %>%
  pull(T_cm)

tibble(
  dry_mean = dry_mean,
  wet_2023 = wet_2023,
  diff = wet_2023 - dry_mean,
  pct_decrease = (dry_mean-wet_2023) / dry_mean * 100
)

```



#### Cummulative GPP and ER 

From overlapping DOY observations in each year. 

```{r, warning=F, size = 'small', echo=F, include=T,fig.width=8, fig.height=8.5}

Csum_grid <- ggarrange(
  BWL_plot,
  BWU_plot,
  GBL_plot,
  GBU_plot,
  ncol = 1, nrow = 4,
  common.legend = TRUE, 
  align=c("v"),
  labels = c("a", "c", "e", "g"),
  label.x = 0.8,  # Move label to the right
  label.y = 1, 
  legend = "bottom")

Csum_grid_er <- ggarrange(
  BWL_plot1,
  BWU_plot1,
  GBL_plot1,
  GBU_plot1,
  ncol = 1, nrow = 4,
  common.legend = TRUE, 
  align=c("v"),
  labels = c("b", "d", "f", "h"),
  label.x = 0.8,  # Move label to the right
  label.y = 1, 
  legend = "bottom")

Csum_grid_T <- ggarrange(
  Csum_grid,
  Csum_grid_er,
  ncol = 2, nrow = 1,
  common.legend = TRUE, 
  legend = "bottom")


Csum_grid_T

```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("./Fig_cumsum_GPP_v2.png", plot = Csum_grid_T, width = 5.5, height = 8.25, units = "in")

```

End of script.
