---
title: 'Figures for MS: Spatiotemporal variation in mountain stream metabolism and nitrogen cycling across contrasting flow regimes '
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
#setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWL_k600/")
lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
```

## Research Questions:

(1) How does stream metabolism vary based on catchment characteristics, antecedent flow conditions, and nitrogen availability?

    -   Figures 2, 3

(2) Is variance in nitrogen demand and instream uptake (either as ammonium or nitrate) more related to biological (GPP, ER, biomass or chlorophyll-a), physical processes (flow and water temperature), or ambient nutrient and organic matter availability? 

    -   Figures 4, 5
    
(3) How does in-stream productivity and nitrogen supply and demand dynamics change within stream catchments and across dry (2021 and 2022) and wet water years (2023)?

    -   Figures 6

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
daoc_date<- Sys.Date()

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
site_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#97b5c2",
  "GBL" = "#DD6E42",
  "GBU" = "#d9cb7c"
)

siteC_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#054fb9",
  "GBL" = "#DD6E42",
  "GBU" = "#DD6E42"
)



catch_colors <- c(
  "BW" = "#054fb9",
  "GB" = "#DD6E42"
)

# Define fill patterns for water years
water_year_patterns <- c("wave","weave","stripe","circle", "stripe", "circle")


# Create a sequence of dates
date_seq <- seq(from = as.Date("2021-03-20"), to = as.Date("2024-10-10"), by = "day")

# Convert the sequence to a dataframe
date_df <- data.frame(date = date_seq)
date_df$site <- "BWL"
date_df$catch <- "BW"
date_df1 <- data.frame(date = date_seq)
date_df1$site <- "GBL"
date_df1$catch <- "GB"
date_df2 <- data.frame(date = date_seq)
date_df2$site <- "GBU"
date_df2$catch <- "GB"
date_df3 <- data.frame(date = date_seq)
date_df3$site <- "BWU"
date_df3$catch <- "BW"

date_datf <- rbind(date_df,date_df1,date_df2, date_df3)

### fxns:
## Fxn for water year:
water_year <- function(data) {
  # Ensure the `date` column exists
  if (!"date" %in% names(data)) {
    stop("The dataframe must contain a column named 'date'.")
  }
  
  # Convert the `date` column to Date format (if not already)
  data$date <- as.Date(data$date)
  
  # Add the water_year column
  data$water_year <- with(data, {
    year <- as.numeric(format(date, "%Y"))
    month <- as.numeric(format(date, "%m"))
    ifelse(month >= 10, year + 1, year)
  })
  
  return(data)
}
## Function to perform ANOVA and post-hoc test within each site
comparisons <- list(
  c("2021", "2022"),
  c("2021", "2023"),
  c("2022", "2023"),
  c("2023", "2024")
)
```

### Figure 1: Map

Not included here.

```{r, warning=F, size = 'small', echo=FALSE, include=F}
### Read in all data
####
date_datf <- water_year(date_datf)

covariat_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_covariate_dat.rds") %>%
  dplyr::select("date", "Site","bulk.density","AFDM_mgg","AFDM_mgcm2", "Chla_ugL_Q", "Pheo_ugL_Q")
summary(covariat_dat)

covariat_dat <- date_datf %>%
  left_join(covariat_dat, by=c("date", "site"="Site"))

bg_nuts <- readRDS("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/NS_chem_dat_nh4_24.rds") %>%
  #mutate(date = as.Date(date, format="%m/%d/%y")) %>%
  dplyr::filter(location=="stream") %>%
    mutate(NO3_mgL_dl = if_else(NO3_mgL_dl < 0.0015, 0.0015, NO3_mgL_dl))%>%
    mutate(NH4_mgL_dl=if_else(NH4_mgL_dl < 0.001, 0.001, NH4_mgL_dl))%>%
    mutate(PO4_ugL_dl=if_else(PO4_ugL_dl < 0.201, 0.201, PO4_ugL_dl))

WQ_dat <- read.csv("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/Stream_Lake_YSI_WaterQuality.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) %>%
  dplyr::select("site", "date","pH") %>%
  dplyr:: group_by(site, date) %>%
  dplyr::summarise(pH=mean(pH, na.rm=T))

covariat_datq <- covariat_dat%>%
  left_join(bg_nuts[ , c("site", "substrate", "date", "NO3_mgL_dl", "NH4_mgL_dl", "PO4_ugL_dl", "DOC_mgL_dl")], by=c("date", "site"))

covariat_datq <- covariat_datq%>%
  left_join(WQ_dat,  by=c("date", "site"))


## bring in SPC data 
### missing 2024 BW SPC
SPC_BWL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWL_SPCv2.rds") %>%
  dplyr:: filter(datetime>as.POSIXct("2021-04-29 24:00:00"))
SPC_BWL$site <- "BWL"
SPC_BWU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWU_SPCv2.rds")
SPC_BWU$site <- "BWU"
SPC_GBL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBL_SPCv2.rds")
SPC_GBL$site <- "GBL"
SPC_GBU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBU_SPCv2.rds")
SPC_GBU$site <- "GBU"

SPC_dat <- rbind(SPC_BWL, SPC_BWU, SPC_GBL, SPC_GBU)
SPC_dat <- SPC_dat %>%
  mutate(date =as.Date(datetime)) 

SPC_dat_day <- SPC_dat %>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    wt= mean(wtr, na.rm=T),
    wt_sd= sd(wtr, na.rm=T),
    SPC_m=mean(SPC, na.rm=T),
    SPC_sd=sd(SPC, na.rm=T))

#hist(SPC_dat_day$wt_sd)
#hist(SPC_dat_day$SPC_sd)

SPC_datD <- SPC_dat_day %>%
   dplyr::filter(wt_sd<4.1 & SPC_sd <35)
  
#hist(SPC_datD$wt_sd)
#hist(SPC_datD$SPC_sd)
  
covariat_datq <- covariat_datq%>%
  left_join(SPC_datD, by=c("date", "site"))

### 
## Bring in morph data : 
## BWL  
BWL_datM <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_BWL_DO_flag_sat_light_v2.rds") %>%
  dplyr::select(datetime, w, v, depth) %>%
  mutate(site="BWL")
summary(BWL_datM) 

## BWU 
BWU_datM <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_BWU_DO_flag_sat_light_v2.rds") %>%
    dplyr::select(datetime, w, v, depth) %>%
  mutate(site="BWU")
summary(BWU_datM)

## GBL 
GBL_datM <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBL_DO_flag_sat_light_v2.rds") %>%
    dplyr::select(datetime, w, v, depth) %>%
  mutate(site="GBL")
summary(GBL_datM)

## GBU 
GBU_datM <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBU_DO_flag_sat_light_v2.rds") %>%
    dplyr::select(datetime, w, v, depth) %>%
  mutate(site="GBU")
summary(GBU_datM)

Morph_dat <- rbind(BWL_datM, BWU_datM, GBL_datM, GBU_datM)

Morph_dat_day <- Morph_dat%>%
  mutate(date =as.Date(datetime))%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    v_m=mean(v, na.rm=T),
    v_sd=sd(v, na.rm=T),
    w_m=mean(w, na.rm=T),
    w_sd=sd(w, na.rm=T),
    depth_m=mean(depth, na.rm=T),
    depth_sd=sd(depth, na.rm=T))


covariat_datq <- covariat_datq%>%
  left_join(Morph_dat_day, by=c("date", "site"))


#### DO ###
## BWL  
BWL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_BWL_mod_dat.rds") %>%
  mutate(site="BWL")
summary(BWL_dat) 

## BWU 
BWU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWU_k600/25_BWU_mod_dat.rds") %>%
  mutate(site="BWU")
summary(BWU_dat)

## GBL 
GBL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBL_mod_dat.rds") %>%
  mutate(site="GBL")
summary(GBL_dat)

## GBU 
GBU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBU_mod_dat.rds") %>%
  mutate(site="GBU")
summary(GBU_dat)



miniDOT_dat <- rbind(BWL_dat, BWU_dat, GBL_dat, GBU_dat)
miniDOT_dat_day <- miniDOT_dat%>%
  mutate(date =as.Date(solar.time))%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    do.obs_m=mean(DO.obs, na.rm=T),
    do.obs_sd=sd(DO.obs, na.rm=T),
    wtr_m=mean(temp.water, na.rm=T),
    wtr_sd=sd(temp.water, na.rm=T),
    Q_m= mean(discharge, na.rm=T),
    Q_sd= sd(discharge, na.rm=T))

covariat_datq <- covariat_datq%>%
  left_join(miniDOT_dat_day, by=c("date", "site"))

### N-uptake metrics
n_up <- read.csv("/Users/kellyloria/Documents/UNR/Ncycle/BTC_N_Table_2025_figure.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) 

covariat_datq <- covariat_datq%>%
  left_join(n_up, by=c("date", "site"))

##================================
## Load metabolism model data
##================================
BWL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWL_Full_2025-02-06_daily.csv") %>%
  dplyr::filter(GPP_daily_Rhat<1.05)%>%
  dplyr::filter(GPP_97.5pct>0)%>%
  dplyr::filter(ER_daily_Rhat<1.05) %>%
  dplyr::filter(ER_2.5pct<0)%>%
  dplyr::filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  dplyr::filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>% 
  mutate(site="BWL",
         catch="BW")

summary(BWL_met_binned)

GBL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBL_Full_2025-02-02_daily.csv") %>%
  dplyr::filter(GPP_daily_Rhat<1.05)%>%
  dplyr::filter(GPP_97.5pct>0)%>%
  dplyr::filter(ER_daily_Rhat<1.05) %>%
  dplyr::filter(ER_2.5pct<0)%>%
  dplyr::filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  dplyr::filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBL",
         catch="GB")


BWU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWU_Full_2025-02-04_daily.csv") %>%
  dplyr::filter(GPP_daily_Rhat<1.05)%>%
  dplyr::filter(GPP_97.5pct>0)%>%
  dplyr::filter(ER_daily_Rhat<1.05) %>%
  dplyr::filter(ER_2.5pct<0)%>%
  dplyr::filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  dplyr::filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="BWU",
         catch="BW")

GBU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBU_Full_2025-02-03_daily.csv") %>%
  dplyr::filter(GPP_daily_Rhat<1.05)%>%
  dplyr::filter(GPP_97.5pct>0)%>%
  dplyr::filter(ER_daily_Rhat<1.05) %>%
  dplyr::filter(ER_2.5pct<0)%>%
  dplyr::filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  dplyr::filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date,GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBU",
         catch="GB")

metab_dat <- rbind(BWL_met_binned, BWU_met_binned, GBL_met_binned, GBU_met_binned)

covariat_datq <- covariat_datq%>%
  left_join(metab_dat, by=c("date", "site", "catch"))
```

Pause to make a column for biomass to chla ratios And also make biomass
in in micro grams to more comparable to chl-a

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

covariat_datq$AFDM_ugcm2 <- (covariat_datq$AFDM_mgcm2 *1000)
#hist(covariat_datq$AFDM_ugcm2 )

covariat_datq$Biom_Chla <- (covariat_datq$AFDM_ugcm2)/(covariat_datq$Chla_ugL_Q)
#hist(covariat_datq$Biom_Chla)

summary(covariat_datq)

```

# Question 1

#### How does stream metabolism vary based on catchment characteristics, antecedent flow conditions, and nitrogen availability?

## Figure 2: Metabolism Time series

-   The magnitude of GPP and ER across all sites

Where GPP is in green, ER is in purple, NEP is in black dataframe:
`metab_dat`

```{r, warning=F, size = 'small', echo=F, fig.width=7, fig.height=8}
metab_TS_plot<- metab_dat %>% 
  ggplot(aes(x = date)) +
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6) +
  # GPP
  geom_errorbar(aes(ymin = GPP_2.5pct, ymax = GPP_97.5pct), width = 0.2, color = "#1f7335", alpha=0.2) +
  geom_point(aes(y = GPP_mean), col = "#1f7335", alpha = 0.8, size = 0.5) +
  # ER
  geom_errorbar(aes(ymin = ER_97.5pct, ymax = ER_2.5pct), width = 0.2, color = "#6c3a8a", alpha=0.2) +
  geom_point(aes(y = ER_mean), col = "#6c3a8a", alpha = 0.8, size = 0.5) +
  # NEP Points
  geom_point(aes(y = NEP_daily_mean), alpha = 0.25, size = 0.15, shape=1) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "12 week", date_labels = "%b-%y") +
  scale_fill_manual(values = c("GPP" = "#1f7335", "ER" = "#6c3a8a")) +
  ylab(expression(Metabolism~(g~O[2]~m^-2~d^-1))) + xlab(NULL) +
  facet_wrap(~site, nrow = 4) 

metab_TS_plot
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/metab_TS_plot_2025.png", plot = metab_TS_plot, width = 5.25, height = 6, units = "in")

```

Time series metabolism plots for each site

```{r, warning=F, size = 'small', echo=F,include=FALSE, fig.width=7, fig.height=8}
metab_TS_plot<- metab_dat %>% 
  dplyr::filter(site=="GBU")%>%
  ggplot(aes(x = date)) +
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6) +
  # GPP
  geom_errorbar(aes(ymin = GPP_2.5pct, ymax = GPP_97.5pct), width = 0.2, color = "#1f7335", alpha=0.2) +
  geom_point(aes(y = GPP_mean), col = "#1f7335", alpha = 0.8, size = 0.5) +
  # ER
  geom_errorbar(aes(ymin = ER_97.5pct, ymax = ER_2.5pct), width = 0.2, color = "#6c3a8a", alpha=0.2) +
  geom_point(aes(y = ER_mean), col = "#6c3a8a", alpha = 0.8, size = 0.5) +
  # NEP Points
  geom_point(aes(y = NEP_daily_mean), alpha = 0.25, size = 0.15, shape=1) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_breaks = "12 week", date_labels = "%b-%y") +
  scale_fill_manual(values = c("GPP" = "#1f7335", "ER" = "#6c3a8a")) +
  ylab(expression(Metabolism~(g~O[2]~m^-2~d^-1))) + xlab(NULL) +
  facet_wrap(~site, nrow = 4) 

metab_TS_plot
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Sup_Figures/ metab_TS_plot_GBU_2025.png", plot = metab_TS_plot, width = 9.25, height = 5, units = "in")

```

```{r, warning=F, size = 'small', echo=F, fig.width=10, fig.height=12}
# Separate out BW 
covariat_datq_BW <- covariat_datq%>%
  dplyr::filter(site=="BWL" | site=="BWU")

# Separate out GB 
covariat_datq_GB <- covariat_datq%>%
  dplyr::filter(site=="GBL" | site=="GBU")
```

#### Descriptive stats to describe stream based differences in

Both BW to GB - large to small streams, as well as variation with
stream.

-   Water temp

-   Flow

-   OM

-   Biomass

-   Chla

-   GPP

-   \|ER\|

### BW catchment:

```{r, warning=F, size = 'small', echo=F, include=T}

summary_tab <- covariat_datq_BW %>%
  summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

substrate_summary <- covariat_datq_BW %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
  group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl, na.rm = TRUE)
  )


# View results
print(summary_tab)
print(substrate_summary)
```

### GB Catchment

```{r, warning=F, size = 'small', echo=F, include=T}

summary_tab <- covariat_datq_GB %>%
  summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

substrate_summary <- covariat_datq_GB %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
  group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl*1000, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl*1000, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl*1000, na.rm = TRUE)
  )


# View results
print(summary_tab)
print(substrate_summary)
```

#### Break out by stream reach location:

##### BWL:

```{r, warning=F, size = 'small', echo=F, include=F}
BWL_DF<- covariat_datq_BW %>%
  dplyr::filter(site == "BWL") 

BWL_DF21 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWL") %>%
  dplyr::filter(water_year==2021)

BWL_DF22 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWL") %>%
  dplyr::filter(water_year==2022)

BWL_DF23 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWL") %>%
  dplyr::filter(water_year==2022)
  
  
summary_tab <- BWL_DF %>%
  summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

substrate_summary <- BWL_DF %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
  group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl*1000, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl*1000, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl*1000, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl*1000, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl*1000, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl*1000, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl, na.rm = TRUE)
  )


# View results
print(summary_tab)
print(substrate_summary)
```

##### BWU:

```{r, warning=F, size = 'small', echo=F, include=F}
BWU_DF <- covariat_datq_BW %>%
  dplyr::filter(site == "BWU")

BWU_DF21 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWU") %>%
  dplyr::filter(water_year==2021)

BWU_DF22 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWU") %>%
  dplyr::filter(water_year==2022)

BWU_DF23 <- covariat_datq_BW %>%
  dplyr::filter(site == "BWU") %>%
  dplyr::filter(water_year==2023)

summary_tab <- BWU_DF %>%
  summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

# Compute mean for substrate-specific nutrients
substrate_summary <- BWU_DF %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
  group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl*1000, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl*1000, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl*1000, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl*1000, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl*1000, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl*1000, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl, na.rm = TRUE)
  )

print(summary_tab)
print(substrate_summary)

```

##### GBL:

```{r, warning=F, size = 'small', echo=F, include=F}
GBL_DF <- covariat_datq_GB %>%
  dplyr::filter(site == "GBL")

GBL_DF21 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBL") %>%
  dplyr::filter(water_year==2021)

GBL_DF22 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBL") %>%
  dplyr::filter(water_year==2022)

GBL_DF23 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBL") %>%
  dplyr::filter(water_year==2023)

GBL_DF24 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBL") %>%
  dplyr::filter(water_year==2024)



summary_tab <- GBL_DF %>%
 summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

# Compute mean for substrate-specific nutrients
substrate_summary <- GBL_DF %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
   group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl*1000, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl*1000, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl*1000, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl*1000, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl*1000, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl*1000, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl, na.rm = TRUE)
  )

print(summary_tab)
print(substrate_summary)

```

##### GBU:

```{r, warning=F, size = 'small', echo=F, include=F}
GBU_DF <- covariat_datq_GB %>%
  dplyr::filter(site == "GBU")

GBU_DF21 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBU") %>%
  dplyr::filter(water_year==2021)

GBU_DF22 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBU") %>%
  dplyr::filter(water_year==2022)

GBU_DF23 <- covariat_datq_GB %>%
  dplyr::filter(site == "GBU") %>%
  dplyr::filter(water_year==2023)

summary_tab <- GBU_DF %>%
 summarise(
    mean_AFDM_mgg = round(mean(AFDM_mgg, na.rm = TRUE), 3),
    min_AFDM_mgg = min(AFDM_mgg, na.rm = TRUE),
    max_AFDM_mgg = max(AFDM_mgg, na.rm = TRUE),

    mean_Chla_ugL_Q = round(mean(Chla_ugL_Q, na.rm = TRUE), 3),
    min_Chla_ugL_Q = min(Chla_ugL_Q, na.rm = TRUE),
    max_Chla_ugL_Q = max(Chla_ugL_Q, na.rm = TRUE),

    mean_AFDM_ugcm2 = round(mean(AFDM_ugcm2, na.rm = TRUE), 3),
    min_AFDM_ugcm2 = min(AFDM_ugcm2, na.rm = TRUE),
    max_AFDM_ugcm2 = max(AFDM_ugcm2, na.rm = TRUE),

    mean_GPP = round(mean(GPP_mean, na.rm = TRUE), 3),
    min_GPP = min(GPP_mean, na.rm = TRUE),
    max_GPP = max(GPP_mean, na.rm = TRUE),

    mean_ER = round(mean(ER_mean, na.rm = TRUE), 3),
    min_ER = min(ER_mean, na.rm = TRUE),
    max_ER = max(ER_mean, na.rm = TRUE),

    mean_wtr_m = round(mean(wtr_m, na.rm = TRUE), 3),
    min_wtr_m = min(wtr_m, na.rm = TRUE),
    max_wtr_m = max(wtr_m, na.rm = TRUE),

    mean_Q_m = round(mean(Q_m, na.rm = TRUE), 3),
    min_Q_m = min(Q_m, na.rm = TRUE),
    max_Q_m = max(Q_m, na.rm = TRUE)
  )

# Compute mean for substrate-specific nutrients
substrate_summary <- GBU_DF %>%
  dplyr::filter(substrate %in% c("sw", "pw")) %>%
   group_by(substrate) %>%
  summarise(
    mean_PO4_ugL_dl = round(mean(PO4_ugL_dl, na.rm = TRUE), 3),
    min_PO4_ugL_dl = min(PO4_ugL_dl, na.rm = TRUE),
    max_PO4_ugL_dl = max(PO4_ugL_dl, na.rm = TRUE),

    mean_NO3_mgL_dl = round(mean(NO3_mgL_dl*1000, na.rm = TRUE), 3),
    min_NO3_mgL_dl = min(NO3_mgL_dl*1000, na.rm = TRUE),
    max_NO3_mgL_dl = max(NO3_mgL_dl*1000, na.rm = TRUE),

    mean_NH4_mgL_dl = round(mean(NH4_mgL_dl*1000, na.rm = TRUE), 3),
    min_NH4_mgL_dl = min(NH4_mgL_dl*1000, na.rm = TRUE),
    max_NH4_mgL_dl = max(NH4_mgL_dl*1000, na.rm = TRUE),

    mean_pH = round(mean(pH, na.rm = TRUE), 3),
    min_pH = min(pH, na.rm = TRUE),
    max_pH = max(pH, na.rm = TRUE),

    mean_DOC_mgL_dl = round(mean(DOC_mgL_dl, na.rm = TRUE), 3),
    min_DOC_mgL_dl = min(DOC_mgL_dl, na.rm = TRUE),
    max_DOC_mgL_dl = max(DOC_mgL_dl, na.rm = TRUE)
  )

print(summary_tab)
print(substrate_summary)

```

##### Kernal density plots of GPP and ER

Colored by years

```{r, warning=F, size = 'small', echo=F,include=FALSE, fig.width=7, fig.height=8}
library(ggplot2)

BWL_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWL" &water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
     scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
          annotate("text", x = 12, y = -2, label = "BWL", size = 6, fontface = "bold", hjust = 0.5) 


# 
# BWU_KD_plot <- ggplot(
#   BWU_DF %>% 
#     filter(site == "BWU" & water_year < 2025) %>%
#     mutate(water_year = factor(water_year, levels = as.character(sort(unique(water_year))))),  # Force correct order
#   aes(x = GPP_mean, y = ER_mean, color = water_year)  
# ) +
#   geom_density_2d(bins = 10, size = 1) + 
#   geom_point(alpha = 0.5) +
#   geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
#   xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#   ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
#    scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
# 
#   scale_color_viridis_d(option = "viridis") +
#   theme_classic()


# 
# BWU_KD_plot <- ggplot(
#   BWU_DF %>%
#     filter(water_year < 2024) %>%
#     mutate(water_year = factor(water_year, levels = sort(unique(water_year)))),  # Ensure correct ordering
#   aes(x = GPP_mean, y = ER_mean, color = water_year)  # `water_year` is now a properly ordered factor
# ) +
#   geom_density_2d(bins = 10, size = 1) +
#   geom_point(alpha = 0.5) +
#   geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
#   scale_x_continuous(position = "top") +  # Move x-axis to the top
#   #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#    xlab(NULL) +
#   ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
#         scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
#      scale_color_viridis_d(option = "viridis") +
#   theme_classic()

BWU_DF_fixed <- BWU_DF %>%
  filter(water_year < 2024) %>%
  mutate(water_year = factor(water_year, levels = c(2021, 2022, 2023, 2024))) %>%
  bind_rows(tibble(GPP_mean = NA, ER_mean = NA, water_year = factor(2024, levels = c(2021, 2022, 2023, 2024))))  # Add dummy row


BWU_KD_plot <- ggplot(BWU_DF_fixed, aes(x = GPP_mean, y = ER_mean, color = water_year)) +
  geom_density_2d(bins = 10, size = 1) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
  xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
  ylab(NULL) +
  #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 12, y = -2, label = "BWU", size = 6, fontface = "bold", hjust = 0.5) 
BWU_KD_plot


# 
# BWU_KD_plot <- ggplot(BWU_DF_fixed, aes(x = GPP_mean, y = ER_mean, color = water_year)) +
#   geom_density_2d(bins = 10, size = 1) +
#   geom_point(alpha = 0.5) +
#   geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
#   scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
#   #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#     #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#    xlab(NULL) +
#   ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
#   scale_color_viridis_d(option = "viridis") +
#   theme_classic()
# 


GBU_DF_clean <- GBU_DF %>%
filter(water_year<2025)%>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)
# 
# str(GBU_DF_clean)
# 
# GBU_KD_plot <- ggplot(GBU_DF_clean, aes(x = log(GPP_mean+1), y = ER_mean, color = as.factor(water_year))) +
#   stat_density_2d(size=1) + 
#   geom_point(alpha=0.25) +
#   geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
#   scale_x_continuous(position = "top") +  # Move x-axis to the top
# xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
#      scale_color_viridis_d(option = "inferno") +
#   theme_classic()



library(MASS)



# Compute 2D density
density_data <- with(GBU_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))

# Convert density data to dataframe
density_df <- data.frame(
  expand.grid(x = density_data$x, y = density_data$y),
  z = as.vector(density_data$z)
)

# Plot with contour lines
GBU_KD_plot <- ggplot() +
  geom_point(data = GBU_DF_clean, aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
    geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 5, size=1) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
    ylab(NULL) +
  #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
    scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 12, y = -2, label = "GBU", size = 6, fontface = "bold", hjust = 0.5) 

GBU_KD_plot


GBL_DF_clean <- GBL_DF %>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# 
# library(MASS)
# 
# # Compute 2D density
# density_data <- with(GBL_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))
# 
# # Convert density data to dataframe
# density_df <- data.frame(
#   expand.grid(x = density_data$x, y = density_data$y),
#   z = as.vector(density_data$z)
# )

# Plot with contour lines
GBL_KD_plot <- ggplot() +
#  geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 10) +
  geom_point(data = GBL_DF_clean%>%filter(water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_viridis_d(option = "viridis",  name = "Water year") +
  theme_classic() +
      annotate("text", x = 12, y = -2, label = "GBL", size = 6, fontface = "bold", hjust = 0.5) 


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4, fig.height=10}
# Modify each plot to rename legend title & adjust legend formatting
BWL_KD_plot <- BWL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

BWU_KD_plot <- BWU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBL_KD_plot <- GBL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBU_KD_plot <- GBU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))
```

### Modified figure 2
#### (Q1) - How does stream metabolism vary based on catchment characteristics, antecedent flow conditions, and nitrogen availability?

```{r, warning=F, size = 'small', echo=F, fig.width=6, fig.height=7}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  BWL_KD_plot,
  BWU_KD_plot,
  GBL_KD_plot,
  GBU_KD_plot,
  labels = c("a", "b", "c", "d"),
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom"
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig2_KD_CH1_grid_v2.png", plot = KD_coeff_grid, width = 6, height = 5.25, units = "in")

```

### General drivers of productivity

Both standing stocks

-   A. Epilithic biomass

-   B. Epilithic chl-a

Ecosystem fluxes

-   C. Gross primary productivity (GPP)

-   E. Absolute value of ecosystem respiration (ER)

Each productivity response is individually modeled within stream using
`glm`

dataframe = `covariat_datq`

#### Fixed effect co-variance ?

Maybe some with flow and water temp but all correlations are under 0.6

##### At Blackwood:

```{r, warning=F, size = 'small', echo=T, fig.width=9, fig.height=9}
cor_data <- covariat_datq_BW[, c(6, 11:14, 18, 28, 30)]

chart.Correlation(cor_data, histogram=TRUE, pch=19)
```

##### At Glenbrook:

```{r, warning=F, size = 'small', echo=T, fig.width=9, fig.height=9}
cor_data <- covariat_datq_GB[, c(6, 11:14, 18, 28, 30)]

chart.Correlation(cor_data, histogram=TRUE, pch=19)
```

### A. Biomass models

##### BW:

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4, fig.height=4}
#What's the response distribution? 
#hist(covariat_datq_BW$AFDM_ugcm2)
#hist(log(covariat_datq_BW$AFDM_ugcm2+1))
```

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
biomass_mod_bw <- lmer(log(AFDM_ugcm2+1)~ 
                         scale(NO3_mgL_dl* 1000) + 
                         scale(NH4_mgL_dl* 1000) + 
                         scale(PO4_ugL_dl) + 
                        scale(AFDM_mgg)+
                        # scale(DOC_mgL_dl)+
                         scale(wtr_m) + 
                         scale(Q_m) + (1|site), data=covariat_datq_BW)

summary(biomass_mod_bw)
vif(biomass_mod_bw)
#hist(residuals(biomass_mod_bw), main = paste(NULL))
r.squaredGLMM(biomass_mod_bw)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=6, fig.height=4}
# Extract fixed effects and confidence intervals
fixed_effects <- broom.mixed::tidy(biomass_mod_bw, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects1 <- fixed_effects[fixed_effects$term != "(Intercept)", ]
fixed_effects1 <- fixed_effects1[order(fixed_effects1$estimate), ]  # Order by estimate
fixed_effects1$Catchment <- "BW"
```

##### GB:

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
biomass_mod_gb <- lmer(log(AFDM_ugcm2+1)~ 
                         scale(NO3_mgL_dl* 1000) + 
                         scale(NH4_mgL_dl* 1000) + 
                         scale(PO4_ugL_dl) + 
                         scale(AFDM_mgg)+
                         #scale(DOC_mgL_dl)+
                         scale(wtr_m) + 
                         scale(Q_m)+(1|site), data=covariat_datq_GB)

summary(biomass_mod_gb)
vif(biomass_mod_gb)
#hist(residuals(biomass_mod_gb), main = paste(NULL))
r.squaredGLMM(biomass_mod_gb)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=6, fig.height=4}
fixed_effects_gb <- broom.mixed::tidy(biomass_mod_gb, effects = "fixed", conf.int = TRUE)

# Filter to exclude intercept and arrange predictors
fixed_effects_gb1 <- fixed_effects_gb[fixed_effects_gb$term != "(Intercept)", ]
fixed_effects_gb1 <- fixed_effects_gb1[order(fixed_effects_gb1$estimate), ]  # Order by estimate
fixed_effects_gb1$Catchment <- "GB"
```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}

fxd_eff <- rbind(fixed_effects_gb1, fixed_effects1)

fxd_eff1 <- fxd_eff %>%
  mutate(Significance = case_when(
    p.value <= 0.05 ~ "sig",
    p.value > 0.05 & p.value <= 0.1  ~ "marg.",
    p.value > 0.1  ~ "not sig."))


term_labels <- c(
  "scale(NO3_mgL_dl * 1000)" = expression(NO[3]~(µg~L^-1)),
  "scale(PO4_ugL_dl)" = expression(PO[4]~(µg~L^-1)),
 # "scale(DOC_mgL_dl)" = expression(DOC~(mg~L^-1)),
  "scale(AFDM_mgg)" = expression(OM~(mg~g^-1)),
  "scale(NH4_mgL_dl * 1000)" = expression(NH[4]~(µ~L^-1)),
  "scale(wtr_m)" = expression(Temp~(degree~C)),
  "scale(Q_m)" = expression(Q~(m^3~s^-1))
)

# Add an offset column based on Catchment
fxd_eff1 <- fxd_eff1 %>%
  mutate(y_position = as.numeric(reorder(term, estimate)) + ifelse(Catchment == "BW", 0.2, 0))

Biomass_coeff_plot <-ggplot(fxd_eff1, aes(x = estimate, y = reorder(term, estimate), color = Catchment, shape = Significance)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, alpha = 0.85, position = position_dodge(width = 0.3)) +
  theme_classic() +
  scale_shape_manual(values = c(4, 19)) +
  scale_color_manual(values = catch_colors) +
  geom_vline(xintercept = 0) +
  labs(
    x = NULL,
    y = "Predictors",
    title = "Epilithic biomass"
  ) +
  scale_y_discrete(labels = term_labels) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12)
  )


Biomass_coeff_plot
```

### B. Chlorophyll-a models

##### BW:

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}

chla_mod_bw <- lmer(log(Chla_ugL_Q+1)~ 
                         scale(NO3_mgL_dl* 1000) + 
                         scale(NH4_mgL_dl* 1000) + 
                         scale(PO4_ugL_dl)+ 
                     scale(AFDM_mgg)+
                         #scale(DOC_mgL_dl)+
                         scale(wtr_m) + 
                         scale(Q_m)+(1|site), data=covariat_datq_BW)

summary(chla_mod_bw)
vif(chla_mod_bw)
#hist(residuals(chla_mod_bw), main = paste(NULL))
r.squaredGLMM(chla_mod_bw)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=6, fig.height=4}
# Extract fixed effects and confidence intervals
fixed_effects_chl <- broom.mixed::tidy(chla_mod_bw, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_chl1 <- fixed_effects_chl[fixed_effects_chl$term != "(Intercept)", ]
fixed_effects_chl1 <- fixed_effects_chl1[order(fixed_effects_chl1$estimate), ]  # Order by estimate
fixed_effects_chl1$Catchment <- "BW"

```

##### GB:


```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}

chla_mod_gb <- lmer(log(Chla_ugL_Q+1)~ 
                      scale(NO3_mgL_dl* 1000) + 
                      scale(NH4_mgL_dl* 1000) + 
                      scale(PO4_ugL_dl)+
                     scale(AFDM_mgg)+
                      #scale(DOC_mgL_dl)+
                      scale(wtr_m) + 
                      scale(Q_m) + (1|site), data=covariat_datq_GB)

summary(chla_mod_gb)
vif(chla_mod_gb)
#hist(residuals(chla_mod_gb), main = paste(NULL))
r.squaredGLMM(chla_mod_gb)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=6, fig.height=4}

fixed_effects_chlg <- broom.mixed::tidy(chla_mod_gb, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_chlg1 <- fixed_effects_chlg[fixed_effects_chlg$term != "(Intercept)", ]
fixed_effects_chlg1 <- fixed_effects_chlg1[order(fixed_effects_chlg1$estimate), ]  # Order by estimate
fixed_effects_chlg1$Catchment <- "GB"
```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
fxd_eff_chla <- rbind(fixed_effects_chlg1, fixed_effects_chl1)

fxd_eff_chla1 <- fxd_eff_chla %>%
  mutate(Significance = case_when(
    p.value <= 0.05 ~ "sig.",
    p.value > 0.05 & p.value <= 0.1  ~ "marg.",
    p.value > 0.1  ~ "not sig."))


term_labels_chla <- c(
  "scale(NO3_mgL_dl * 1000)" = expression(NO[3]~(µg~L^-1)),
  "scale(PO4_ugL_dl)" = expression(PO[4]~(µg~L^-1)),
   # "scale(DOC_mgL_dl)" = expression(DOC~(mg~L^-1)),
  "scale(AFDM_mgg)" = expression(OM~(mg~g^-1)),
  "scale(NH4_mgL_dl * 1000)" = expression(NH[4]~(µ~L^-1)),
  "scale(wtr_m)" = expression(Temp~(degree~C)),
  "scale(Q_m)" = expression(Q~(m^3~s^-1))
)

# Add an offset column based on Catchment
fxd_eff_chla1 <- fxd_eff_chla1 %>%
  mutate(y_position = as.numeric(reorder(term, estimate)) + ifelse(Catchment == "BW", 0.2, 0))


chla_coeff_plot <-ggplot(fxd_eff_chla1, aes(x = estimate, y = reorder(term, estimate), color = Catchment, shape = Significance)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, alpha = 0.85, position = position_dodge(width = 0.3)) +
  theme_classic() +
  scale_shape_manual(values = c(1,4, 19)) +
  scale_color_manual(values = catch_colors) +
  geom_vline(xintercept = 0) +
  labs(
    x = NULL,
    y = NULL,
    title = "Epilithic chl-a"
  ) +
  scale_y_discrete(labels = term_labels_chla) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12)
  )
chla_coeff_plot

#########
```

### C. GPP models

##### BW:

```{r, warning=F, size = 'small', echo=T, include=F, fig.width=6, fig.height=4}
# What's the response distribution? 

covariat_datq_BW_na <- covariat_datq_BW %>%
  drop_na(GPP_mean) %>%
  mutate(ab_ER = ER_mean*-1)

covariat_datq_BW_na$Q_m <- (log(covariat_datq_BW_na$Q_m+1))

# Possibility still too skewed to use a normal distribution in lmer. 

```

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
gpp_mod_bw <- lmer(log(GPP_mean+1) ~ 
                      scale(NO3_mgL_dl* 1000) + 
                      scale(NH4_mgL_dl* 1000) + 
                      scale(PO4_ugL_dl)+
                      scale(AFDM_mgg)+
                      #scale(DOC_mgL_dl)+
                      scale(wtr_m) +
                      scale(Q_m)+ (1|site), 
                  data=covariat_datq_BW_na)

summary(gpp_mod_bw)
vif(gpp_mod_bw)
#hist(residuals(gpp_mod_bw),main = paste(NULL))
r.squaredGLMM(gpp_mod_bw)
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=6, fig.height=4}

# Extract fixed effects and confidence intervals
fixed_effects_nep <- broom.mixed::tidy(gpp_mod_bw, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_nep <- fixed_effects_nep[fixed_effects_nep$term != "(Intercept)", ]
fixed_effects_nep <- fixed_effects_nep[order(fixed_effects_nep$estimate), ]  # Order by estimate
fixed_effects_nep$Catchment <- "BW"
```

##### GB:

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
# What's the response distribution? 

covariat_datq_GB_na <- covariat_datq_GB %>%
    drop_na(GPP_mean) %>%
  mutate(ab_ER = ER_mean*-1)

  #drop_na(GPP_mean)
```

Highly skewed even when log transformed.

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
gpp_mod_gb <- lmer(log(GPP_mean+1) ~ 
                      scale(NO3_mgL_dl* 1000) + 
                      scale(NH4_mgL_dl* 1000) + 
                      scale(PO4_ugL_dl)+ 
                             scale(AFDM_mgg)+
                      #scale(DOC_mgL_dl)+
                      scale(wtr_m) + 
                      scale(Q_m) + (1|site), data=covariat_datq_GB_na)

summary(gpp_mod_gb)
vif(gpp_mod_gb)
#hist(residuals(gpp_mod_gb), main = paste(NULL))
r.squaredGLMM(gpp_mod_gb)

```

```{r, warning=F, size = 'small', echo=T, fig.width=6, fig.height=4}

fixed_effects_nep1 <- broom.mixed::tidy(gpp_mod_gb, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_nep1 <- fixed_effects_nep1[fixed_effects_nep1$term != "(Intercept)", ]
fixed_effects_nep1 <- fixed_effects_nep1[order(fixed_effects_nep1$estimate), ]  # Order by estimate
fixed_effects_nep1$Catchment <- "GB"
```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
fxd_eff_nep <- rbind(fixed_effects_nep, fixed_effects_nep1)

fxd_eff_nep1 <- fxd_eff_nep %>%
  mutate(Significance = case_when(
    p.value <= 0.05 ~ "sig.",
    p.value > 0.05 & p.value <= 0.1  ~ "marg.",
    p.value > 0.1  ~ "not sig."))


term_labels_nep <- c(
  "scale(NO3_mgL_dl * 1000)" = expression(NO[3]~(µg~L^-1)),
  "scale(PO4_ugL_dl)" = expression(PO[4]~(µg~L^-1)),
    "scale(AFDM_mgg)" = expression(OM~(mg~g^-1)),
 # "scale(DOC_mgL_dl)" = expression(DOC~(mg~L^-1)),
  "scale(NH4_mgL_dl * 1000)" = expression(NH[4]~(µ~L^-1)),
  "scale(wtr_m)" = expression(Temp~(degree~C)),
  "scale(Q_m)" = expression(Q~(m^3~s^-1))
)

# Add an offset column based on Catchment
fxd_eff_nep1 <- fxd_eff_nep1 %>%
  mutate(y_position = as.numeric(reorder(term, estimate)) + ifelse(Catchment == "BW", 0.2, 0))


nep_coeff_plot <-ggplot(fxd_eff_nep1, aes(x = estimate, y = reorder(term, estimate), color = Catchment, shape = Significance)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, alpha = 0.85, position = position_dodge(width = 0.3)) +
  theme_classic() +
  scale_shape_manual(values = c(1,4, 19)) +
  scale_color_manual(values = catch_colors) +
  geom_vline(xintercept = 0) +
  labs(
    x = "Scaled effect size",
    y = "Predictors",
    title = "GPP"
  ) +
  scale_y_discrete(labels = term_labels_nep) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12)
  )
```

### D. \|ER\| models

##### BW:

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
er_mod_bw <- lmer(log(ab_ER+1) ~ 
                      scale(NO3_mgL_dl* 1000) + 
                      scale(NH4_mgL_dl* 1000) + 
                      scale(PO4_ugL_dl)+
                      scale(AFDM_mgg)+
                      #scale(DOC_mgL_dl)+
                      scale(wtr_m)  +
                      scale(Q_m) + (1|site), 
                  data=covariat_datq_BW_na)

summary(er_mod_bw)
vif(er_mod_bw)
#hist(residuals(er_mod_bw), main = paste(NULL))
r.squaredGLMM(er_mod_bw)

```

```{r, warning=F, size = 'small', echo=T, fig.width=6, fig.height=4}
# Extract fixed effects and confidence intervals
fixed_effects_er <- broom.mixed::tidy(er_mod_bw, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_er <- fixed_effects_er[fixed_effects_er$term != "(Intercept)", ]
fixed_effects_er <- fixed_effects_er[order(fixed_effects_er$estimate), ]  # Order by estimate
fixed_effects_er$Catchment <- "BW"
```

##### GB:

```{r, warning=F, size = 'small', echo=T, fig.width=3, fig.height=3.5}
er_mod_gb <- lmer(log(ab_ER+1) ~ 
                      scale(NO3_mgL_dl* 1000) + 
                      scale(NH4_mgL_dl* 1000) + 
                      scale(PO4_ugL_dl)+
                      scale(AFDM_mgg)+
                     # scale(DOC_mgL_dl)+
                      scale(wtr_m) +
                      scale(Q_m)+ (1|site), 
                  data=covariat_datq_GB_na)

summary(er_mod_gb)
vif(er_mod_gb)
#hist(residuals(er_mod_gb), main = paste(NULL))
r.squaredGLMM(er_mod_gb)
```

```{r, warning=F, size = 'small', echo=T, fig.width=6, fig.height=4}

fixed_effects_er1 <- broom.mixed::tidy(er_mod_gb, effects = "fixed", conf.int = TRUE)
# Filter to exclude intercept and arrange predictors
fixed_effects_er1 <- fixed_effects_er1[fixed_effects_er1$term != "(Intercept)", ]
fixed_effects_er1 <- fixed_effects_er1[order(fixed_effects_er1$estimate), ]  # Order by estimate
fixed_effects_er1$Catchment <- "GB"
```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
fxd_eff_er <- rbind(fixed_effects_er, fixed_effects_er1)

fxd_eff_er1 <- fxd_eff_er %>%
  mutate(Significance = case_when(
    p.value <= 0.05 ~ "sig.",
    p.value > 0.05 & p.value <= 0.1  ~ "marg.",
    p.value > 0.1  ~ "not sig."))


term_labels_er <- c(
  "scale(NO3_mgL_dl * 1000)" = expression(NO[3]~(µg~L^-1)),
  "scale(PO4_ugL_dl)" = expression(PO[4]~(µg~L^-1)),
    "scale(AFDM_mgg)" = expression(OM~(mg~g^-1)),
 # "scale(DOC_mgL_dl)" = expression(DOC~(mg~L^-1)),
  "scale(NH4_mgL_dl * 1000)" = expression(NH[4]~(µ~L^-1)),
  "scale(wtr_m)" = expression(Temp~(degree~C)),
  "scale(Q_m)" = expression(Q~(m^3~s^-1))
)

# Add an offset column based on Catchment
fxd_eff_er1 <- fxd_eff_er1 %>%
  mutate(y_position = as.numeric(reorder(term, estimate)) + ifelse(Catchment == "BW", 0.2, 0))

er_coeff_plot <-ggplot(fxd_eff_er1, aes(x = estimate, y = reorder(term, estimate), color = Catchment, shape = Significance)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0, alpha = 0.85, position = position_dodge(width = 0.3)) +
  theme_classic() +
  scale_shape_manual(values = c(1,4, 19)) +
  scale_color_manual(values = catch_colors) +
  geom_vline(xintercept = 0) +
  labs(
    x = "Scaled effect size",
    y = NULL,
    title = "|ER|"
  ) +
  scale_y_discrete(labels = term_labels_er) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.title.x = element_text(size = 12)
  )

er_coeff_plot

```

### Figure 3: General drivers for biomass, chl-a and GPP

#### (Q1) - How does stream metabolism vary based on catchment characteristics, antecedent flow conditions, and nitrogen availability?

Mainly addresses how productivity is associated with nutrients and water
quality:

```{r, warning=F, size = 'small', echo=F, fig.width=6, fig.height=6}

### big grid 
coeff_grid <- ggarrange(
  Biomass_coeff_plot,
  chla_coeff_plot,
  nep_coeff_plot,
  er_coeff_plot,
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  labels=c("a", "b", "c", "d"),
  legend = "bottom")
# 
# 
# coeff_grid <- ggarrange(
#   Biomass_coeff_plot + theme(legend.position = "bottom") + 
#     guides(color = guide_legend(nrow = 2), shape = guide_legend(nrow = 3)),
#   chla_coeff_plot + theme(legend.position = "bottom") + 
#     guides(color = guide_legend(nrow = 2), shape = guide_legend(nrow = 3)),
#   nep_coeff_plot + theme(legend.position = "bottom") + 
#     guides(color = guide_legend(nrow = 2), shape = guide_legend(nrow = 3)),
#   er_coeff_plot + theme(legend.position = "bottom") + 
#     guides(color = guide_legend(nrow = 3), shape = guide_legend(nrow = 4)),
#   labels=c("a", "b", "c", "d"),
#   ncol = 1, nrow = 4,
#   common.legend = TRUE,
#   legend = "bottom"
# )

coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig4_coeff_ecol_CH1_grid.png", plot = coeff_grid, width = 5.5, height = 5.25, units = "in")

```

In general GPP may be more a function of physical conditions but is
partially associated with NH4.

# Part2/ Question 2

### Nitrogen demand and supply dynamics

#### Calculate nitrogen supply and demand:

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
# Constants
ra <- 0.5  # Autotrophic respiration coefficient (Hall & Tank 2003)
C_Nauto <- 16  # Autotrophic C:N ratio (Stelzer & Lamberti 2001)
C_Nhetero <- 20  # Heterotrophic C:N ratio (Hall & Tank 2003)
HGE <- 0.05  # Heterotrophic Growth Efficiency (Hall & Tank 2003)

# Calculate components of nitrogen demand
covariat_nitrogen2 <- covariat_datq %>%
  filter(substrate=="sw")%>% # select just surface water samples
  group_by(site) %>% 
  arrange(date) %>%  # Ensure chronological order
  mutate( ## 17 NAs
    v_mi = ifelse(is.na(v_m), 
                  rollapplyr(v_m, width = 3, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE), 
                  v_m),
    w_mi = ifelse(is.na(w_m), 
                  rollapplyr(w_m, width = 3, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE), 
                  w_m)
  ) %>%
  # Carry forward any remaining NAs
  mutate(
    v_mi = zoo::na.locf(v_mi, na.rm = FALSE),
    v_mi = zoo::na.locf(v_mi, fromLast = TRUE, na.rm = FALSE),
    w_mi = zoo::na.locf(w_mi, na.rm = FALSE),
    w_mi = zoo::na.locf(w_mi, fromLast = TRUE, na.rm = FALSE)
  ) %>%
  # If there are still missing values, replace them with the site mean
  mutate(
    v_mi = ifelse(is.na(v_mi), mean(v_m, na.rm = TRUE), v_mi),
    w_mi = ifelse(is.na(w_mi), mean(w_m, na.rm = TRUE), w_mi)
  ) %>%
  ungroup()%>%
  ## should already be at detection limits
  # filter(water_year < 2024) %>%
  # mutate(NO3_mgL_dl = if_else(NO3_mgL_dl < 0.0015, 0.0015, NO3_mgL_dl))%>%
  #   mutate(NH4_mgL_dl=if_else(NH4_mgL_dl < 0.001, 0.001, NH4_mgL_dl))%>%
  #   mutate(PO4_ugL_dl=if_else(PO4_ugL_dl < 0.201, 0.201, PO4_ugL_dl))%>%
  mutate(
    
    ## ** Pause here to double check Q units
    Q_Ls = Q_m * 1000, # flow from csm to Ls
    # Autotrophic respiration (raGPP)
    raGPP = GPP_mean * ra,
    # Autotrophic assimilation of N
    Auto_N_assim = GPP_mean / C_Nauto,
    # Heterotrophic respiration (Rh)
    Rh = ER_mean - raGPP,
    # Heterotrophic assimilation of N
    Hetero_N_assim = (Rh * HGE) / C_Nhetero,
    # Total nitrogen demand
    Ndemand = Auto_N_assim + Hetero_N_assim, # unit should be g N m-2 d-1
    # calculate reach length in m
    reachL = c(((v_mi*10)*w_mi*86.4)/K600_daily_mean),
    Ndemand = if_else(Ndemand < 0, 0.0001, Ndemand), # have demand be essentially zero 
   # Calculate no3 supply 
   NO3_supply = c(((86400*Q_Ls*NO3_mgL_dl)/(w_mi*reachL))/1000),
   # Calculate nh3 supply 
   NH4_supply = c(((86400*Q_Ls*NH4_mgL_dl)/(w_mi*reachL))/1000), ## unit should be g N m-2 d-1
   PO4_supply = c(((86400*Q_Ls*(PO4_ugL_dl/1000))/(w_mi*reachL))/1000) ## unit should be g N m-2 d-1
   )


```

### Plots of Ratios of NH4+- N supply to demand, ratios NO3-- N supply to demand

```{r, warning=F, size = 'small', echo=T, include=F}
covariat_NS_ND <- covariat_nitrogen2%>%
  mutate(jday = yday(date))%>%
  dplyr:: select(date, jday, site, catch, water_year, Ndemand, NO3_supply, NH4_supply) %>%
  mutate(NO3_SupDem = (NO3_supply/Ndemand),
         NH4_SupDem = (NH4_supply/Ndemand))

N_ratio_plots <- ggplot(covariat_NS_ND, aes(x = log(NO3_supply+1), y = Ndemand, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    geom_point( aes(x = log(NH4_supply+1), y = Ndemand, color = site),
                size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
  scale_color_manual(values = siteC_colors) +
      scale_shape_manual(values = c(15,0,17,2)) +
  xlim(0,1) +
   ylab(expression(Nitrogen~demand~(g~N~m^-2~d^-1))) +
     xlab(expression(log(Nitrogen~supply+1)~(g~N~m^-2~d^-1))) +
  theme_classic()


NO3_supp_plots <- ggplot(covariat_NS_ND%>%filter(NO3_supply<60), aes(y = log(NO3_supply+1), x = jday, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    geom_point( aes(y = log(NH4_supply+1), x = jday, color = site, shape=site), 
                size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = siteC_colors) +
  xlim(15,360) +
  #  ylim(0,5) +
   xlab(NULL) +
     ylab(expression(log(Supply+1)~(g~N~m^-2~d^-1))) +
  theme_classic() +
  scale_shape_manual(values = c(15,0,17,2)) +
    theme(legend.position = "none")  # Remove legend

N_demand_plot1 <- ggplot(covariat_NS_ND, aes(y = (Ndemand), x = jday, color = site,  shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.8)) +
  scale_color_manual(values = siteC_colors) +
   xlab("Day of year") +
    xlim(15,360) +
    scale_shape_manual(values = c(15,0,17,2)) +
   ylab(expression(Demand~~(g~N~m^-2~d^-1))) +
  theme_classic() +
    theme(legend.position = "none")  # Remove legend
```

##### Summary of reach based surface water supply

```{r, warning=F, size = 'small', echo=F, include=T}
## Summary stats for days with N supply and demand

covariat_NS_check <- covariat_NS_ND%>%
  filter(NO3_SupDem<=0.9)
length(unique(covariat_NS_check$date))

covariat_NH4_check <- covariat_NS_ND%>%
  filter(NH4_SupDem<=0.9)
length(unique(covariat_NH4_check$date))


covariat_NS_check2 <- covariat_NS_ND%>%
  filter(NO3_SupDem>=0.99)

length(unique(covariat_NS_check2$date))


covariat_NS_ND_sum <- covariat_nitrogen2%>%
  group_by(catch)%>%
  summarise(
    N_supplym =mean(NO3_supply+NH4_supply, na.rm=T),
    NO3_supplym =mean(NO3_supply, na.rm=T),
            NO3_supplymin =min(NO3_supply, na.rm=T),
           NO3_supplymax =max(NO3_supply, na.rm=T),
            NH4_supplym =mean(NH4_supply, na.rm=T),
           NH4_supplymin =min(NH4_supply, na.rm=T),
            NH4_supplymax =max(NH4_supply, na.rm=T),
          PO4_supplym =mean(PO4_supply, na.rm=T),
           PO4_supplymin =min(PO4_supply, na.rm=T),
            PO4_supplymax =max(PO4_supply, na.rm=T)
          )
covariat_NS_ND_sum
```

## Figure 4: Nitrogen supply and demand seasonally at each reach

### (Q2) - Is variance in nitrogen demand and instream uptake (either as ammonium or nitrate) more related to biological (GPP, ER, biomass or chlorophyll-a), physical processes (flow and water temperature), or ambient nutrient and organic matter availability? 

```{r, warning=F, size = 'small', echo=T, fig.width=8, fig.height=4}

# n_s_grid <- ggarrange(
#   NO3_supp_plots,
#   N_demand_plot1,
#   labels = c("b","c"),
#   ncol = 1, nrow = 2)


n_s_grid <- ggarrange(
  NO3_supp_plots,
  N_demand_plot1,
  labels = c("b", "c"),
  ncol = 1, nrow = 2,
  label.x = 0.95,  # Move label to the right
  label.y = 1,     # Keep label at the top
  hjust = 1,       # Right-align the label
  vjust = 1        # Top-align the label
)

# 
# # Create the top row with extra spacing on both sides
# #left_col <- ggdraw() + draw_plot(N_ratio_plots, x = 0.15, height = 0.85)  # Centered with 25% narrower width
# left_col <- ggdraw() + 
#   draw_plot(N_ratio_plots, x = 0.15, y = 0.075, height = 0.85) 

### big grid 
Ndyn_grid <- ggarrange(
  N_ratio_plots,
  n_s_grid,
  ncol = 2, nrow = 1,
    labels = c("a", ""),
  common.legend = TRUE, 
   widths = c(1.75, 2.25),
  legend = "bottom")

Ndyn_grid

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Nsupply_grid_CH1.png", plot = Ndyn_grid, width = 9.25, height = 4.75, units = "in")

```




### Plots of Ratios of NH4+- N supply to demand, ratios NO3-- N supply to demand colored by DOY

```{r, warning=F, size = 'small', echo=T, include=F}
covariat_NS_ND <- covariat_nitrogen2%>%
  mutate(jday = yday(date))%>%
  dplyr:: select(date, jday, site, catch, water_year, Ndemand, NO3_supply, NH4_supply) %>%
  mutate(NO3_SupDem = (NO3_supply/Ndemand),
         NH4_SupDem = (NH4_supply/Ndemand))


## adjust the demand y axis to be from 0 to 1 

N_ratio_plots <- ggplot(covariat_NS_ND, aes(x = log(NO3_supply+1), y = Ndemand, color = jday, shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    geom_point( aes(x = log(NH4_supply+1), y = Ndemand, color = jday),
                size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
    scale_color_viridis_c(option = "magma",  name = "Day of year") +
      scale_shape_manual(values = c(15,0,17,2)) +
  xlim(0,1) +
  ylim(0, 1) +
   ylab(expression(Nitrogen~demand~(g~N~m^-2~d^-1))) +
     xlab(expression(log(Nitrogen~supply+1)~(g~N~m^-2~d^-1))) +
  theme_classic()


NO3_supp_plots <- ggplot(covariat_NS_ND%>%filter(NO3_supply<60), aes(y = log(NO3_supply+1), x = jday, color = jday, shape=site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    geom_point( aes(y = log(NH4_supply+1), x = jday, color = jday, shape=site), 
                size = 3, alpha = 0.85, position = position_dodge(width = 0.3)) +
    scale_color_viridis_c(option = "viridis", name = "Day of year") +  # Use _c for continuous data
  xlim(15,360) +
  #  ylim(0,5) +
   xlab(NULL) +
     ylab(expression(log(Supply+1)~(g~N~m^-2~d^-1))) +
  theme_classic() +
  scale_shape_manual(values = c(15,0,17,2)) +
    theme(legend.position = "none")  # Remove legend


N_demand_plot1 <- ggplot(covariat_NS_ND, aes(y = (Ndemand), x = jday, color = jday, shape = site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.8)) +
  scale_color_viridis_c(option = "viridis", name = "Day of year") +  # Use _c for continuous data
  xlab("Day of year") +
  xlim(15, 360) +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  ylab(expression(Demand~~(g~N~m^-2~d^-1))) +
  theme_classic() +
  theme(legend.position = "none")  # Remove legend

N_demand_plot1

```


```{r, warning=F, size = 'small', echo=T, fig.width=8, fig.height=4}

n_s_grid <- ggarrange(
  NO3_supp_plots,
  N_demand_plot1,
  labels = c("b", "c"),
  ncol = 1, nrow = 2,
  label.x = 0.95,  # Move label to the right
  label.y = 1,     # Keep label at the top
  hjust = 1,       # Right-align the label
  vjust = 1        # Top-align the label
)


Ndyn_grid <- ggarrange(
  N_ratio_plots,
  n_s_grid,
  ncol = 2, nrow = 1,
    labels = c("a", ""),
  common.legend = TRUE, 
   widths = c(1.75, 2.25),
  legend = "bottom")

Ndyn_grid

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Nsupply_grid_CH1_v2.png", plot = Ndyn_grid, width = 9.25, height = 4.75, units = "in")

```



##### Calculations of flow normalized NO3 and NH4 average daily loads by water year:

```{r, warning=F, size = 'small', echo=T, include=FALSE,fig.width=6, fig.height=4}
WY_nloads <- covariat_nitrogen2 %>%
  mutate(mon = month(date)) %>%
  filter(mon > 6 & mon < 11) %>%
  filter(water_year < 2024) %>%
  group_by(site, water_year) %>%
  mutate(Ls = Q_m * 1000,  # Convert cubic meters per second to liters per second
         NO3_N_sec = NO3_mgL_dl * Ls,  # mg/L * L/s = mg/s
         NH4_N_sec = NH4_mgL_dl * Ls,  
         NO3_N_mg = NO3_N_sec * 86.4,  # / 1e6 if want kg Convert mg/s to kg/day (1 kg = 1e6 mg)
         NH4_N_mg = NH4_N_sec * 86.4) %>%
  summarise(NO3_Q = mean(NO3_N_mg, na.rm = TRUE),
            NO3_Qsd = sd(NO3_N_mg, na.rm = TRUE),
            NH4_Q = mean(NH4_N_mg, na.rm = TRUE),
            NH4_Qsd = sd(NH4_N_mg, na.rm = TRUE))

WY_nloads
```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
n_up_sum <-n_up%>%
  filter(success=="yes") %>%
  dplyr::group_by(method, site) %>%
  dplyr::summarise(
    Uadd_ug_L = mean(Uadd_ug_L_min, na.rm=T),
    Uadd_ug_L_MIN = min(Uadd_ug_L_min, na.rm=T),
    Uadd_ug_L_MAX = max(Uadd_ug_L_min, na.rm=T),
    sw = mean(sw_m, na.rm=T),
    sw_MIN = min(sw_m, na.rm=T),
    sw_MAX = max(sw_m, na.rm=T),
    Vf_m = mean(Vf_m_min, na.rm=T),
    Vf_m_MIN = min(Vf_m_min, na.rm=T),
    Vf_m_MAX = max(Vf_m_min, na.rm=T))

nup_sum_plot <-ggplot(n_up_sum, aes(x = site, y = Uadd_ug_L, color = site, shape = method)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Uadd_ug_L_MIN, ymax = Uadd_ug_L_MAX),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() + ylab(expression(U[t]~(µg~m^-2~min^-1)))+
  # ylab(expression(U[t]~(µg~L^-1~min^-1)))+
  xlab(NULL) +
  #scale_shape_manual(values = c(4, 19)) +
  scale_color_manual(values = siteC_colors) #+ theme(axis.text.x = element_text(angle = 45, hjust = 1))


nvf_sum_plot <-ggplot(n_up_sum, aes(x = site, y = Vf_m, color = site, shape = method)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = Vf_m_MIN, ymax = Vf_m_MAX),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic()  + ylab(expression(V[f]~(m~min^-1)))+
 # ylab(expression(V[f]~(µg~m^-1~min^-1)))+
  xlab(NULL) +
  #scale_shape_manual(values = c(4, 19)) +
  scale_color_manual(values = siteC_colors) #+ theme(axis.text.x = element_text(angle = 45, hjust = 1))


nsw_sum_plot <-ggplot(n_up_sum, aes(x = site, y = sw, color = site, shape = method)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin = sw_MIN, ymax = sw_MAX),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic()  + ylab(expression(S[w]~(m)))+
  xlab(NULL) +
  #scale_shape_manual(values = c(4, 19)) +
  scale_color_manual(values = siteC_colors) #+ theme(axis.text.x = element_text(angle = 45, hjust = 1))



```

##### Plot for nitrogen cycling by reach location

Figure 5 a-c

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=5, fig.height=3}
# Combine plots with a shared legend
Nup_grid <- ggarrange(
  nup_sum_plot,
  nvf_sum_plot,
  nsw_sum_plot,
  ncol = 3, nrow = 1,
  labels=c("a", "b", "c"), 
  common.legend = TRUE,
  legend = "right"
)

Nup_grid <- Nup_grid + 
  theme(legend.box = "vertical",        # Arrange legends vertically
        legend.text = element_text(size = 10),  # Adjust legend text size if needed
        legend.title = element_text(size = 12)) +
  guides(
    color = guide_legend(nrow = 1),    # Put colors on one row
    shape = guide_legend(nrow = 1)    # Put shapes on another row
  )

Nup_grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Nuptake_CH1_gridv2.png", plot = Nup_grid, width = 7, height = 2.5, units = "in")
```

#### Correlations between N uptake and ...

-   

    (d) biomass

-   

    (e) chl-a

-   

    (f) GPP

-   

    (g) ER

-   

    (h) Q

-   

    (i) temp

###### Dataframe for successfull N uptatke

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}
covariat_N <- covariat_datq %>%
  filter(Uadd_ug_L_min < 121) %>%
  filter(success=="yes")%>%
  dplyr::select(site, catch, date, water_year, Chla_ugL_Q, method, Uadd_ug_L_min, 
         Uadd_min_sd, sw_m, sw_sd, Vf_m_min, vf_min_sd, GPP_mean, ER_mean,
         AFDM_ugcm2, Q_m, wtr_m, AFDM_mgg) %>%
    drop_na(Uadd_ug_L_min) 

covariat_NU <- covariat_N %>% distinct()


#hist(covariat_NU$Uadd_ug_L_min,  main = paste(NULL))
#hist(log(covariat_NU$Uadd_ug_L_min+1),  main = paste(NULL))

covariat_NU_B <- covariat_NU%>%
  filter(catch=="BW")

covariat_NU_G <- covariat_NU%>%
  filter(catch=="GB")

```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}
U_mod_OM <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(AFDM_mgg) + (1|site), data=covariat_NU_B)
summary(U_mod_OM)
#hist(residuals(U_mod_OM),main = paste(NULL))

U_mod_bio_r2_vals <- r.squaredGLMM(U_mod_OM)
U_mod_bio_r2_vals
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}
U_mod_OM <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(AFDM_mgg) + (1|site), data=covariat_NU_G)
summary(U_mod_OM)
#hist(residuals(U_mod_OM),main = paste(NULL))

U_mod_bio_r2_vals <- r.squaredGLMM(U_mod_OM)
U_mod_bio_r2_vals
```

#### d) Uptake and epilithic biomass:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}
U_mod_bio <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(AFDM_ugcm2) + (1|site), data=covariat_NU_B)
summary(U_mod_bio)
#hist(residuals(U_mod_bio),main = paste(NULL))

U_mod_bio_r2_vals <- r.squaredGLMM(U_mod_bio)
U_mod_bio_conditional_r2 <- round(U_mod_bio_r2_vals[2], 2)

```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}
U_mod_biog <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(AFDM_ugcm2) + (1|site), data=covariat_NU_G)
summary(U_mod_biog)
#hist(residuals(U_mod_bio),main = paste(NULL))
U_mod_bio_r2_valsg <- r.squaredGLMM(U_mod_bio)
U_mod_bio_conditional_r2g <- round(U_mod_bio_r2_vals[2], 2)

```

#### e) Uptake and chl-a:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_chla <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(Chla_ugL_Q) + (1|site), data=covariat_NU_B)

summary(U_mod_chla)
#hist(residuals(U_mod_chla),main = paste(NULL))
r.squaredGLMM(U_mod_chla)

U_mod_chla_r2_vals <- r.squaredGLMM(U_mod_chla)
U_mod_chla_conditional_r2 <- round(U_mod_chla_r2_vals[2], 2)
```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_chlag <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale((Chla_ugL_Q)) + (1|site), data=covariat_NU_G)

summary(U_mod_chlag)
#hist(residuals(U_mod_chlag),main = paste(NULL))
r.squaredGLMM(U_mod_chlag)

U_mod_chla_r2_vals <- r.squaredGLMM(U_mod_chlag)
U_mod_chla_conditional_r2 <- round(U_mod_chla_r2_vals[2], 2)
```

#### f) Uptake and GPP:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_GPP <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(log(GPP_mean+1)) + (1|site), data=covariat_NU_B)

summary(U_mod_GPP)
#hist(residuals(U_mod_GPP), main = paste(NULL))

U_mod_GPP_r2_vals <- r.squaredGLMM(U_mod_GPP)
U_mod_GPP_conditional_r2 <- round(U_mod_GPP_r2_vals[2], 2)
```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}

U_mod_GPPg <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(log(GPP_mean+1)) + (1|site), data=covariat_NU_G)

summary(U_mod_GPPg)
#hist(residuals(U_mod_GPP), main = paste(NULL))
U_mod_GPP_r2_vals <- r.squaredGLMM(U_mod_GPPg)
U_mod_GPP_conditional_r2 <- round(U_mod_GPP_r2_vals[2], 2)
```

#### g) Uptake and \|ER\|:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_ER <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale((ER_mean*-1)) + (1|site), data=covariat_NU_B)

summary(U_mod_ER)
#hist(residuals(U_mod_ER), main = paste(NULL))
U_mod_ER_r2_vals <- r.squaredGLMM(U_mod_ER)
U_mod_ER_conditional_r2 <- round(U_mod_ER_r2_vals[2], 2)
```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_ER <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale((ER_mean*-1)) + (1|site), data=covariat_NU_G)

summary(U_mod_ER)
#hist(residuals(U_mod_ER), main = paste(NULL))
U_mod_ER_r2_vals <- r.squaredGLMM(U_mod_ER)
U_mod_ER_conditional_r2 <- round(U_mod_ER_r2_vals[2], 2)

```

#### h) Uptake and Q:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_Q <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(log(Q_m+1)) + (1|site), data=covariat_NU_B)

summary(U_mod_Q)
#hist(residuals(U_mod_Q), main = paste(NULL))
U_mod_Q_r2_vals <- r.squaredGLMM(U_mod_Q)
U_mod_Q_conditional_r2 <- round(U_mod_Q_r2_vals[2], 2)
```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_Qg <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(log(Q_m)+1) + (1|site), data=covariat_NU_B)

summary(U_mod_Qg)
#hist(residuals(U_mod_Qg), main = paste(NULL))
U_mod_Q_r2_vals <- r.squaredGLMM(U_mod_Qg)
U_mod_Q_conditional_r2 <- round(U_mod_Q_r2_vals[2], 2)
```

#### i) Uptake and temp:

##### BW:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_temp <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(wtr_m) + (1|site), data=covariat_NU_B)

summary(U_mod_temp)
#hist(residuals(U_mod_temp), main = paste(NULL))
r.squaredGLMM(U_mod_temp)
```

##### GB:

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}

U_mod_tempg <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(wtr_m) + (1|site), data=covariat_NU_G)

summary(U_mod_tempg)
#hist(residuals(U_mod_temp), main = paste(NULL))
r.squaredGLMM(U_mod_tempg)
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_NO3 <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(NO3_mgL_dl) + (1|site), data=covariat_datq_BW)

summary(U_mod_NO3)
#hist(residuals(U_mod_NO3), main = paste(NULL))
r.squaredGLMM(U_mod_NO3)
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_NO3 <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(NO3_mgL_dl) + (1|site), data=covariat_datq_GB%>%filter(NO3_mgL_dl<0.7780))

summary(U_mod_NO3)
#hist(residuals(U_mod_NO3), main = paste(NULL))
r.squaredGLMM(U_mod_NO3)
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_NH4 <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(NH4_mgL_dl) + (1|site), data=covariat_datq_BW%>%filter(substrate=="pw"))

summary(U_mod_NH4)
#hist(residuals(U_mod_NH4), main = paste(NULL))
r.squaredGLMM(U_mod_NH4)
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
U_mod_NH4 <- lmer(log(Uadd_ug_L_min+1) ~ 
                      scale(NH4_mgL_dl) + (1|site), data=covariat_datq_GB%>%filter(substrate=="pw"))

summary(U_mod_NH4)
#hist(residuals(U_mod_NH4), main = paste(NULL))
r.squaredGLMM(U_mod_NH4)
```

#### Plots of Ut and all linear model covariates by stream

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
U_OM_plt<- ggplot(covariat_datq, aes(x=AFDM_mgg, y = Uadd_ug_L_min, color =catch, shape=method)) + 
  geom_point(size=2, alpha=0.75) + theme_classic() + 
   ylab(expression(U[t]~(µg~L^-1~min^-1)))+
  xlab(expression(OM~(mg~g^-1)))+
  scale_color_manual(values = catch_colors)

U_biomass_plt <- ggplot(covariat_NU, aes(x = AFDM_ugcm2, y = log(Uadd_ug_L_min + 1))) + 
  geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() +
 # geom_smooth(method = "lm", aes(color = catch), se = FALSE, alpha=0.5, lty="dashed") + 
  #ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  ylab(" ") +
  xlab(expression(Biomass~(µg~cm^2))) +
  scale_color_manual(values = catch_colors)

U_Chla_plt <- ggplot(covariat_NU, aes(x = log(Chla_ugL_Q+1), y = log(Uadd_ug_L_min + 1))) + 
  geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
  geom_smooth(data = subset(covariat_NU), 
              aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) + 
  ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  xlab(expression(log(Chla+1)~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)


U_GPP_plt<-ggplot(covariat_NU, aes(x=log(GPP_mean+1),  y = log(Uadd_ug_L_min+1))) + 
    geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
    geom_smooth(data = subset(covariat_NU, catch == "GB"), 
              aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) + 
    ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
    xlab(expression(log(GPP+1)~(g~O[2]~m^-2~d^-1))) +
  scale_color_manual(values = catch_colors)

U_ER_plt<-ggplot(covariat_NU, aes(x=(ER_mean*-1),  y = log(Uadd_ug_L_min+1))) + 
    geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
      geom_smooth(data = subset(covariat_NU, catch == "GB"), 
              aes(color = catch), method = "lm", se = FALSE, alpha = 0.5, lty="dashed") +
        geom_smooth(data = subset(covariat_NU, catch == "BW"), 
              aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) +
  geom_smooth(method = "lm", aes(color = catch), se = FALSE, alpha=0.5) + 
   #  ylab(expression(log(U[t]+1)~(µg~L^-1~min^-1)))+
  ylab(" ") +
          xlab(expression(abs(ER)~(g~O[2]~m^-2~d^-1))) +
  scale_color_manual(values = catch_colors)

U_temp_plt<-ggplot(covariat_NU, aes(x=wtr_m,  y = log(Uadd_ug_L_min+1), color =catch, shape=method)) + 
  geom_point(size=2, alpha=0.75) + theme_classic() + 
    ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  #ylab(" ") +
        xlab(expression(Temp~(degree*C))) +
  scale_color_manual(values = catch_colors)

U_Q_plt<-ggplot(covariat_NU, aes(x=log(Q_m+1), y = log(Uadd_ug_L_min+1))) + 
      geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
  geom_smooth( aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) +
    #ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  ylab(" ") +
     xlab(expression(log(Q+1)~(m^3~s^-1))) +
  scale_color_manual(values = catch_colors)



U_NO3_plt<-ggplot(covariat_datq%>%filter(NO3_mgL_dl < 0.7780, !is.na(method), substrate=="pw"), 
                  aes(x=NO3_mgL_dl*1000, y = log(Uadd_ug_L_min+1))) + 
      geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
    xlim(0,290) +
 # geom_smooth(data = subset(covariat_datq, catch == "GB"), 
  #            aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) +
    ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  #ylab(" ") +
  xlab(expression(Ambient~NO[3]^-1~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)

U_NH4_plt<-ggplot(covariat_datq%>%filter(!is.na(method) & substrate=="pw", NH4_mgL_dl < 0.15), aes(x=NH4_mgL_dl*1000, y = log(Uadd_ug_L_min+1))) + 
      geom_point(aes(color = catch, shape = method), size = 2, alpha = 0.75) + 
  theme_classic() + 
 # xlim(0,300) +
  geom_smooth(aes(color = catch), method = "lm", se = FALSE, alpha = 0.5) +
    #ylab(expression(log(U[t] + 1) ~ (µg~L^-1~min^-1))) +
  ylab(" ") +
  xlab(expression(Ambient~NH[4]^+1~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=5}

Nup_grid2 <- ggarrange(
  U_biomass_plt,
  U_Chla_plt,
  U_GPP_plt,
  U_ER_plt,
  U_temp_plt,
  U_Q_plt,
  ncol = 3, nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

```

### Figure 5: General drivers of either NO3 or NH4 uptake rates (Ut)

### (Q2) - Is variance in nitrogen demand and instream uptake (either as ammonium or nitrate) more related to biological (GPP, ER, biomass or chlorophyll-a), physical processes (flow and water temperature), or ambient nutrient and organic matter availability? 

Point color represents site, point shape by N uptake method triangles
are no3 and circles are nh4

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=8, fig.height=7}
#Nup_grid3 <- ggarrange(
  #Nup_grid,
  # nup_sum_plot, # a
  # nvf_sum_plot, # b
  # nsw_sum_plot, # c
#   U_biomass_plt, # d
#   U_GPP_plt, # g
#   U_temp_plt, # i
#   U_NO3_plt, # k
#   U_Chla_plt, # f
#   U_ER_plt, # h
#   U_Q_plt, # j
#   U_NH4_plt, # l
#   labels = c(#"a", "b", "c",
#              "d", "f", "h", "j", 
#              "e", "g", "i", "k"),
#   ncol = 3, nrow = 3,
#   common.legend = TRUE,
#   legend = "right"
# )
# 
# Nup_grid3
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4.5, fig.height=12}
Nup_grid3 <- ggarrange(
  U_Chla_plt,
  U_biomass_plt,
  U_GPP_plt,
  U_ER_plt,
  U_temp_plt, # i
  U_Q_plt, # j
  U_NO3_plt, # k
  U_NH4_plt, # l
  labels = c("a", "b", "c", "d", 
             "e", "f", "g", "h"),
  ncol = 2, nrow = 4,
  common.legend = TRUE,
  legend = "right",
  label.x = 0.85,  
  label.y = 0.95   
)

Nup_grid3
```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4.75, fig.height=7}
 
# # Create the top row with extra spacing on both sides
# #top_row <- ggdraw() + draw_plot(Nup_grid, x = 0.05, width = 0.95)  # Centered with 25% narrower width
# 
# # Arrange the plots
# Nup_grid4 <- plot_grid(
#   Nup_grid, Nup_grid3,
#   ncol = 1, rel_heights = c(1.1, 3.90)  # Keeps the bottom row proportionally larger
# )
# 
# # Display the plot
# print(Nup_grid4)
# # 
# # 
# # 
# # 
# # Nup_grid4 <- ggarrange(
# #   Nup_grid,
# #   Nup_grid3,
# #   ncol = 1, nrow = 2,
# #   common.legend = TRUE,
# #   legend = "right"
# #   ##
# # )
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Nuptake_supp_CH1_grid.png", plot = Nup_grid3, width = 5.15, height = 8.5, units = "in")

```

### Metabolism by water year:

###### Dataframes:

For metabolism by water year March - October (2021-2024)

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
metab_dat<- water_year(metab_dat)

metab_plots_WY <- metab_dat %>%
  mutate(mon= month(date))%>%
  filter(mon>2 & mon<11) %>%
  filter(water_year<2025)%>%
  group_by(site, water_year) %>%
  summarise(GPP = mean(GPP_mean, na.rm=T),
            GPP_sd = sd(GPP_mean, na.rm=T),
            ER = mean(ER_mean, na.rm=T),
            ER_sd = sd(ER_mean, na.rm=T),
            GPP_se = sd(GPP_mean, na.rm = TRUE) / sqrt(length(GPP_mean)),
            ER_se = sd(ER_mean, na.rm = TRUE) / sqrt(length(ER_mean))
            ) %>%
  ungroup() %>%  # Remove grouping to calculate overall means
    mutate(
        GPP_se = if_else(is.na(GPP_se), mean(GPP_se, na.rm = TRUE), GPP_se), 
        ER_se = if_else(is.na(ER_se), mean(ER_se, na.rm = TRUE), ER_se)
        )
  
```

Metabolimsm regression analysis :

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}

metab_WY_lmdf <- metab_dat %>%
  mutate(mon= month(date))%>%
  filter(mon>2 & mon<11) %>%
  filter(water_year<2025)%>%
  mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",
    TRUE ~ NA_character_  
  ))
```

For nitrogen dynamics by water year March - October (2021-2023)

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
nuts_plots_WY <- covariat_nitrogen2 %>%
  mutate(mon= month(date))%>%
  filter(mon>3 & mon<10) %>%
    filter(NO3_supply<200) %>%
  filter(water_year<2024) %>%
  group_by(site, water_year) %>%
  summarise(
            NO3_supplym = mean(NO3_supply, na.rm=T),
            NO3_supply_sd = sd(NO3_supply, na.rm=T),
            NH4_supplym = mean(NH4_supply, na.rm=T),
            NH4_supply_sd = sd(NH4_supply, na.rm=T),
            PO4_supplym = mean(PO4_supply, na.rm=T),
            PO4_supply_sd = sd(PO4_supply, na.rm=T),
            Ndemandm = mean(Ndemand, na.rm=T),
            Ndemand_sd = sd(Ndemand, na.rm=T),
            NH4_supply_se = sd(NH4_supply, na.rm = TRUE) / sqrt(length(NH4_supply)),
            NO3_supply_se = sd(NO3_supply, na.rm = TRUE) / sqrt(length(NO3_supply)),
            Ndemand_se = sd(Ndemand, na.rm = TRUE) / sqrt(length(Ndemand))
            ) %>%
  ungroup() %>%  # Remove grouping to calculate overall means
    mutate(
        NH4_supply_se = if_else(is.na(NH4_supply_se), mean(NH4_supply_se, na.rm = TRUE), NH4_supply_se), 
        NO3_supply_se = if_else(is.na(NO3_supply_se), mean(NO3_supply_se, na.rm = TRUE), NO3_supply_se),
        Ndemand_se = if_else(is.na(Ndemand_se), mean(Ndemand_se, na.rm = TRUE), Ndemand_se)
        )
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
covariat_WY <- covariat_nitrogen2 %>% # was covariat_datq
    mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",
    TRUE ~ NA_character_  
  )) %>%
  mutate(mon= month(date))%>%
  filter(mon>2 & mon<11) %>%
  filter(water_year<2025) 

covariat_WY<- covariat_WY%>%
  dplyr::select(site, catch, date, water_year, Chla_ugL_Q, method, Uadd_ug_L_min, 
         Ndemand, NO3_supply, NH4_supply, PO4_supply, WY_lab, GPP_mean, ER_mean,
         AFDM_ugcm2, Q_m, wtr_m) 

covariat_WY <- covariat_WY %>% distinct()

covariat_WY_sum <- covariat_WY %>%
  group_by(catch, WY_lab) %>%
  summarise(
    Ndemandtm = mean(Ndemand, na.rm=T),
    NH4_supplym = mean(NH4_supply, na.rm=T), 
    NO3_supplym = mean(NO3_supply, na.rm=T)
    )

metab_WY_sum <- metab_WY_lmdf %>%
  group_by(catch, WY_lab) %>%
  summarise(
    GPP_wym = mean(GPP_mean, na.rm=T),
    ER_wym = mean(ER_mean, na.rm=T)
    )
# pull out streams 
covariat_WY_B <- covariat_WY %>%
  filter(catch=="BW") 

covariat_WY_G <- covariat_WY %>%
  filter(catch=="GB") 

```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
(3.23-0.79)/3.23

(13.680257-10.864423)/13.680257

(0.20-0.08)/0.20

(6.211836-3.511)/6.211836


```

##### Linear models for metabolism across dry or wet years

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3}
lmm_WY_GPP <- lmer(GPP_mean ~ WY_lab + (1|site) + (1|catch), data = metab_WY_lmdf)
summary(lmm_WY_GPP)
#hist(residuals(lmm_WY_GPP), main = paste(NULL))
r.squaredGLMM(lmm_WY_GPP)

lmm_WY_GPP_BW <- lmer(GPP_mean ~ WY_lab + (1|site), data = metab_WY_lmdf%>%filter(catch=="BW"))
summary(lmm_WY_GPP_BW)
#hist(residuals(lmm_WY_GPP_BW), main = paste(NULL))
r.squaredGLMM(lmm_WY_GPP_BW)

lmm_WY_GPP_GB <- lmer(GPP_mean ~ WY_lab + (1|site), data = metab_WY_lmdf%>%filter(catch=="GB"))
summary(lmm_WY_GPP_GB)
#hist(residuals(lmm_WY_GPP_GB), main = paste(NULL))
r.squaredGLMM(lmm_WY_GPP_GB)

lmm_WY_ER <- lmer((ER_mean*-1) ~ WY_lab + (1|site) + (1|catch), data = metab_WY_lmdf)
summary(lmm_WY_ER)
#hist(residuals(lmm_WY_ER), main = paste(NULL))
r.squaredGLMM(lmm_WY_ER)

lmm_WY_ER_BW <- lmer((ER_mean*-1) ~ WY_lab + (1|site), data = metab_WY_lmdf%>%filter(catch=="BW"))
summary(lmm_WY_ER_BW)
#hist(residuals(lmm_WY_ER_BW), main = paste(NULL))
r.squaredGLMM(lmm_WY_ER_BW)

lmm_WY_ER_GB <- lmer((ER_mean*-1) ~ WY_lab + (1|site), data =  metab_WY_lmdf%>%filter(catch=="GB"))
summary(lmm_WY_ER_GB)
#hist(residuals(lmm_WY_ER_GB), main = paste(NULL))
r.squaredGLMM(lmm_WY_ER_GB)
```

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=3, fig.height=3.5}

lmm_WY_NO3_Q <- lmer(NO3_supply ~ WY_lab + (1|site) + (1|catch),, data = covariat_WY)
summary(lmm_WY_NO3_Q)
#hist(residuals(lmm_WY_NO3_Q), main = paste(NULL))
r.squaredGLMM(lmm_WY_NO3_Q)


lmm_WY_NH4_Q <- lmer(NH4_supply ~ WY_lab + (1|site) + (1|catch), data = covariat_WY)
summary(lmm_WY_NH4_Q)
#hist(residuals(lmm_WY_NH4_Q), main = paste(NULL))
r.squaredGLMM(lmm_WY_NH4_Q)

lmm_WY_demand_Q <- lmer(Ndemand ~ WY_lab + (1|site) + (1|catch), data = covariat_WY)
summary(lmm_WY_demand_Q)
#hist(residuals(lmm_WY_demand_Q),main = paste(NULL))
r.squaredGLMM(lmm_WY_demand_Q)

```

##### plots

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
GPP_plots <- ggplot(metab_plots_WY%>%filter(water_year<2024), aes(x = water_year, y = GPP, color = site, shape=site)) +
  geom_point(size = 4, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin =GPP-GPP_se, ymax = GPP+GPP_se),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() + ylab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
  xlab(NULL) + scale_color_manual(values = siteC_colors) +
  theme(legend.position =  "none") +
    scale_shape_manual(values = c(15,0,17,2)) +
    scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))

GPP_plots  

ER_plots <- ggplot(metab_plots_WY%>%filter(water_year<2024), aes(x = water_year, y = ER, color = site, shape=site)) +
  geom_point(size = 4, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin =ER-ER_se, ymax = ER+ER_se),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() + ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  xlab(NULL) + scale_color_manual(values = siteC_colors) +
    theme(legend.position =  "none") +
      scale_shape_manual(values = c(15,0,17,2)) +
    scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))

ER_plots  

```

```{r, warning=F, size = 'small', echo=F, include=FALSE,fig.width=6, fig.height=4}
# 
NO3_plots <- ggplot(nuts_plots_WY, aes(x = (water_year), y = NO3_supplym, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin=NO3_supplym-(NO3_supply_se), ymax = NO3_supplym+(NO3_supply_se)),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() + ylab(expression(NO[3]~supply~(g~N~m^-2~d^-1))) +
  #ylim(0,50)+
  xlab(NULL) + scale_color_manual(values = siteC_colors) +
      scale_shape_manual(values = c(15,0,17,2)) +
      scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))


NH4_plots <- ggplot(nuts_plots_WY, aes(x = (water_year), y = NH4_supplym, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.1)) +
  geom_line(position = position_dodge(width = 0.1))+
  geom_errorbar(aes(ymin=NH4_supplym-NH4_supply_se, ymax = NH4_supplym+NH4_supply_se),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.1)) +
   # ylim(0,50)+
  theme_classic() + ylab(expression(NH[4]~supply~(g~N~m^-2~d^-1))) + # mg N m-2 day -1
  xlab(NULL) + scale_color_manual(values = siteC_colors) +
        scale_shape_manual(values = c(15,0,17,2)) +
        scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))
NH4_plots  

N_demand_plots <- ggplot(nuts_plots_WY, aes(x = (water_year), y = Ndemandm, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3)) +
  geom_errorbar(aes(ymin =(Ndemandm-Ndemand_se), ymax = (Ndemandm+Ndemand_se)),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() +   
  ylab(expression(Nitrogen~demand~(g~N~m^-2~d^-1))) +
 #ylim(0,300)+
          scale_shape_manual(values = c(15,0,17,2)) +
  xlab(NULL) + scale_color_manual(values = siteC_colors)  +
        scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))
N_demand_plots

```

### Figure 6: Average annual differences in GPP, ER, N-demand and N-supply by water year

Values averaged across March to October observations

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=8, fig.height=6}

WY_n_grid <- ggarrange(
  N_demand_plots,
  NO3_plots,
  NH4_plots,
  labels = c("c","d", "e"),
  ncol = 3, nrow = 1,
  common.legend = TRUE, 
  legend = "bottom")


### big grid 
WY_n_grid2 <- ggarrange(
  GPP_plots,
  ER_plots,
  ncol = 2, nrow = 1,
    labels = c("a","b"))


### big grid 
WY_n_grid3 <- ggarrange(
  WY_n_grid2,
  WY_n_grid,
  ncol = 1, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom")

WY_n_grid3

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/WY_n_grid_CH1.png", plot = WY_n_grid3, width = 8, height = 6.25, units = "in")

```

### Extra?

Relationship between Ut and n-demand? not really there.

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
names(covariat_nitrogen2)

plot(covariat_nitrogen2$Uadd_ug_L_min,covariat_nitrogen2$Ndemand)
cor.test(covariat_nitrogen2$Uadd_ug_L_min,covariat_nitrogen2$Ndemand)


U_temp_plt<-ggplot(covariat_nitrogen2, aes(x=Ndemand,  y = log(Uadd_ug_L_min+1), color =catch, shape=method)) + 
  geom_point(size=2, alpha=0.75) + 
  theme_classic() + 
  geom_smooth(method="lm")


covariat_WY <- covariat_datq %>%
    mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",

    TRUE ~ NA_character_  
  )) %>%
  filter(Uadd_ug_L_min<120)


lmm_WY_Ut <- lmer(Uadd_ug_L_min ~ WY_lab + (1|site) + (1|catch), data = covariat_WY)
summary(lmm_WY_Ut)
#hist(residuals(lmm_WY_ER))
r.squaredGLMM(lmm_WY_Ut)

lmm_WY_Ut <- lmer(Uadd_ug_L_min ~ WY_lab + method + (1|site) + (1|catch), data = covariat_WY)
summary(lmm_WY_Ut)
#hist(residuals(lmm_WY_ER))
r.squaredGLMM(lmm_WY_Ut)

lmm_WY_Ut <- lmer(Uadd_ug_L_min ~ WY_lab + (1|site) + (1|catch), data = covariat_WY%>%filter(method=="NO3"))
summary(lmm_WY_Ut)
#hist(residuals(lmm_WY_ER))
r.squaredGLMM(lmm_WY_Ut)


lmm_WY_Ut <- lmer(Uadd_ug_L_min ~ WY_lab + (1|site) + (1|catch), data = covariat_WY%>%filter(method=="NH3"))
summary(lmm_WY_Ut)
#hist(residuals(lmm_WY_ER))
r.squaredGLMM(lmm_WY_Ut)



lmm_WY_sw <- lmer(sw_m ~ WY_lab + method + (1|site) + (1|catch), data = covariat_WY)
summary(lmm_WY_sw)
#hist(residuals(lmm_WY_sw))
r.squaredGLMM(lmm_WY_sw)

uptake_yr <- covariat_WY %>%
  group_by(site, method) %>%
  dplyr::summarise(
    Uadd_ug_L_min = mean(Uadd_ug_L_min, na.rm=T),
    sw_m = mean(sw_m, na.rm=T),
    vf = mean(vf, na.rm=T))
  

uptake_yr <- covariat_WY %>%
  group_by(method) %>%
  dplyr::summarise(
    Uadd_ug_L_min = mean(Uadd_ug_L_min, na.rm=T),
    sw_m = mean(sw_m, na.rm=T),
    vf = mean(vf, na.rm=T))

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
(17.93211-10.89327)/17.93211

(0.001728073-0.001147727)/0.001728073

(34.78636-23.26239)/34.78636
```

### Correlations btwn GPP and other standing stock estiamtes of productivity

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

mod1 <- lmer(log(GPP_mean+1) ~ scale(AFDM_ugcm2) +  (1|site) + (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)


mod1 <- lmer(log(GPP_mean+1) ~ scale(Chla_ugL_Q) +  (1|site) + (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)

mod1 <- lmer(log(Chla_ugL_Q+1) ~ scale(AFDM_ugcm2) +  (1|site) + (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)

mod1 <- lmer(log((ER_mean*-1)+1) ~ scale(AFDM_ugcm2) + (1|site) + (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)


mod1 <- lmer(log((ER_mean*-1)+1)  ~ scale(AFDM_mgg) + (1|site)+ (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)

mod1 <- lmer(log((ER_mean*-1)+1)  ~ scale(Chla_ugL_Q) +  (1|site) + (1|catch), data = covariat_datq)
summary(mod1)
#hist(residuals(mod1))
r.squaredGLMM(mod1)
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

BM_cor_plt <- ggplot(covariat_datq, aes(x = log(AFDM_ugcm2+1), y = log(GPP_mean+1))) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
  theme_classic() + 
      geom_smooth(data = covariat_datq%>%filter(catch=="BW") , aes(color = "grey50"), method = "lm", se = FALSE, alpha = 0.5) + 
   ylab(expression(log(GPP+1)~(g~O[2]~m^-2~d^-1))) +
  xlab(expression(log(Biomass+1)~(µg~cm^-2))) +
  scale_color_manual(values = catch_colors)

BM_cor_plt1 <- ggplot(covariat_datq, aes(x = log(Chla_ugL_Q+1), y = log(GPP_mean+1))) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
  theme_classic() + 
      geom_smooth(aes(color = "grey50"), method = "lm", se = FALSE, alpha = 0.5) + 
  ylab(expression(log(GPP+1)~(g~O[2]~m^-2~d^-1))) +
  xlab(expression(log(Chla+1)~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)

BM_cor_plt2 <- ggplot(covariat_datq, aes(y = log(Chla_ugL_Q+1), x = log(AFDM_ugcm2+1))) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
      geom_smooth(aes(color = "grey50"), method = "lm", se = FALSE, alpha = 0.5) + 
  theme_classic() + 
  xlab(expression(log(Biomass+1)~(µg~cm^-2))) +
  ylab(expression(log(Chla+1)~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)


BM_cor_plt3 <- ggplot(covariat_datq, aes(x = log(AFDM_ugcm2+1), y = log((ER_mean*-1)+1))) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
  theme_classic() + 
  ylab(expression(log(abs(ER)+1)~(g~O[2]~m^-2~d^-1))) +
  xlab(expression(log(Biomass+1)~(µg~cm^-2))) +
  scale_color_manual(values = catch_colors)


BM_cor_plt4 <- ggplot(covariat_datq, aes(x = log(Chla_ugL_Q+1), y = log((ER_mean*-1))+1)) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
  theme_classic() + 
   # geom_smooth(aes(color = catch), method = "lm", se = FALSE, alpha = 0.5, lty="dashed") + 
    ylab(expression(log(abs(ER)+1)~(g~O[2]~m^-2~d^-1))) +
  xlab(expression(log(Chla+1)~(µg~L^-1))) +
  scale_color_manual(values = catch_colors)


BM_cor_plt5 <- ggplot(covariat_datq, aes(x = log(AFDM_mgg+1), y = log((ER_mean*-1))+1)) + 
  geom_point(aes(color = catch), size = 2, alpha = 0.55) + 
  theme_classic() + 
    ylab(expression(log(abs(ER)+1)~(g~O[2]~m^-2~d^-1))) +
  xlab(expression(log(OM+1)~(mg~g^-1))) +
  scale_color_manual(values = catch_colors)

```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=8, fig.height=7}
BP_grid <- ggarrange(
  BM_cor_plt,
  BM_cor_plt1,
  BM_cor_plt2,
  BM_cor_plt3,
  BM_cor_plt4,
  BM_cor_plt5,
  labels = c("a", "b", "c",
             "d", "e", "f"),
  ncol = 3, nrow = 2,
  common.legend = TRUE,
  legend = "right"
)

BP_grid
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/BP_grid_CH1.png", plot = BP_grid, width = 8.5, height = 4.75, units = "in")

```

### Looking at stoichiometeric ratios of N:P overtime?

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

#summary(covariat_NS_ND)
#
# Stoich_df <- covariat_nitrogen2 %>%
#   mutate(
#     NO3_ugL = c(NO3_mgL_dl*1000),
#     NH4_ugL = c(NH4_mgL_dl*1000),
#     Tot_N_ugL = c(NO3_ugL+ NH4_ugL),
#     N_P_ugL = (Tot_N_ugL/PO4_ugL_dl))


Stoich_df <- covariat_nitrogen2 %>%
  mutate(
    jday= yday(date),
    NO3_uM = (NO3_mgL_dl * 1000) / 62.0049,  # Convert NO3 to µmol/L
    NH4_uM = (NH4_mgL_dl * 1000) / 18.0385,  # Convert NH4 to µmol/L
    PO4_uM = PO4_ugL_dl / 30.9738,  # Convert PO4 to µmol/L
    Tot_N_uM = NO3_uM + NH4_uM,  # Total nitrogen in µmol/L
    N_P_ratio = if_else(PO4_uM > 0, Tot_N_uM / PO4_uM, NA_real_)  # Avoid division by zero
  )
```

### N:P

Looks like N:P \<16 across all observations

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=9, fig.height=4}

N_P_ratio_plots <- ggplot(Stoich_df %>%filter(water_year<2024), aes(x = jday, y = N_P_ratio, color = site)) +
  geom_point(size = 2, alpha = 0.75, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = site_colors) +
   ylab(expression(N:P~(umol~L^-1))) +
     xlab("Day of year") +
  facet_grid(.~water_year) +
  theme_classic()

N_P_ratio_plots
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=5, fig.height=4}
N_P_ratio_plots <- ggplot(Stoich_df %>%filter(water_year<2024), aes(y = GPP_mean, x = N_P_ratio, color = site)) +
  geom_point(size = 2, alpha = 0.75, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = site_colors) +
   xlab(expression(N:P~(umol~L^-1))) +
  #facet_grid(.~water_year) +
  theme_classic()

N_P_ratio_plots
```

#### Statisical relationship?

Maybe? GPP and N:P are slightly negatively related. But falls apart when
both catchments are modeled independently

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=5, fig.height=4}

mod <- lmer((GPP_mean) ~ scale(N_P_ratio) + (1|catch), data = Stoich_df %>%filter(water_year<2024))
summary(mod)
#hist(residuals(mod), main = paste(NULL))
r.squaredGLMM(mod)

modb <- lmer((GPP_mean) ~ scale(N_P_ratio) + (1|site), data = Stoich_df %>%filter(water_year<2024 & catch=="BW"))
summary(modb)
#hist(residuals(modb), main = paste(NULL))
r.squaredGLMM(modb)


modg <- lmer((GPP_mean) ~ scale(N_P_ratio) + (1|site), data = Stoich_df %>%filter(water_year<2024 & catch=="GB"))
summary(modg)
#hist(residuals(modg), main = paste(NULL))
r.squaredGLMM(modg)
```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=5, fig.height=4}

N_P_ratio_plots <- ggplot(Stoich_df %>%filter(water_year<2024), aes(y = ER_mean, x = N_P_ratio, color = site)) +
  geom_point(size = 2, alpha = 0.75, position = position_dodge(width = 0.3)) +
  scale_color_manual(values = site_colors) +
   xlab(expression(N:P~(umol~L^-1))) +
  #facet_grid(.~water_year) +
  theme_classic()

N_P_ratio_plots
```

#### Statisical relationship?

No

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=5, fig.height=4}

mod1 <- lmer((ER_mean*-1) ~ scale(N_P_ratio) + (1|catch), data = Stoich_df %>%filter(water_year<2024))
summary(mod1)
#hist(residuals(mod1), main = paste(NULL))
r.squaredGLMM(mod1)
```

#### Table of mean N:P ratios by site

```{r, warning=F, size = 'small', echo=F, fig.width=5, fig.height=12}

Stoich_table <- Stoich_df %>%
  group_by(site)  %>%
  summarise(
    Tot_N_uMm = round(mean(Tot_N_uM, na.rm=T),3),
    N_P_ratiom = round(mean(N_P_ratio, na.rm=T),4),
    N_P_ratiomin = round(min(N_P_ratio, na.rm=T),4),
    N_P_ratiomax = round(max(N_P_ratio, na.rm=T),3),
    rangeNP = c(N_P_ratiomax-N_P_ratiomin)
)

Stoich_table
```

```{r, warning=F, size = 'small', echo=F, fig.width=5, include=FALSE,fig.height=4}

Stoich_plot <- Stoich_df %>%
  filter(water_year<2024) %>%
  group_by(site, water_year) %>%
  dplyr::summarise(
    Tot_N_uMm = round(mean(Tot_N_uM, na.rm = TRUE), 3),
    N_P_ratiom = round(mean(N_P_ratio, na.rm = TRUE), 4),
    N_P_sd = round(sd(N_P_ratio, na.rm = TRUE), 4),
    N_P_se = round(sd(N_P_ratio, na.rm = TRUE) / sqrt(sum(!is.na(N_P_ratio))), 4)  # Standard Error
  ) %>%
  mutate(
    N_P_sd = if_else(is.na(N_P_sd), N_P_ratiom, N_P_sd),  # Replace NA SD with mean
    N_P_se = if_else(is.na(N_P_se), 0, N_P_se)  # Replace NA SE with 0
  )


N_P_plots <- ggplot(Stoich_plot, aes(x = (water_year), y = N_P_ratiom, color = site, shape=site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_line(position = position_dodge(width = 0.3))+
  geom_errorbar(aes(ymin =N_P_ratiom-N_P_se, ymax = N_P_ratiom+N_P_se),  width = 0.05, height = 0, alpha = 0.75, position = position_dodge(width = 0.3)) +
  theme_classic() + ylab(expression(N:P~(µM~L^-1))) +
 #ylim(0,300)+
 scale_shape_manual(values = c(15,0,17,2)) +
  xlab(NULL) + scale_color_manual(values = siteC_colors)  +
        scale_x_continuous(breaks = seq(floor(min(nuts_plots_WY$water_year)), ceiling(max(nuts_plots_WY$water_year)), by = 1), labels = function(x) as.character(round(x)))
N_P_plots

```

## Moddified figure 6?

#### (3) How does in-stream productivity and nitrogen supply and demand dynamics change within stream catchments and across dry (2021 and 2022) and wet water years (2023)?


```{r, warning=F, size = 'small', echo=F, fig.width=7, fig.height=5}

WY_n_grid <- ggarrange(
  N_demand_plots,
  N_P_plots,
  NO3_plots,
  NH4_plots,
  labels = c("c","d","e", "f"),
  ncol = 4, nrow = 1,
  common.legend = TRUE, 
  legend = "bottom")


### big grid 
WY_n_grid2 <- ggarrange(
  GPP_plots,
  ER_plots,
  ncol = 2, nrow = 1,
    labels = c("a","b"))


# ### big grid 
# WY_n_grid3 <- ggarrange(
#   WY_n_grid2,
#   WY_n_grid,
#   ncol = 1, nrow = 2,
#   common.legend = TRUE, 
#   legend = "bottom")
# 
# WY_n_grid3



library(ggpubr)
library(cowplot)

# Create the top row with extra spacing on both sides
top_row <- ggdraw() + draw_plot(WY_n_grid2, x = 0.05, width = 0.95)  # Centered with 25% narrower width

# Arrange the plots
WY_n_grid3 <- plot_grid(
  top_row, WY_n_grid,
  ncol = 1, rel_heights = c(4, 5)  # Keeps the bottom row proportionally larger
)

# Display the plot
print(WY_n_grid3)


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/WY_n_grid_CH1_v2.png", plot = WY_n_grid3, width = 8.5, height = 5.25, units = "in")

```


