---
title: 'Antecedent flow conditions for MS'
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
#setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWL_k600/")
lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB", "plotly", "MASS",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
```

### MS: Spatiotemporal variation in mountain stream metabolism and nitrogen cycling across contrasting flow regimes

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
daoc_date<- Sys.Date()

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
site_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#97b5c2",
  "GBL" = "#DD6E42",
  "GBU" = "#d9cb7c"
)

siteC_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#054fb9",
  "GBL" = "#DD6E42",
  "GBU" = "#DD6E42"
)



catch_colors <- c(
  "BW" = "#054fb9",
  "GB" = "#DD6E42"
)

# Define fill patterns for water years
water_year_patterns <- c("wave","weave","stripe","circle", "stripe", "circle")


# Create a sequence of dates
date_seq <- seq(from = as.Date("2021-03-20"), to = as.Date("2024-10-10"), by = "day")

# Convert the sequence to a dataframe
date_df <- data.frame(date = date_seq)
date_df$site <- "BWL"
date_df$catch <- "BW"
date_df1 <- data.frame(date = date_seq)
date_df1$site <- "GBL"
date_df1$catch <- "GB"
date_df2 <- data.frame(date = date_seq)
date_df2$site <- "GBU"
date_df2$catch <- "GB"
date_df3 <- data.frame(date = date_seq)
date_df3$site <- "BWU"
date_df3$catch <- "BW"

date_datf <- rbind(date_df,date_df1,date_df2, date_df3)

### fxns:
## Fxn for water year:
water_year <- function(data) {
  # Ensure the `date` column exists
  if (!"date" %in% names(data)) {
    stop("The dataframe must contain a column named 'date'.")
  }
  
  # Convert the `date` column to Date format (if not already)
  data$date <- as.Date(data$date)
  
  # Add the water_year column
  data$water_year <- with(data, {
    year <- as.numeric(format(date, "%Y"))
    month <- as.numeric(format(date, "%m"))
    ifelse(month >= 10, year + 1, year)
  })
  
  return(data)
}
## Function to perform ANOVA and post-hoc test within each site
comparisons <- list(
  c("2021", "2022"),
  c("2021", "2023"),
  c("2022", "2023"),
  c("2023", "2024")
)
```

```{r, warning=F, size = 'small', echo=FALSE, include=F}
date_datf <- water_year(date_datf)

covariat_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_covariate_dat_AFDM.rds") %>%
  dplyr::select("date", "Site","bulk.density","AFDM_mgg","AFDM_mgcm2", "Chla_ugL_Q", "Pheo_ugL_Q")
summary(covariat_dat)

covariat_dat <- date_datf %>%
  left_join(covariat_dat, by=c("date", "site"="Site"))

flow_df <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_flow_df.rds")

covariat_dat <- covariat_dat %>%
    left_join(flow_df, by=c("date", "site"))


bg_nuts <- readRDS("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/NS_chem_dat_nh4_24.rds") %>%
  filter(location=="stream") %>%
    mutate(NO3_mgL_dl = if_else(NO3_mgL_dl < 0.0015, 0.0015, NO3_mgL_dl))%>%
    mutate(NH4_mgL_dl=if_else(NH4_mgL_dl < 0.001, 0.001, NH4_mgL_dl))%>%
    mutate(PO4_ugL_dl=if_else(PO4_ugL_dl < 0.201, 0.201, PO4_ugL_dl)) %>%
  mutate(PO4_ugL_dl=if_else(DOC_mgL_dl < 0.25, 0.125, DOC_mgL_dl)) %>%
  group_by(site, shore, date, depth, location, substrate) %>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")


##### pivot wider 
bg_nuts_wide <- bg_nuts %>%
  dplyr::select(site, date, substrate, NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill) %>%
  group_by(site, date,substrate) %>% 
  summarise(across(c(NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill), mean, na.rm = TRUE), .groups = "drop") %>% 
  pivot_wider(names_from = substrate, values_from = c(NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill), names_sep = "_")

######


WQ_dat <- read.csv("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/Stream_Lake_YSI_WaterQuality.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) %>%
  dplyr::select("site", "date","pH") %>%
  dplyr:: group_by(site, date) %>%
  dplyr::summarise(pH=mean(pH, na.rm=T))

covariat_datq <- covariat_dat%>%
  left_join(bg_nuts_wide[ , c("site", "date", "NO3_mgL_dl_sw", "NO3_mgL_dl_pw", "NH4_mgL_dl_sw", 
                              "NH4_mgL_dl_pw", "PO4_ugL_dl_sw","PO4_ugL_dl_pw", "DOC_mgL_dl_sw",
                              "DOC_mgL_dl_pw", "pH_infill_sw",  "pH_infill_pw")], by=c("date", "site"))

covariat_datq <- covariat_datq%>%
  left_join(WQ_dat,  by=c("date", "site"))


## bring in SPC data 
### missing 2024 BW SPC
SPC_BWL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWL_SPCv2.rds") %>%
  filter(datetime>as.POSIXct("2021-04-29 24:00:00"))
SPC_BWL$site <- "BWL"
SPC_BWU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWU_SPCv2.rds")
SPC_BWU$site <- "BWU"
SPC_GBL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBL_SPCv2.rds")
SPC_GBL$site <- "GBL"
SPC_GBU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBU_SPCv2.rds")
SPC_GBU$site <- "GBU"

SPC_dat <- rbind(SPC_BWL, SPC_BWU, SPC_GBL, SPC_GBU)
SPC_dat <- SPC_dat %>%
  mutate(date =as.Date(datetime)) 

SPC_dat_day <- SPC_dat %>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    wt= mean(wtr, na.rm=T),
    wt_sd= sd(wtr, na.rm=T),
    SPC_m=mean(SPC, na.rm=T),
    SPC_sd=sd(SPC, na.rm=T))

#hist(SPC_dat_day$wt_sd)
#hist(SPC_dat_day$SPC_sd)

SPC_datD <- SPC_dat_day %>%
  filter(wt_sd<4.1 & SPC_sd <35)
  
#hist(SPC_datD$wt_sd)
#hist(SPC_datD$SPC_sd)
  
covariat_datq <- covariat_datq%>%
  left_join(SPC_datD, by=c("date", "site"))

### 
## Bring in morph data : 
morph_df <- read.csv("/Users/kellyloria/Documents/UNR/MSMmetab/stream_morphology_core.csv")%>%
  filter(quality =="trust") %>%
    mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%dT%H:%M:%SZ"),
           date=as.Date(datetime)) %>%
  rename(site="Site")


Morph_dat_day <- morph_df%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    v_m=mean(v_estimation, na.rm=T),
    w_m=mean(w, na.rm=T),
    depth_measure=mean(z, na.rm=T))


covariat_datq <- covariat_datq%>%
  left_join(Morph_dat_day, by=c("date", "site"))

#### DO ###
## BWL  
BWL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_BWL_mod_dat.rds") %>%
  mutate(site="BWL")
summary(BWL_dat) 

## BWU 
BWU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWU_k600/25_BWU_mod_dat.rds") %>%
  mutate(site="BWU")
summary(BWU_dat)

## GBL 
GBL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBL_mod_dat.rds") %>%
  mutate(site="GBL")
summary(GBL_dat)

## GBU 
GBU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBU_mod_dat.rds") %>%
  mutate(site="GBU")
summary(GBU_dat)


miniDOT_dat <- rbind(BWL_dat, BWU_dat, GBL_dat, GBU_dat)
miniDOT_dat_day <- miniDOT_dat%>%
  mutate(date =as.Date(solar.time))%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    do.obs_m=mean(DO.obs, na.rm=T),
    do.obs_sd=sd(DO.obs, na.rm=T),
    wtr_m=mean(temp.water, na.rm=T),
    wtr_sd=sd(temp.water, na.rm=T))

covariat_datq <- covariat_datq%>%
  left_join(miniDOT_dat_day, by=c("date", "site"))

### N-uptake metrics
n_up <- read.csv("/Users/kellyloria/Documents/UNR/Ncycle/BTC_N_Table_2025_figure.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) 

covariat_datq <- covariat_datq%>%
  left_join(n_up, by=c("date", "site"))

##================================
## Load metabolism model data
##================================
BWL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWL_Full_2025-02-06_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>% 
  mutate(site="BWL",
         catch="BW")

summary(BWL_met_binned)

GBL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBL_Full_2025-02-02_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBL",
         catch="GB")


BWU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWU_Full_2025-02-04_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="BWU",
         catch="BW")

GBU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBU_Full_2025-02-03_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date,GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBU",
         catch="GB")

metab_dat <- rbind(BWL_met_binned, BWU_met_binned, GBL_met_binned, GBU_met_binned)

covariat_datq <- covariat_datq%>%
  left_join(metab_dat, by=c("date", "site", "catch"))


### read in streamlight data
## BWL  
BWL_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_BWL.rds")
summary(BWL_SL_dat) 

## BWU 
BWU_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_BWU.rds") 
summary(BWU_SL_dat)

## GBL 
GBL_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_GBL.rds") 
summary(GBL_SL_dat)

## GBU 
GBU_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_GBU.rds") 
summary(GBU_SL_dat)

light_mod_df <- rbind(BWL_SL_dat, BWU_SL_dat, GBL_SL_dat, GBU_SL_dat)

light_mod_df_day <- light_mod_df%>%
  mutate(date=as.Date(local_time)) %>%
  group_by(site, date)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%
  dplyr::select(site, date, LAI, PAR_inc, PAR_bc, veg_shade, bank_shade, PAR_surface)


covariat_datq <- covariat_datq%>%
  left_join(light_mod_df_day, by=c("date", "site"))


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

covariat_datq$AFDM_ugcm2 <- (covariat_datq$AFDM_mgcm2 *1000)
#hist(covariat_datq$AFDM_ugcm2 )

covariat_datq$Biom_Chla <- (covariat_datq$AFDM_ugcm2)/(covariat_datq$Chla_ugL_Q)
#hist(covariat_datq$Biom_Chla)

summary(covariat_datq)

```

Pause to make a column for biomass to chla ratios And also make biomass
in in micro grams to more comparable to chl-a

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

covariat_datq$AFDM_ugcm2 <- (covariat_datq$AFDM_mgcm2 *1000)
#hist(covariat_datq$AFDM_ugcm2 )

covariat_datq$Biom_Chla <- (covariat_datq$AFDM_ugcm2)/(covariat_datq$Chla_ugL_Q)
#hist(covariat_datq$Biom_Chla)
```

Additive columns for combined surface and pore water nutrient as tot\_
in ugL

```{r, warning=F, size = 'small', echo=T, include=F, fig.width=6, fig.height=4}

covariat_datq$tot_NO3_ugL <- c(covariat_datq$NO3_mgL_dl_pw + covariat_datq$NO3_mgL_dl_sw) *1000

covariat_datq$tot_NH4_ugL <- c(covariat_datq$NH4_mgL_dl_pw + covariat_datq$NH4_mgL_dl_sw) *1000

covariat_datq$tot_PO4_ugL <- c(covariat_datq$PO4_ugL_dl_pw + covariat_datq$PO4_ugL_dl_pw)

covariat_datq$tot_DOC_ugL <- c(covariat_datq$DOC_mgL_dl_pw + covariat_datq$DOC_mgL_dl_sw) *1000


summary(covariat_datq)

```

```{r, warning=F, size = 'small', echo=F, fig.width=10, fig.height=12}
# Separate out BW 
covariat_datq_BW <- covariat_datq%>%
  dplyr::filter(site=="BWL" | site=="BWU")

# Separate out GB 
covariat_datq_GB <- covariat_datq%>%
  dplyr::filter(site=="GBL" | site=="GBU")
```

##### Kernal density plots of GPP and ER

Colored by years

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=7, fig.height=8}
library(ggplot2)

BWL_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWL" &water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
     scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
          annotate("text", x = 7, y = -2, label = "BW Lower", size = 5, fontface = "bold", hjust = 0.5) 

BWU_DF_fixed <- covariat_datq%>%dplyr::filter(site=="BWU") %>%
  filter(water_year < 2024) %>%
  mutate(water_year = factor(water_year, levels = c(2021, 2022, 2023, 2024))) %>%
  bind_rows(tibble(GPP_mean = NA, ER_mean = NA, water_year = factor(2024, levels = c(2021, 2022, 2023, 2024))))  # Add dummy row


BWU_KD_plot <- ggplot(BWU_DF_fixed, aes(x = GPP_mean, y = ER_mean, color = water_year)) +
  geom_density_2d(bins = 10, size = 1) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
  xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
  ylab(NULL) +
  #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 7, y = -2, label = "BW Upper", size = 5, fontface = "bold", hjust = 0.5) 
BWU_KD_plot



GBU_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBU") %>%
filter(water_year<2025)%>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# Compute 2D density
density_data <- with(GBU_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))

# Convert density data to dataframe
density_df <- data.frame(
  expand.grid(x = density_data$x, y = density_data$y),
  z = as.vector(density_data$z)
)

# Plot with contour lines
GBU_KD_plot <- ggplot() +
  geom_point(data = GBU_DF_clean, aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
    geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 5, size=1) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
    ylab(NULL) +
  #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
    scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 7, y = -2, label = "GB Upper", size = 5, fontface = "bold", hjust = 0.5) 

GBU_KD_plot


GBL_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBL") %>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# 
# library(MASS)
# 
# # Compute 2D density
# density_data <- with(GBL_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))
# 
# # Convert density data to dataframe
# density_df <- data.frame(
#   expand.grid(x = density_data$x, y = density_data$y),
#   z = as.vector(density_data$z)
# )

# Plot with contour lines
GBL_KD_plot <- ggplot() +
#  geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 10) +
  geom_point(data = GBL_DF_clean%>%filter(water_year<2025), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_viridis_d(option = "viridis",  name = "Water year") +
  theme_classic() +
      annotate("text", x = 7, y = -2, label = "GB Lower", size = 5, fontface = "bold", hjust = 0.5) 


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4, fig.height=10}
# Modify each plot to rename legend title & adjust legend formatting
BWL_KD_plot <- BWL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

BWU_KD_plot <- BWU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBL_KD_plot <- GBL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBU_KD_plot <- GBU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))
```

### Modified figure 2

```{r, warning=F, size = 'small', echo=F, fig.width=9, fig.height=7}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  BWL_KD_plot,
  BWU_KD_plot,
  GBL_KD_plot,
  GBU_KD_plot,
  labels = c("a", "b", "c", "d"),
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom"
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig2_KD_CH1_grid_v2.png", plot = KD_coeff_grid, width = 6, height = 5.25, units = "in")

```

### Focal Question:

### How does stream metabolism vary based on antecedent flow conditions, and nitrogen availability across small and large stream?

Can we frame wet years as a shrinking productivity window? Or is it more
a function of a mismatch in phenology ques, where peak light occurs at
unsuitable streamflows (e.g. to high) for epilithic algae?

##### Ideas:

(1) Figure out what flow threshold at each site we can't really model
    GPP at?

    (a) How large is that window from year to year?

    (b) How is the cumulative GPP during that time?

    (c) How is the cumulative PAR and water temp?

<!-- -->

(2) Center dynamics on that flow threshold
    (a) X-axis could be time since peak "snow pack metric" or "flow
        metric"

<!-- -->

(3) Double check there isn't evidence of UV suppression, should double
    check magnitude of surface PAR

```{r, warning=F, size = 'small', echo=F, include=FALSE, include=F}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
  filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
    filter(site=="BWL") %>%
  filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

plot_data_GPP <- covariat_datq %>%
      filter(site=="BWL") %>%
  filter(!is.na(GPP_mean)) %>%
  dplyr:: select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt <- ggplot(covariat_datq, aes(x = date)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data = filter(plot_data_DO, has_do), aes(y = 0.095), alpha=0.3, color = "black", size = 1) + # DO presence tick marks
  geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.090), alpha= 0.3, color = "#579467", size = 1) + # GPP presence tick marks
    geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.085), alpha= 0.02, color = "#fffcfd", size = 1) + # GPP presence tick marks

  scale_x_date(date_breaks = "3 months", date_labels = "%b-%y",  limits = date_range) +  # Labels every 2 months
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Lower")
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, include=F}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
  filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
    filter(site=="BWU") %>%
  filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

plot_data_GPP <- covariat_datq %>%
      filter(site=="BWU") %>%
  filter(!is.na(GPP_mean)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt_BWU <- ggplot(covariat_datq, aes(x = date)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data = filter(plot_data_DO, has_do), aes(y = 0.095), alpha=0.3, color = "black", size = 1) + # DO presence tick marks
  geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.090), alpha= 0.3, color = "#579467", size = 1) + # GPP presence tick marks
    geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.085), alpha= 0.02, color = "#fffcfd", size = 1) + # GPP presence tick marks

  scale_x_date(date_breaks = "3 months", date_labels = "%b-%y",  limits = date_range) +  # Labels every 2 months
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "BW Upper")
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, include=F}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
  filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
    filter(site=="GBL") %>%
  filter(!is.na(do.obs_m)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

plot_data_GPP <- covariat_datq %>%
      filter(site=="GBL") %>%
  filter(!is.na(GPP_mean)) %>%
   dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt_GBL <- ggplot(covariat_datq, aes(x = date)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data = filter(plot_data_DO, has_do), aes(y = 0.095), alpha=0.3, color = "black", size = 1) + # DO presence tick marks
  geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.090), alpha= 0.3, color = "#579467", size = 1) + # GPP presence tick marks
    geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.085), alpha= 0.02, color = "#fffcfd", size = 1) + # GPP presence tick marks

  scale_x_date(date_breaks = "3 months", date_labels = "%b-%y",  limits = date_range) +  # Labels every 2 months
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Lower")
```

```{r, warning=F, size = 'small', echo=F, include=FALSE, include=F}
# Filter the dataset to only include rows where do.obs_m or GPP_mean are not NA
plot_data_date <- covariat_datq %>%
  filter(!is.na(do.obs_m))

plot_data_DO <- covariat_datq %>%
    filter(site=="GBU") %>%
  filter(!is.na(do.obs_m)) %>%
  dplyr::select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

plot_data_GPP <- covariat_datq %>%
  filter(site=="GBU") %>%
  filter(!is.na(GPP_mean)) %>%
  dplyr:: select(date, do.obs_m, GPP_mean) %>%
  mutate(has_do = !is.na(do.obs_m), has_GPP = !is.na(GPP_mean))

# Determine the date range
date_range <- range(plot_data_date$date, na.rm = TRUE)

# Create the number line plot
templt_GBU <- ggplot(covariat_datq, aes(x = date)) +
  #geom_segment(aes(x = date_range[1], xend = date_range[2], y = 0, yend = 0), linewidth = 0.2) + # Main line
  geom_point(data = filter(plot_data_DO, has_do), aes(y = 0.095), alpha=0.3, color = "black", size = 1) + # DO presence tick marks
  geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.090), alpha= 0.3, color = "#579467", size = 1) + # GPP presence tick marks
    geom_point(data = filter(plot_data_GPP, has_GPP), aes(y = 0.085), alpha= 0.02, color = "#fffcfd", size = 1) + # GPP presence tick marks

  scale_x_date(date_breaks = "3 months", date_labels = "%b-%y",  limits = date_range) +  # Labels every 2 months
  scale_y_continuous(name = "", breaks = NULL) +
  theme_classic() + xlab(NULL) +
    labs(title = "GB Upper",
       caption = "Black: DO Observed | Green: GPP Mean")
```

#### Data resolution for DO and metab

DO in black, metabolism in green

```{r, warning=F, size = 'small', echo=T, fig.width=10, fig.height=4.5}
### big grid 
WY_n_grid2 <- ggarrange(
  templt,
  templt_BWU,
  templt_GBL,
  templt_GBU,
  ncol = 1, nrow = 4)

WY_n_grid2
```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Sup_Figures/metab_dat_resolution.png", plot = WY_n_grid2, width = 8.25, height = 4.25, units = "in")

```

#### Get snotel data

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
library(snotelr)

# read in the sites data for all SNOTEL sites
target_snow_data <- snotelr::snotel_download(network = "sntl",
                                             site_id = c(848, 615), 
                                             metric = TRUE, internal = TRUE)
target_snow_data$date<- as.Date(target_snow_data$date)

# transform for water year:
target_snow_dat<- water_year(target_snow_data)

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}


#Trim for columns 
target_snow_df <- target_snow_dat %>%
  filter(water_year>2003) %>%
  filter(date < as.Date("2024-10-10")) %>%
  mutate(site = case_when(
    site_id == 615 ~ "GBL",
    site_id %in% c(848) ~ "BWL")) %>%
  dplyr::select(site_name,site_id,date,snow_water_equivalent,precipitation_cumulative,
         temperature_max,temperature_min,temperature_mean,precipitation, water_year,
         site)

summary(target_snow_df)


result <- target_snow_df %>%
  filter(date > as.Date("2020-09-30"))%>%
  arrange(site, date) %>%
  group_by(site) %>%
  mutate(
    daily_delta_SWE = snow_water_equivalent - lag(snow_water_equivalent, default = 0),
    snow_water_equivalent_cumulative = cumsum(daily_delta_SWE),
    precipitation_quality = ifelse(daily_delta_SWE > 0, "Accumulate", "Melt"), # not really rain
    rain = ifelse(daily_delta_SWE < 0, 1, 0)  # again just a really simple proxy
  ) %>%
  ungroup()


 summarize_transition_to_zero <- function(data) {
  data %>%
    filter(lubridate::month(date) >= 4 & lubridate::month(date) <= 7) %>%
    arrange(site, date) %>%
    group_by(site, WaterYear = lubridate::year(date) + if_else(lubridate::month(date) < 10, 0, 1)) %>%
    summarize(
      transition_to_zero = date[which.max(
        (daily_delta_SWE == 0 & lead(daily_delta_SWE == 0, default = FALSE) &
         lead(daily_delta_SWE == 0, n = 8, default = FALSE) &
         lag(daily_delta_SWE, default = 0) == 0) &
         (lead(daily_delta_SWE == 0, n = 8) & lag(daily_delta_SWE == 0, n = 5))
      )],
      .groups = 'drop',
      melt_out_DOY = lubridate::yday(transition_to_zero),
    ) %>%
    filter(!is.na(transition_to_zero))
 }
 

summarize_peak_SWE_date <- function(data) {
  data %>%
    filter(lubridate::month(date) >= 1 & lubridate::month(date) <= 6) %>%
    arrange(site, date) %>%
    group_by(site, WaterYear = lubridate::year(date) + if_else(lubridate::month(date) < 10, 0, 1)) %>%
    summarize(
      peak_SWE_date = date[which.max(snow_water_equivalent)],
      peak_SWE_day_of_year = lubridate::yday(date[which.max(snow_water_equivalent)]),
      max_SWE = max(snow_water_equivalent)
    ) %>%
    ungroup()
}

sum_melt <- summarize_transition_to_zero(result)

sum_peakSWE <- summarize_peak_SWE_date(result)

joint_snow<- sum_peakSWE %>%
left_join(sum_melt, by=c("site", "WaterYear"))

```

#### Rough time series plots to start getting at the antecedent flow conditions

```{r, warning=F, size = 'small', echo=F,include=FALSE, fig.width=10, fig.height=6}

TS_plot <- ggplot(covariat_datq%>%mutate(DOY=yday(date))%>%filter(site=="BWL")%>%filter(water_year<2024), aes(x = DOY, y = PAR_surface, color =as.factor(water_year), shape = site)) +
  geom_point(size = 2, alpha = 0.5, position = position_dodge(width = 0.3)) +
 # geom_line(aes(x = DOY, y=PAR_inc )) +
    geom_line(aes(x = DOY, y=(Q_m*100))) +
 scale_color_viridis_d(option = "viridis") +
  geom_point(size = 2, alpha = 0.7) +
  #scale_color_manual(values = siteC_colors) +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  ylab(expression(PAR~at~stream~surface~(mu~mol~m^-2~s^-1))) +
  theme_classic() + facet_grid(site~.)

TS_plot



TS_plot1 <- ggplot(covariat_datq%>%filter(water_year<2024), aes(x = Q_m, y = GPP_mean, color =as.factor(water_year), shape = site)) +
  geom_point(size = 2, alpha = 0.5, position = position_dodge(width = 0.3)) +
 # geom_line(aes(x = DOY, y=PAR_inc )) +
 #scale_color_viridis_d(option = "viridis") +
  geom_point(size = 2, alpha = 0.7) +
  #scale_color_manual(values = siteC_colors) +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  #ylab(expression(PAR~at~stream~surface~(mu~mol~m^-2~s^-1))) +
  theme_classic() + facet_grid(site~.)

TS_plot1
```

### (1) Figure out what flow threshold at each site we can't really model GPP at?

#### flow stats for when GPP was modeled

Max Q where GPP was obtained...

-   BWL : 2.763 cms

-   BWU : 5.990 cms

-   GBL : 0.622 cms

-   GBU : 0.299 cms

```{r, warning=F, size = 'small', echo=F, include=T, include=F}
covariat_datq_BWL_na <- covariat_datq %>%
  filter(site=="BWL") %>%
  drop_na(GPP_mean)
# max flow: 2.76302
# 3rd q: 0.18728
# mean q: 0.23213

covariat_datq_BWU_na <- covariat_datq %>%
    filter(site=="BWU") %>%
  drop_na(GPP_mean)

# max flow: 5.989955
# 3rd q: 0.278157
# mean q: 0.665920


covariat_datq_GBL_na <- covariat_datq %>%
    filter(site=="GBL") %>%
  drop_na(GPP_mean)

# max flow: 0.62202
# 3rd q: 0.03807
# mean q: 0.04228


covariat_datq_GBU_na <- covariat_datq %>%
    filter(site=="GBU") %>%
  drop_na(GPP_mean)

# max flow: 0.298702
# 3rd q: 0.037678
# mean q: 0.029493
```

#### Get DOY dates for max light and flow

```{r, warning=F, size = 'small', echo=F, include=FALSE, include=F}
sum_peak_Q <- function(data) {
  data %>%
    arrange(site, date) %>%
    group_by(site, water_year) %>%
    ## filter for spring runoff
    mutate(mon = month(date))%>%
    filter(mon >1 & mon <8) %>%
    summarize(
      peak_Q_date = date[which.max(Q_m)],
      peak_Q_DOY = lubridate::yday(date[which.max(Q_m)]),
      max_Q = max(Q_m, na.rm=T)
    ) %>%
    ungroup()
}



sum_peak_SUR_PAR <- function(data) {
  data %>%
    arrange(site, date) %>%
    group_by(site, water_year) %>%
    ## filter for spring runoff
    mutate(mon = month(date))%>%
    summarize(
      peak_PAR_date = date[which.max(PAR_surface)],
      peak_PAR_DOY = lubridate::yday(date[which.max(PAR_surface)]),
      max_PAR = max(PAR_surface, na.rm=T),
      peak_LAI_date = date[which.max(LAI)],
      peak_LAI_DOY = lubridate::yday(date[which.max(LAI)]),
      max_LAI = max(LAI, na.rm=T)
    ) %>%
    ungroup()
}


sum_par <- sum_peak_SUR_PAR(covariat_datq)

sum_Q_df <- sum_peak_Q(covariat_datq)

joint_WY<- sum_Q_df %>%
left_join(sum_par, by=c("site", "water_year"))

```

##### Manually identify baseflow onset

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}

p <- covariat_datq %>%
  mutate(yday=yday(date))%>%
  filter(site == "GBU") %>%
  ggplot(aes(x = yday, y = Q_m, col = as.factor(water_year))) +
  geom_point() +
  theme_bw() 

# Convert ggplot to plotly
p_plotly <- ggplotly(p)
p_plotly
```

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}
### make dataframe of baseflow onset at each reach 
BF_df <- data.frame(
  site = c("BWL", "BWL", "BWL", "BWL",
           "BWU", "BWU", "BWU",
           "GBL", "GBL", "GBL", "GBL",
           "GBU", "GBU", "GBU", "GBU"),
  
  water_year = c(2021, 2022, 2023, 2024,
                 2021, 2022, 2023,
                 2021, 2022, 2023, 2024,
                 2021, 2022, 2023, 2024),
  
  BF_DOY = c(167, 190, 219, 189,
             164, 185, 216,
             130, 142, 200, 165,
             140, 155, 200, 165)
)

```

#### Looking at Cumulative GPP

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}
# Ensure the date column is in Date format
covariat_datq1 <- covariat_datq %>%
  mutate(
    # Define water year DOY: Oct 1 is DOY 1, Sep 30 is DOY 365 (or 366 for leap years)
    DOY = as.numeric(difftime(date, ymd(paste0(water_year - 1, "-09-30")), units = "days"))
  )

# Compute cumulative sum of GPP_mean for each site and water year
covariat_cumsum <- covariat_datq1 %>%
  mutate(yday = yday(date),
         year = year(date)) %>%
  arrange(site, year, yday) %>%
  group_by(site, year) %>%
  mutate(GPP_cumsum = cumsum(ifelse(is.na(GPP_mean), 0, GPP_mean)))


# Merge covariat_cumsum with joint_WY to include peak_PAR_DOY
covariat_cumsum <- covariat_cumsum %>%
  left_join(joint_WY %>% dplyr::select(site, water_year, peak_PAR_DOY), 
            by = c("site" = "site", "year" = "water_year")) %>%
  left_join(BF_df %>% dplyr::select(site, water_year, BF_DOY), 
            by = c("site" = "site", "year" = "water_year"))

# Plot cumulative GPP across years with vertical lines for peak_PAR_DOY
ggplot(covariat_cumsum %>% filter(year < 2024), 
       aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line() + 
  labs(x = "Day of Water Year", y = "Cumulative GPP", color = "Water Year") +
  theme_classic() + 
  facet_grid(site ~ ., scales = "free")
```

Dotted lines for onset of baseflow Dashed lines for peak PAR

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}

# Plot cumulative GPP across years with vertical lines for peak_PAR_DOY
ggplot(covariat_cumsum %>% filter(year < 2024), 
       aes(x = yday, y = GPP_cumsum, color = factor(year))) +
  geom_line() + 
  geom_vline(aes(xintercept = peak_PAR_DOY, color = factor(year)), 
             linetype = "dashed", alpha = 0.9) + 
  geom_vline(aes(xintercept = BF_DOY, color = factor(year)), 
             linetype = "dotted", alpha = 0.9) +
  labs(x = "Day of Water Year", y = "Cumulative GPP", color = "Water Year") +
  theme_classic() + 
  facet_grid(site ~ ., scales = "free")

```

#### Looking at cumulative GPP + light + flow at each reach:

### BW Lower

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}

scale_factor_Qm <- max(covariat_cumsum$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum$Q_m, na.rm = TRUE)
scale_factor_PAR <- max(covariat_cumsum$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum$PAR_surface, na.rm = TRUE)

BWL_plot <- ggplot(covariat_cumsum %>% filter(site == "BWL"), aes(x = yday)) +
  # Cumulative GPP (Primary Y-axis)
  geom_line(aes(y = GPP_cumsum, color = GPP_mean), size = 1.5) + 
  
  # Color scale for GPP_mean: lighter for lower values and darker for higher values
  scale_color_gradient(low = "#b3d4a9", high = "#284a31") +  # Lighter to darker green
  
  scale_y_continuous(
    sec.axis = sec_axis(~ . * scale_factor_Qm, name = "Flow (cms)")  # Adjust `scale_factor_Qm`
  ) +
  
  # Flow (Q_m) as a second y-axis (scaled)
  geom_line(aes(y = Q_m * scale_factor_Qm), color = "#1c67a3", alpha = 0.9) + 
  
  # PAR_surface normalized (third variable, rescaled)
  geom_line(aes(y = PAR_surface * scale_factor_PAR), color = "goldenrod", linetype = "dotted") +
  
  labs(x = "Day of Water Year", color = "GPP Mean", 
       title = "BW Lower") +
    ylab(expression(Cumulative~GPP~(g~O[2]~m^-2~d^-1))) +
  theme_classic() + 
  facet_grid(year ~ ., scales = "free") 
BWL_plot

```

### BW Upper

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}
BWU_plot <- ggplot(covariat_cumsum %>% filter(site == "BWU" & water_year<2024), aes(x = yday)) +
  # Cumulative GPP (Primary Y-axis)
  geom_line(aes(y = GPP_cumsum, color = GPP_mean), size = 1.5) + 
  
  # Color scale for GPP_mean: lighter for lower values and darker for higher values
  scale_color_gradient(low = "#b3d4a9", high = "#284a31") +  # Lighter to darker green
  
  scale_y_continuous(
    sec.axis = sec_axis(~ . * scale_factor_Qm, name = "Flow (cms)")  # Adjust `scale_factor_Qm`
  ) +
  
  # Flow (Q_m) as a second y-axis (scaled)
  geom_line(aes(y = Q_m * scale_factor_Qm), color = "#1c67a3", alpha = 0.9) + 
  
  # PAR_surface normalized (third variable, rescaled)
  geom_line(aes(y = PAR_surface * scale_factor_PAR), color = "goldenrod", linetype = "dotted") +
  
  labs(x = "Day of Water Year", color = "GPP Mean", 
       title = "BW Upper") +
  ylab(expression(Cumulative~GPP~(g~O[2]~m^-2~d^-1))) +
  theme_classic() + 
  facet_grid(year ~ ., scales = "free") 
BWU_plot

```

### GB Lower

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}

covariat_cumsum_gb<- covariat_cumsum%>%filter(site=="GBL")

scale_factor_Qm <- max(covariat_cumsum_gb$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum_gb$Q_m, na.rm = TRUE)
scale_factor_PAR <- max(covariat_cumsum_gb$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum_gb$PAR_surface, na.rm = TRUE)


GBL_plot <- ggplot(covariat_cumsum_gb, aes(x = yday)) +
  # Cumulative GPP (Primary Y-axis)
  geom_line(aes(y = GPP_cumsum, color = GPP_mean), size = 1.5) + 
  
  # Color scale for GPP_mean: lighter for lower values and darker for higher values
  scale_color_gradient(low = "#b3d4a9", high = "#284a31") +  # Lighter to darker green
  
  scale_y_continuous(
    sec.axis = sec_axis(~ . * scale_factor_Qm, name = "Flow (cms)")  # Adjust `scale_factor_Qm`
  ) +
  
  # Flow (Q_m) as a second y-axis (scaled)
  geom_line(aes(y = Q_m * scale_factor_Qm), color = "#1c67a3", alpha = 0.9) + 
  
  # PAR_surface normalized (third variable, rescaled)
  geom_line(aes(y = PAR_surface * scale_factor_PAR), color = "goldenrod", linetype = "dotted") +
  
  labs(x = "Day of Water Year", color = "GPP Mean", 
       title = "GB Lower") +
  ylab(expression(Cumulative~GPP~(g~O[2]~m^-2~d^-1))) +
  theme_classic() + 
  facet_grid(year ~ ., scales = "free") 
GBL_plot

```

### GB Upper

```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}

covariat_cumsum_gb<- covariat_cumsum%>%filter(site=="GBU")

scale_factor_Qm <- max(covariat_cumsum_gb$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum_gb$Q_m, na.rm = TRUE)
scale_factor_PAR <- max(covariat_cumsum_gb$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum_gb$PAR_surface, na.rm = TRUE)


GBU_plot <- ggplot(covariat_cumsum_gb, aes(x = yday)) +
  # Cumulative GPP (Primary Y-axis)
  geom_line(aes(y = GPP_cumsum, color = GPP_mean), size = 1.5) + 
  
  # Color scale for GPP_mean: lighter for lower values and darker for higher values
  scale_color_gradient(low = "#b3d4a9", high = "#284a31") +  # Lighter to darker green
  
  scale_y_continuous(
    sec.axis = sec_axis(~ . * scale_factor_Qm, name = "Flow (cms)")  # Adjust `scale_factor_Qm`
  ) +
  
  # Flow (Q_m) as a second y-axis (scaled)
  geom_line(aes(y = Q_m * scale_factor_Qm), color = "#1c67a3", alpha = 0.9) + 
  
  # PAR_surface normalized (third variable, rescaled)
  geom_line(aes(y = PAR_surface * scale_factor_PAR), color = "goldenrod", linetype = "dotted") +
  
  labs(x = "Day of Water Year", color = "GPP Mean", 
       title = "GB Upper") +
  ylab(expression(Cumulative~GPP~(g~O[2]~m^-2~d^-1))) +
  theme_classic() + 
  facet_grid(year ~ ., scales = "free") 
GBU_plot

```


```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}
### Leaf area index 2021-2023

plot <- ggplot(joint_WY %>% filter(water_year < 2024)) +
  geom_point(aes(x = water_year, y = peak_LAI_DOY, color = as.factor(water_year), shape = site), 
             size = 3, alpha = 0.95) +
  
  geom_line(aes(x = water_year, y = peak_LAI_DOY, group = site), 
            alpha = 0.95) +  # Ensure lines connect correctly
  
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  theme_classic() +facet_grid(site ~ .)  # Keeps each site in a separate panel

plot

```


```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}
### Surface PAR 2021-2023

plot <- ggplot(joint_WY %>% filter(water_year < 2024)) +
  geom_point(aes(x = water_year, y = peak_PAR_DOY, color = as.factor(water_year), shape = site), 
             size = 3, alpha = 0.95) +
  
  geom_line(aes(x = water_year, y = peak_PAR_DOY, group = site), 
            alpha = 0.95) +  # Ensure lines connect correctly
  
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  theme_classic() +
  facet_grid(site ~ .)  # Keeps each site in a separate panel

plot

```


```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}
### Peak spring Q 2021-2023

plot <- ggplot(joint_WY %>% filter(water_year < 2024)) +
  geom_point(aes(x = water_year, y = peak_Q_DOY, color = as.factor(water_year), shape = site), 
             size = 3, alpha = 0.95) +
  geom_line(aes(x = water_year, y = peak_Q_DOY, group = site), 
            alpha = 0.95) +  # Ensure lines connect correctly
  
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  theme_classic() +
  facet_grid(site ~ .)  # Keeps each site in a separate panel

plot

```

### Trying to parse out why the GPP signal isn’t always decernamble: 

#### Logistic flow plots

Where 0 = DO/Temp data but either no metabolism estimate or GPP/ER = 0

Where 1 = DO/Temp data and GPP > 0 and ER < 0


```{r, echo=FALSE, error=FALSE, warning=FALSE, include=F, message=FALSE}

BWL_logQ_df_gpp <- covariat_datq %>%
  filter(site == "BWL") %>%
  filter(date > as.Date("2021-04-28")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
    is.na(GPP_mean) ~ 0,

    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(GPP_mean) & GPP_mean > 0.0001 ~ 1,

    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


BWL_logQ_df_er <- covariat_datq %>%
  filter(site == "BWL") %>%
  filter(date > as.Date("2021-04-28")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
      is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(ER_mean) & ER_mean < -0.0001 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


####
BWU_logQ_df_gpp <- covariat_datq %>%
  filter(site == "BWU") %>%
  filter(date > as.Date("2021-06-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
    is.na(GPP_mean) & is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(GPP_mean) & GPP_mean > 0.01 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


BWU_logQ_df_er <- covariat_datq %>%
  filter(site == "BWU") %>%
  filter(date > as.Date("2021-06-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
      is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(ER_mean) & ER_mean < -0.01 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


####
GBL_logQ_df_gpp <- covariat_datq %>%
  filter(site == "GBL") %>%
  filter(date > as.Date("2021-03-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
    is.na(GPP_mean) & is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(GPP_mean) & GPP_mean > 0.00001 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


GBL_logQ_df_er <- covariat_datq %>%
  filter(site == "GBL") %>%
  filter(date > as.Date("2021-03-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
      is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(ER_mean) & ER_mean < -0.00001 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))



####
GBU_logQ_df_gpp <- covariat_datq %>%
  filter(site == "GBU") %>%
  filter(date > as.Date("2021-03-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
    is.na(GPP_mean) & is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(GPP_mean) & GPP_mean > 0.0001 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))


GBU_logQ_df_er <- covariat_datq %>%
  filter(site == "GBU") %>%
  filter(date > as.Date("2021-03-29")) %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(site, catch, date, jday, water_year, do.obs_m, wtr_m,
                Q_m, depth_m, GPP_mean, ER_mean, K600_daily_mean) %>%
  mutate(GPP_bi = case_when(
    # GPP_bi = 0 when do.obs_m, wtr_m, Q_m, and depth_m are present AND metabolism estimates are missing
    !is.na(do.obs_m) & !is.na(wtr_m) & !is.na(Q_m) & !is.na(depth_m) & 
      is.na(ER_mean) ~ 0,
    # GPP_bi = 1 when GPP_mean > 0 AND ER_mean < 0, ensuring both are not NA
    !is.na(ER_mean) & ER_mean < -0.0001 ~ 1,
    # Default case: set GPP_bi = 0
    TRUE ~ 0  
  ))

```

#### GPP and flow thresholds?

```{r, warning=F, size = 'small', echo=T, include=T, fig.width=7, fig.height=3.75}

BWL_logQ_plt <- ggplot(BWL_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE, linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
  #facet_grid(water_year ~ ., scales = "free") +
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Lower")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}
BWU_logQ_plt <- ggplot(BWU_logQ_df_gpp %>% filter(water_year < 2024), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
 # facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Upper")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}

logQ_pltGBL <- ggplot(GBL_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Lower")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}

logQ_plt_GBU <- ggplot(GBU_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Upper")

```


```{r, warning=F, size = 'small', echo=F, include=T, fig.width=9, fig.height=5}
grid <- ggarrange(
  BWL_logQ_plt,
  BWU_logQ_plt,
  logQ_pltGBL,
  logQ_plt_GBU,
    common.legend = T, 
  legend = "bottom",
  ncol = 2, nrow = 2)
grid
```

Logistic growth lines color by year, individial points color by day of year (light= later)/

#### ER and flow thresholds?

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}

BWL_logQ_plt <- ggplot(BWL_logQ_df_er %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(ER[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
  #facet_grid(water_year ~ ., scales = "free") +
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Lower")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}
BWU_logQ_plt <- ggplot(BWU_logQ_df_er %>% filter(water_year < 2024), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(ER[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
 # facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Upper")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}

logQ_pltGBL <- ggplot(GBL_logQ_df_er %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(ER[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Lower")

```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=7, fig.height=3.75}

logQ_plt_GBU <- ggplot(GBU_logQ_df_er %>% filter(water_year < 2025), aes(x = log(Q_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) + # Custom colors for years

  # Labels and theme
  ylab(expression(ER[bi]~(0~or~1))) +
  xlab(expression(log(Q[m]))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Upper")

```



```{r, warning=F, size = 'small', echo=F, include=T, fig.width=9, fig.height=5}
grid <- ggarrange(
  BWL_logQ_plt,
  BWU_logQ_plt,
  logQ_pltGBL,
  logQ_plt_GBU,
    common.legend = T, 
  legend = "bottom",
  ncol = 2, nrow = 2)
grid
```


##### Table of flow thresholds where metabolism is unreasonable to model:

```{r, echo=T, error=FALSE, warning=FALSE, include=T, message=FALSE}
# Function to fit logistic regression and extract parameters
fit_logistic_stats <- function(data) {
  model <- glm(GPP_bi ~ log(Q_m), data = data, family = binomial)
  coef_vals <- coef(model)  # Extract coefficients
  
  # Calculate steepness (slope) and midpoint
  slope <- coef_vals[2]  
  midpoint <- -coef_vals[1] / coef_vals[2]  

  return(data.frame(
    water_year = unique(data$water_year),
    slope = slope,
    midpoint = midpoint
  ))
}

```

 
#### BW Lower

GPP:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}
# Apply function to each water_year
logistic_stats_gpp <- BWL_logQ_df_gpp %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()
# View the results
print(logistic_stats_gpp)

```

ER:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}

logistic_stats_er <- BWL_logQ_df_er %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()


print(logistic_stats_er)
```


#### BW Upper


GPP:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}
# Apply function to each water_year
logistic_stats_gpp <- BWU_logQ_df_gpp %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()
# View the results
print(logistic_stats_gpp)

```

ER:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}

logistic_stats_er <- BWU_logQ_df_er %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()


print(logistic_stats_er)
```


#### GB Lower

GPP:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}
# Apply function to each water_year
logistic_stats_gpp <- GBL_logQ_df_gpp %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()
# View the results
print(logistic_stats_gpp)

```

ER:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}

logistic_stats_er <- GBL_logQ_df_er %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()


print(logistic_stats_er)
```


#### GB Upper

GPP:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}
# Apply function to each water_year
logistic_stats_gpp <- GBU_logQ_df_gpp %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()
# View the results
print(logistic_stats_gpp)

```

ER:

```{r, echo=FALSE, error=FALSE, warning=FALSE, echo=FALSE, include=T, message=FALSE}

logistic_stats_er <- GBU_logQ_df_er %>%
  filter(water_year < 2025) %>%
  group_by(water_year) %>%
  group_split() %>%
  lapply(fit_logistic_stats) %>%
  bind_rows()


print(logistic_stats_er)
```


#### Water temp thersholds? 


```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=7, fig.height=3.75}

logQ_plt1 <- ggplot(BWL_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(wtr_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Temp))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Lower")



logQ_plt2 <- ggplot(BWU_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(wtr_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Temp))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Upper")


logQ_plt3 <- ggplot(GBL_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(wtr_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Temp))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Lower")

logQ_plt4 <- ggplot(GBU_logQ_df_gpp %>% filter(water_year < 2025), aes(x = log(wtr_m), y = GPP_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Temp))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Upper")

```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=9, fig.height=5}
grid <- ggarrange(
  logQ_plt1,
  logQ_plt2,
  logQ_plt3,
  logQ_plt4, 
  common.legend = T, 
  legend = "bottom",
  ncol = 2, nrow = 2)
grid
```


#### Water temp ~ Q thersholds? 
```{r, echo=T, error=FALSE, warning=FALSE, include=T, message=FALSE}
BWU_logQ_df_peak <- BWU_logQ_df_er %>%
  group_by(water_year) %>%
  filter(wtr_m == max(wtr_m, na.rm = TRUE)) %>%
  summarise(jday_peak_temp = min(jday, na.rm = TRUE))  # Get the day of the year for peak temperature

# Merge the peak temperature timing back to the original data
BWU_logQ_df_metab <- BWU_logQ_df_gpp %>%
  left_join(BWU_logQ_df_peak, by = "water_year")

# Step 2: Plot logistic regression with peak temperature timing (jday_peak_temp) on the x-axis
BWU_logistic_plt <- ggplot(BWU_logQ_df_metab %>% filter(water_year < 2025), aes(x = jday_peak_temp, y = GPP_bi)) +
  # Points colored by Q_m (flow)
  geom_point(aes(fill = log(Q_m)), color = "black", shape = 21, alpha = 0.9) +  
  scale_fill_viridis_c(name = "log(Q_m + 1)") +  # Color by log(Q_m)
  
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), 
              method = "glm", method.args = list(family = "binomial"), 
              se = FALSE) +
  
  # Custom colors for different years
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", 
                                "2023" = "#003049", "2024" = "#d1ac00")) + 
  
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab("Julian Day of Peak Temperature") +
  theme_classic() +
  
  # Facet by water year
  facet_grid(water_year ~ ., scales = "free") +
  
  # Adjust legends
  guides(fill = guide_colorbar(title = "log(Q_m + 1)"), 
         color = guide_legend(title = "Water Year"))

BWU_logistic_plt


```



##### Based on logistic growth
Flow thresholds based on midpoint in metabolism logistic plots 
```{r, echo=T, error=FALSE, warning=FALSE, include=T, message=FALSE}

BWL_logQ_df <- BWL_logQ_df_gpp %>%
  mutate(Q_bi = case_when(
    (water_year == 2021 & Q_m > 1.27) ~ 0,
    (water_year == 2022 & Q_m > 0.10) ~ 0,
    (water_year == 2023 & Q_m > 0.007) ~ 0,
    (water_year == 2024 & Q_m > 0.154) ~ 0,
    (water_year == 2021 & Q_m < 1.27) ~ 1,
    (water_year == 2022 & Q_m < 0.10) ~ 1,
    (water_year == 2023 & Q_m < 0.007) ~ 1,
    (water_year == 2024 & Q_m < 0.154) ~ 1,
    TRUE ~ 0
  ))




BWU_logQ_df <- BWU_logQ_df_gpp %>%
  mutate(Q_bi = case_when(
    (water_year == 2021 & Q_m > 0.015) ~ 0,
    (water_year == 2022 & Q_m > 0.016) ~ 0,
    (water_year == 2023 & Q_m > 0.940) ~ 0,
    (water_year == 2024 & Q_m > 0.037) ~ 0,
    (water_year == 2021 & Q_m < 0.015) ~ 1,
    (water_year == 2022 & Q_m < 0.016) ~ 1,
    (water_year == 2023 & Q_m < 0.940) ~ 1,
    (water_year == 2024 & Q_m < 0.037) ~ 1,
    TRUE ~ 0
  ))



GBL_logQ_df <- GBL_logQ_df_gpp %>%
  mutate(Q_bi = case_when(
    (water_year == 2021 & Q_m > 0.161) ~ 0,
    (water_year == 2022 & Q_m > 0.161) ~ 0,
    (water_year == 2023 & Q_m > 0.161) ~ 0,
    (water_year == 2024 & Q_m > 0.0015) ~ 0,
    (water_year == 2021 & Q_m < 0.161) ~ 1,
    (water_year == 2022 & Q_m < 0.161) ~ 1,
    (water_year == 2023 & Q_m < 0.161) ~ 1,
    (water_year == 2024 & Q_m < 0.0015) ~ 1,
    TRUE ~ 0
  ))


GBU_logQ_df <- GBU_logQ_df_gpp %>%
  mutate(Q_bi = case_when(
    (water_year == 2021 & Q_m >  0.0323) ~ 0,
    (water_year == 2022 & Q_m >  0.0113) ~ 0,
    (water_year == 2023 & Q_m > 0.0117) ~ 0,
    (water_year == 2024 & Q_m > 0.0539) ~ 0,
    (water_year == 2021 & Q_m <  0.0323) ~ 1,
    (water_year == 2022 & Q_m <  0.0113) ~ 1,
    (water_year == 2023 & Q_m < 0.0117) ~ 1,
    (water_year == 2024 & Q_m < 0.0539) ~ 1,
    TRUE ~ 0
  ))
```



```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=7, fig.height=3.75}

logQ_plt1 <- ggplot(BWL_logQ_df %>% filter(water_year < 2025), aes(x = (wtr_m), y = Q_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  ylab(expression(GPP[bi]~(0~or~1))) +
  xlab(expression(log(Temp))) +
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Lower")



logQ_plt2 <- ggplot(BWU_logQ_df %>% filter(water_year < 2025), aes(x = (wtr_m), y = Q_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "BW Upper")


logQ_plt3 <- ggplot(GBL_logQ_df %>% filter(water_year < 2025), aes(x = (wtr_m), y = Q_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Lower")

logQ_plt4 <- ggplot(GBU_logQ_df %>% filter(water_year < 2025), aes(x = (wtr_m), y = Q_bi)) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "glm", method.args = list(family = "binomial"), se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  theme_classic() +
#  facet_grid(water_year ~ ., scales = "free") +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  +
  labs(caption = "GB Upper")

```

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=9, fig.height=5}
grid <- ggarrange(
  logQ_plt1,
  logQ_plt2,
  logQ_plt3,
  logQ_plt4, 
  common.legend = T, 
  legend = "bottom",
  ncol = 2, nrow = 2)
grid
```



###

Q ~ K600 relationships

```{r, warning=F, size = 'small', echo=F, include=T, fig.width=11, fig.height=9}

logQ_plt1 <- ggplot(covariat_datq %>% filter(water_year < 2025) %>%mutate(jday=yday(date)), aes(x = log(wtr_m), y = log(K600_daily_mean))) +
  # Points colored by GPP_mean
  geom_point(aes(fill = jday), color = "black", shape = 21, alpha = 0.9) +  
  #scale_fill_gradient(low = "#b3d4a9", high = "#284a31") +  # Gradient for GPP_mean
  scale_fill_viridis_c(name="day of year") +  # Gradient for GPP_mean
  # Logistic regression curve colored by water_year
  geom_smooth(aes(color = as.factor(water_year)), method = "lm", se = FALSE,  linewidth=0.5) +
  scale_color_manual(values = c("2021" = "#D62828", "2022" = "#F77F00", "2023" = "#003049", "2024" = "#d1ac00")) +
  # Labels and theme
  theme_classic() +
 facet_wrap(water_year ~ site) +
  # Separate legends for points and lines
  guides(fill = guide_colorbar(title = "Day of Year"), 
         color = guide_legend(title = "Water Year"))  

logQ_plt1
```





### Code sandbox:



```{r, echo=FALSE, error=FALSE, warning=FALSE, include=T, message=FALSE}
# library(ggplot2)
# library(dplyr)
# library(patchwork)
# 
# # Ensure numeric values for scaling factors
# scale_factor_Qm <- max(covariat_cumsum$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum$Q_m, na.rm = TRUE)
# scale_factor_PAR <- max(covariat_cumsum$GPP_cumsum, na.rm = TRUE) / max(covariat_cumsum$PAR_surface, na.rm = TRUE)
# 
# # First Plot: Cumulative GPP
# p1 <- ggplot(covariat_cumsum %>% filter(site == "BWL"), aes(x = yday, y = GPP_cumsum)) +
#   geom_line(aes(color = GPP_mean), size = 1.5) + 
#   scale_color_gradient(low = "#b3d4a9", high = "#284a31") +  
#   labs(y = "Cumulative GPP", x = NULL, color = "GPP Mean") +
#   theme_classic() +
#   facet_grid(year ~ ., scales = "free")
# 
# # Second Plot: Flow (Q_m)
# p2 <- ggplot(covariat_cumsum %>% filter(site == "BWL"), aes(x = yday, y = Q_m)) +
#   geom_line(color = "#1c67a3", size = 1) +  
#   labs(y = "Flow (Q_m)", x = NULL) +
#   theme_classic() +
#   facet_grid(year ~ ., scales = "free")
# 
# # Third Plot: PAR_surface with scaled GPP and Flow
# p3 <- ggplot(covariat_cumsum %>% filter(site == "BWL"), aes(x = yday)) +
#   geom_line(aes(y = GPP_mean, color = "Cumulative GPP"), size = 1.5) + 
#   geom_line(aes(y = PAR_surface / scale_factor_PAR, color = "PAR_surface (scaled)"), 
#             linetype = "dotted", size = 1.2) +
#   geom_line(aes(y = Q_m * scale_factor_Qm, color = "Flow (scaled)"), size = 1) +
#   scale_color_manual(values = c(
#     "PAR_surface (scaled)" = "goldenrod", 
#     "Cumulative GPP" = "#284a31",  # Fixing the color issue
#     "Flow (scaled)" = "#1c67a3"
#   )) +
#   labs(y = "PAR_surface", x = "Day of Water Year", color = "Variables") +
#   theme_classic() +
#   theme(axis.title.y = element_text(color = "goldenrod")) + 
#   facet_grid(year ~ ., scales = "free")
# 
# # Extract y-axis labels
# ylab_p1 <- wrap_elements(get_plot_component(p1, "ylab-l"))
# ylab_p2 <- wrap_elements(get_plot_component(p2, "ylab-l"))
# 
# # Extract y-axes
# yaxis_p1 <- wrap_elements(get_y_axis(p1))
# yaxis_p2 <- wrap_elements(get_y_axis(p2))
# 
# # Arrange plots using patchwork
# final_plot <- ylab_p1 + yaxis_p1 + ylab_p2 + yaxis_p2 + p3 +
#   plot_layout(widths = c(3, 1, 3, 1, 40)) 
# 
# # Print the final plot
# print(final_plot)


```
