---
title: 'Antecedent flow conditions for MS'
author: "Kelly Loria"
date: "`r Sys.Date()`"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F, include=FALSE}
### Packages
#setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWL_k600/")
lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern", "broom.mixed",
         "glmmTMB", "plotly", "MASS",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
```

### MS: Spatiotemporal variation in mountain stream metabolism and nitrogen cycling across contrasting flow regimes

```{r, warning=F, size = 'small', echo=FALSE, include=FALSE}
daoc_date<- Sys.Date()

lapply(c("plyr","dplyr","ggplot2","cowplot",
         "lubridate","tidyverse", "reshape2", "ggpubr", "ggpattern",
         "lme4", "lmerTest", "MuMIn", "PerformanceAnalytics", "car"), require, character.only=T)
detach("package:plyr", unload = TRUE)
site_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#97b5c2",
  "GBL" = "#DD6E42",
  "GBU" = "#d9cb7c"
)

siteC_colors <- c(
  "BWL" = "#054fb9",
  "BWU" = "#054fb9",
  "GBL" = "#DD6E42",
  "GBU" = "#DD6E42"
)



catch_colors <- c(
  "BW" = "#054fb9",
  "GB" = "#DD6E42"
)

# Define fill patterns for water years
water_year_patterns <- c("wave","weave","stripe","circle", "stripe", "circle")


# Create a sequence of dates
date_seq <- seq(from = as.Date("2021-03-20"), to = as.Date("2024-10-10"), by = "day")

# Convert the sequence to a dataframe
date_df <- data.frame(date = date_seq)
date_df$site <- "BWL"
date_df$catch <- "BW"
date_df1 <- data.frame(date = date_seq)
date_df1$site <- "GBL"
date_df1$catch <- "GB"
date_df2 <- data.frame(date = date_seq)
date_df2$site <- "GBU"
date_df2$catch <- "GB"
date_df3 <- data.frame(date = date_seq)
date_df3$site <- "BWU"
date_df3$catch <- "BW"

date_datf <- rbind(date_df,date_df1,date_df2, date_df3)

### fxns:
## Fxn for water year:
water_year <- function(data) {
  # Ensure the `date` column exists
  if (!"date" %in% names(data)) {
    stop("The dataframe must contain a column named 'date'.")
  }
  
  # Convert the `date` column to Date format (if not already)
  data$date <- as.Date(data$date)
  
  # Add the water_year column
  data$water_year <- with(data, {
    year <- as.numeric(format(date, "%Y"))
    month <- as.numeric(format(date, "%m"))
    ifelse(month >= 10, year + 1, year)
  })
  
  return(data)
}
## Function to perform ANOVA and post-hoc test within each site
comparisons <- list(
  c("2021", "2022"),
  c("2021", "2023"),
  c("2022", "2023"),
  c("2023", "2024")
)
```

```{r, warning=F, size = 'small', echo=FALSE, include=F}
date_datf <- water_year(date_datf)

detach("package:plyr", unload = TRUE)


covariat_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_covariate_dat_AFDM.rds") %>%
  dplyr::select("date", "Site","bulk.density","AFDM_mgg","AFDM_mgcm2", "Chla_ugL_Q", "Pheo_ugL_Q")
summary(covariat_dat)

covariat_dat <- date_datf %>%
  left_join(covariat_dat, by=c("date", "site"="Site"))

flow_df <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_flow_df.rds")

covariat_dat <- covariat_dat %>%
    left_join(flow_df, by=c("date", "site"))


bg_nuts <- 
  readRDS("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/NS_chem_dat_nh4_24.rds") %>%
  filter(location=="stream") %>%
  mutate(NO3_mgL_dl = if_else(NO3_mgL_dl < 0.0015, 0.0015, NO3_mgL_dl))%>%
    mutate(NH4_mgL_dl=if_else(NH4_mgL_dl < 0.001, 0.001, NH4_mgL_dl))%>%
    mutate(PO4_ugL_dl=if_else(PO4_ugL_dl < 0.201, 0.201, PO4_ugL_dl)) %>%
  mutate(PO4_ugL_dl=if_else(DOC_mgL_dl < 0.25, 0.125, DOC_mgL_dl)) %>%
  group_by(site, shore, date, depth, location, substrate) %>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")


##### pivot wider 
bg_nuts_wide <- bg_nuts %>%
  dplyr::select(site, date, substrate, NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill) %>%
  group_by(site, date,substrate) %>% 
  summarise(across(c(NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill), mean, na.rm = TRUE), .groups = "drop") %>% 
  pivot_wider(names_from = substrate, values_from = c(NO3_mgL_dl, NH4_mgL_dl, PO4_ugL_dl, DOC_mgL_dl, pH_infill), names_sep = "_")

######


WQ_dat <- read.csv("/Users/kellyloria/Documents/LittoralMetabModeling/RawData/WaterChem/Stream_Lake_YSI_WaterQuality.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) %>%
  dplyr::select("site", "date","pH") %>%
  dplyr:: group_by(site, date) %>%
  dplyr::summarise(pH=mean(pH, na.rm=T))

covariat_datq <- covariat_dat%>%
  left_join(bg_nuts_wide[ , c("site", "date", "NO3_mgL_dl_sw", "NO3_mgL_dl_pw", "NH4_mgL_dl_sw", 
                              "NH4_mgL_dl_pw", "PO4_ugL_dl_sw","PO4_ugL_dl_pw", "DOC_mgL_dl_sw",
                              "DOC_mgL_dl_pw", "pH_infill_sw",  "pH_infill_pw")], by=c("date", "site"))

covariat_datq <- covariat_datq%>%
  left_join(WQ_dat,  by=c("date", "site"))


## bring in SPC data 
### missing 2024 BW SPC
SPC_BWL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWL_SPCv2.rds") %>%
  filter(datetime>as.POSIXct("2021-04-29 24:00:00"))
SPC_BWL$site <- "BWL"
SPC_BWU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_BWU_SPCv2.rds")
SPC_BWU$site <- "BWU"
SPC_GBL <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBL_SPCv2.rds")
SPC_GBL$site <- "GBL"
SPC_GBU <- readRDS("/Users/kellyloria/Documents/UNR/MSMmetab/23_CleanDat/24_GBU_SPCv2.rds")
SPC_GBU$site <- "GBU"

SPC_dat <- rbind(SPC_BWL, SPC_BWU, SPC_GBL, SPC_GBU)
SPC_dat <- SPC_dat %>%
  mutate(date =as.Date(datetime)) 

SPC_dat_day <- SPC_dat %>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    wt= mean(wtr, na.rm=T),
    wt_sd= sd(wtr, na.rm=T),
    SPC_m=mean(SPC, na.rm=T),
    SPC_sd=sd(SPC, na.rm=T))

#hist(SPC_dat_day$wt_sd)
#hist(SPC_dat_day$SPC_sd)

SPC_datD <- SPC_dat_day %>%
  filter(wt_sd<4.1 & SPC_sd <35)
  
#hist(SPC_datD$wt_sd)
#hist(SPC_datD$SPC_sd)
  
covariat_datq <- covariat_datq%>%
  left_join(SPC_datD, by=c("date", "site"))

### 
## Bring in morph data : 
morph_df <- read.csv("/Users/kellyloria/Documents/UNR/MSMmetab/stream_morphology_core.csv")%>%
  filter(quality =="trust") %>%
    mutate(datetime = as.POSIXct(datetime, format="%Y-%m-%dT%H:%M:%SZ"),
           date=as.Date(datetime)) %>%
  rename(site="Site")


Morph_dat_day <- morph_df%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    v_m=mean(v_estimation, na.rm=T),
    w_m=mean(w, na.rm=T),
    depth_measure=mean(z, na.rm=T))


covariat_datq <- covariat_datq%>%
  left_join(Morph_dat_day, by=c("date", "site"))

#### DO ###
## BWL  
BWL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_BWL_mod_dat.rds") %>%
  mutate(site="BWL")
summary(BWL_dat) 

## BWU 
BWU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/BWU_k600/25_BWU_mod_dat.rds") %>%
  mutate(site="BWU")
summary(BWU_dat)

## GBL 
GBL_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBL_mod_dat.rds") %>%
  mutate(site="GBL")
summary(GBL_dat)

## GBU 
GBU_dat <- readRDS("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/input_data/25_GBU_mod_dat.rds") %>%
  mutate(site="GBU")
summary(GBU_dat)


miniDOT_dat <- rbind(BWL_dat, BWU_dat, GBL_dat, GBU_dat)
miniDOT_dat_day <- miniDOT_dat%>%
  mutate(date =as.Date(solar.time))%>%
  dplyr::group_by(site, date) %>%
  dplyr::summarise(
    do.obs_m=mean(DO.obs, na.rm=T),
    do.obs_sd=sd(DO.obs, na.rm=T),
    wtr_m=mean(temp.water, na.rm=T),
    wtr_sd=sd(temp.water, na.rm=T))

covariat_datq <- covariat_datq%>%
  left_join(miniDOT_dat_day, by=c("date", "site"))

### N-uptake metrics
n_up <- read.csv("/Users/kellyloria/Documents/UNR/Ncycle/BTC_N_Table_2025_figure.csv")%>%
  mutate(date = as.Date(date, format="%m/%d/%y")) 

covariat_datq <- covariat_datq%>%
  left_join(n_up, by=c("date", "site"))

##================================
## Load metabolism model data
##================================
BWL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWL_Full_2025-02-06_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>% 
  mutate(site="BWL",
         catch="BW")

summary(BWL_met_binned)

GBL_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBL_Full_2025-02-02_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBL",
         catch="GB")


BWU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/BWU_Full_2025-02-04_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date, GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="BWU",
         catch="BW")

GBU_met_binned <- read_csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/final_output_2025/GBU_Full_2025-02-03_daily.csv") %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05) %>%
  mutate(
    GPP_mean = ifelse(GPP_mean < 0, 0.001, GPP_mean)) %>%
  filter(ER_mean < 0.01) %>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  dplyr::select(date,GPP_mean, ER_mean, NEP_daily_mean, K600_daily_mean,
                GPP_2.5pct, GPP_97.5pct, 
                ER_2.5pct, ER_97.5pct,
                K600_daily_2.5pct, K600_daily_predlog_97.5pct,
                GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  mutate(site="GBU",
         catch="GB")

metab_dat <- rbind(BWL_met_binned, BWU_met_binned, GBL_met_binned, GBU_met_binned)

covariat_datq <- covariat_datq%>%
  left_join(metab_dat, by=c("date", "site", "catch"))


### read in streamlight data
## BWL  
BWL_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_BWL.rds")
summary(BWL_SL_dat) 

## BWU 
BWU_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_BWU.rds") 
summary(BWU_SL_dat)

## GBL 
GBL_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_GBL.rds") 
summary(GBL_SL_dat)

## GBU 
GBU_SL_dat <- readRDS("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/data/CH1_StreamLightMod_GBU.rds") 
summary(GBU_SL_dat)

light_mod_df <- rbind(BWL_SL_dat, BWU_SL_dat, GBL_SL_dat, GBU_SL_dat)

light_mod_df_day <- light_mod_df%>%
  mutate(date=as.Date(local_time)) %>%
  group_by(site, date)%>%
  summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%
  dplyr::select(site, date, LAI, PAR_inc, PAR_bc, veg_shade, bank_shade, PAR_surface)


covariat_datq <- covariat_datq%>%
  left_join(light_mod_df_day, by=c("date", "site"))


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
# temp summary save file 


light_mod_df_day1 <- water_year(light_mod_df_day)

light_data_HA1 <- light_mod_df_day1%>%
  mutate(yday=yday(date)) %>%
  filter(yday<253) %>% # 253
  mutate(catch = case_when(
    site== "BWL"~ "BW",
    site== "BWU"~ "BW",
    site== "GBL"~ "GB",
    site== "GBU"~ "GB",
    TRUE ~ NA_character_  )) %>%
  # mutate(cum_PAR_surface = cumsum(PAR_surface)) %>%
  dplyr::group_by(site) %>%
  summarise(mean_LAI = round(mean(LAI, na.rm=T),2),
            mean_veg_shade = round(mean(veg_shade, na.rm = T),2),
            mean_PAR_surface = round(mean(PAR_surface, na.rm = T),2),
            min_PAR_surface = min(PAR_surface, na.rm = T),
            max_PAR_surface = max(PAR_surface, na.rm = T),
            )


light_data_HA1


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

covariat_datq$AFDM_ugcm2 <- (covariat_datq$AFDM_mgcm2 *1000)
#hist(covariat_datq$AFDM_ugcm2 )

covariat_datq$Biom_Chla <- (covariat_datq$AFDM_ugcm2)/(covariat_datq$Chla_ugL_Q)
#hist(covariat_datq$Biom_Chla)

summary(covariat_datq)

```

Pause to make a column for biomass to chla ratios And also make biomass
in in micro grams to more comparable to chl-a

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

covariat_datq$AFDM_ugcm2 <- (covariat_datq$AFDM_mgcm2 *1000)
#hist(covariat_datq$AFDM_ugcm2 )

covariat_datq$Biom_Chla <- (covariat_datq$AFDM_ugcm2)/(covariat_datq$Chla_ugL_Q)
#hist(covariat_datq$Biom_Chla)
```

Additive columns for combined surface and pore water nutrient as tot\_
in ugL

```{r, warning=F, size = 'small', echo=T, include=F, fig.width=6, fig.height=4}

covariat_datq$tot_NO3_ugL <- c(covariat_datq$NO3_mgL_dl_pw + covariat_datq$NO3_mgL_dl_sw) *1000

covariat_datq$tot_NH4_ugL <- c(covariat_datq$NH4_mgL_dl_pw + covariat_datq$NH4_mgL_dl_sw) *1000

covariat_datq$tot_PO4_ugL <- c(covariat_datq$PO4_ugL_dl_pw + covariat_datq$PO4_ugL_dl_pw)

covariat_datq$tot_DOC_ugL <- c(covariat_datq$DOC_mgL_dl_pw + covariat_datq$DOC_mgL_dl_sw) *1000


summary(covariat_datq)

```

```{r, warning=F, size = 'small', echo=F, fig.width=10, fig.height=12}
# Separate out BW 
covariat_datq_BW <- covariat_datq%>%
  dplyr::filter(site=="BWL" | site=="BWU")

# Separate out GB 
covariat_datq_GB <- covariat_datq%>%
  dplyr::filter(site=="GBL" | site=="GBU")
```

##### Kernal density plots of GPP and ER

Colored by years

```{r, warning=F, size = 'small', echo=F, include=FALSE, fig.width=7, fig.height=8}
library(ggplot2)

BWL_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWL" & water_year<2023), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
   scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
#   scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
          annotate("text", x = 7, y = -2, label = "BW Lower", size = 5, fontface = "bold", hjust = 0.5) 
# 
# BWU_DF_fixed <- covariat_datq%>%dplyr::filter(site=="BWU" & water_year<2022) 
# #%>%
#  # mutate(water_year = factor(water_year, levels = c(2021, 2022))) %>%
# #  bind_rows(tibble(GPP_mean = NA, ER_mean = NA, water_year = factor(2022, levels = c(2021, 2022))))  # Add dummy row


# BWU_KD_plot <- ggplot(BWU_DF_fixed, aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
#   geom_density_2d(bins = 10, size = 1) +
#   geom_point(alpha = 0.5) +
#   geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
#   scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
#   scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
#   xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
#   ylab(NULL) +
#   #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
#   #  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
#   scale_color_viridis_d(option = "viridis", name = "Water year") +
#   theme_classic() +
#         annotate("text", x = 7, y = -2, label = "BW Upper", size = 5, fontface = "bold", hjust = 0.5) 
# BWU_KD_plot


BWU_KD_plot <- ggplot(covariat_datq%>%dplyr::filter(site=="BWU" & water_year<2023), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year))) +
  geom_density_2d(bins=10, size=1) + 
  geom_point(alpha=0.5)+
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
 ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
  scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
#   scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
            annotate("text", x = 7, y = -2, label = "BW Upper", size = 5, fontface = "bold", hjust = 0.5) 



GBU_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBU") %>%
filter(water_year<2023)%>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# Compute 2D density
density_data <- with(GBU_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))

# Convert density data to dataframe
density_df <- data.frame(
  expand.grid(x = density_data$x, y = density_data$y),
  z = as.vector(density_data$z)
)

# Plot with contour lines
GBU_KD_plot <- ggplot() +
  geom_point(data = GBU_DF_clean, aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
    geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 5, size=1) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
    ylab(NULL) +
  #ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
    scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
    scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
    scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +

 # scale_color_viridis_d(option = "viridis", name = "Water year") +
  theme_classic() +
        annotate("text", x = 7, y = -2, label = "GB Upper", size = 5, fontface = "bold", hjust = 0.5) 

GBU_KD_plot


GBL_DF_clean <- covariat_datq%>%dplyr::filter(site=="GBL") %>%
  filter(!is.na(GPP_mean) & !is.na(ER_mean))%>%
  dplyr::select(date, water_year, GPP_mean, ER_mean)

# 
# library(MASS)
# 
# # Compute 2D density
# density_data <- with(GBL_DF_clean, kde2d(GPP_mean, ER_mean, n = 100))
# 
# # Convert density data to dataframe
# density_df <- data.frame(
#   expand.grid(x = density_data$x, y = density_data$y),
#   z = as.vector(density_data$z)
# )

# Plot with contour lines
GBL_KD_plot <- ggplot() +
#  geom_contour(data = density_df, aes(x = x, y = y, z = z), color = "black", bins = 10) +
  geom_point(data = GBL_DF_clean%>%filter(water_year<2023), aes(x = GPP_mean, y = ER_mean, color = as.factor(water_year)), alpha = 0.5) +
  geom_abline(slope = -1, intercept = 0, linetype = "dashed", color = "black") +  # 1:1 line
  scale_x_continuous(position = "top") +
    #xlab(expression(GPP~(g~O[2]~m^-2~d^-1))) +
   xlab(NULL) +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1))) +
  scale_y_continuous(limits = c(-30, 0)) +  # Ensure zero is visible
      scale_x_continuous(position = "top", limits = c(0, 13)) +  # Set x limits & move axis
 #scale_color_viridis_d(option = "viridis",  name = "Water year") +
    scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
  theme_classic() +
      annotate("text", x = 7, y = -2, label = "GB Lower", size = 5, fontface = "bold", hjust = 0.5) 


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=4, fig.height=10}
# Modify each plot to rename legend title & adjust legend formatting
BWL_KD_plot <- BWL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

BWU_KD_plot <- BWU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBL_KD_plot <- GBL_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))

GBU_KD_plot <- GBU_KD_plot + 
  guides(color = guide_legend(title = "Water year", nrow = 2))
```

### Modified figure 2

```{r, warning=F, size = 'small', echo=F, fig.width=9, fig.height=7}
# Arrange plots with shared legend at the bottom
KD_coeff_grid <- ggarrange(
  BWL_KD_plot,
  BWU_KD_plot,
  GBL_KD_plot,
  GBU_KD_plot,
  labels = c("a", "b", "c", "d"),
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  legend = "bottom"
)

KD_coeff_grid
```

```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig2_KD_CH1_grid_21_22.png", plot = KD_coeff_grid, width = 6, height = 5.25, units = "in")

```




### **Figure 2**
Setting: Showing the delayed onset of baseflow in wetter years.
Color as water year. 

DOY - April 1 = 91
DOY - May 1 = 121
DOY -June 1 = 152
DOY - Sept 1 = 244
DOY - Oct 1 = 274

```{r, warning=F, size = 'small', echo=F, include=FALSE}

names(covariat_datq)

covariat_datq<-covariat_datq%>%
  mutate(yday=yday(date))


overall_pf<- covariat_datq %>%
    filter(yday >= 92, yday <= 153) %>%
  group_by(site, water_year) %>%
  summarise(
    max_pf = max(Q_m, na.rm = TRUE),
    min_pf = min(Q_m, na.rm = TRUE),
    max_pf_yday = yday[which.max(Q_m)],
    min_pf_yday = yday[which.min(Q_m)],
    mean_pf = mean(Q_m, na.rm = TRUE),
    mean_pf_yday = yday[which.min(abs(Q_m - mean_pf))]) %>%
  ungroup()


overall_bf<- covariat_datq %>%
    filter(yday >= 170, yday <= 274) %>%
  group_by(site) %>%
  summarise(
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q = mean(Q_m, na.rm = TRUE),
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]) %>%
  ungroup()


overall_bf_df <- covariat_datq %>%
  left_join(overall_bf, by = c("site"))


steepest_slopes <- overall_bf_df %>%
    filter(yday >= 153, yday <= 274) %>%
  #  arrange(site, water_year, yday) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2023 ~ mean_q + 0.031,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2024 ~ mean_q + 0.01,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
    mutate(
    Q_m = case_when(
      site == "GBL" & water_year == 2021 ~ Q_m + 0.001,  # Adjust ave_q for GBL in 2021
      TRUE ~ Q_m)) %>%
  group_by(site, water_year) %>%
  summarise(
    ave_q = mean(mean_q, na.rm = TRUE),
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]
   # max_slope = max(slope, na.rm = TRUE),
  #  max_slope_yday = yday[which(slope == max(slope, na.rm = TRUE))[1]],
  #  min_slope = min(slope, na.rm = TRUE),
  #  min_slope_yday = yday[which(slope == min(slope, na.rm = TRUE))[1]]
  ) %>%
  ungroup() 

```

```{r, warning=F, size = 'small', echo=F, include=F}
# Merge covariat_datq with steepest_slopes to get mean_q_yday
Baseflowplots <- covariat_datq %>%
  dplyr::select(date, site, catch, water_year, Q_m, yday) %>%
  left_join(steepest_slopes %>% 
              dplyr::select(site, water_year, mean_q_yday, ave_q), by = c("site", "water_year","yday"= "mean_q_yday")) %>%
  left_join(overall_pf %>% 
              dplyr::select(site, water_year, max_pf, max_pf_yday), by = c("site", "water_year", "yday"= "max_pf_yday"))  %>%filter(site=="BWL" | site=="GBL")


Baseflowplot1 <- Baseflowplots%>%
  mutate(
    ave_q = case_when(
      site == "GBL" & water_year == 2021 ~ ave_q - 0.006,  # Adjust ave_q for GBL in 2021
      TRUE ~ ave_q))


Baseflowplot <- ggplot(Baseflowplot1 %>% 
                         filter(water_year < 2025), 
                       aes(x = yday, y = Q_m, color = as.factor(water_year))) +
  geom_line() +  
  scale_color_viridis_d(name = "Year") +  
  xlim(91, 234) + 
  facet_wrap(.~site, scales = "free") + 
  theme_classic() + 
  ylab(expression(Streamflow~(m^3~s^-1))) +
  xlab("Day of year") +

  # Vertical lines at ave_q day
  geom_vline(data = Baseflowplot1 %>% filter(water_year < 2025),
             aes(xintercept = yday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) +

  # Points for max_pf and ave_q
  geom_point(aes(x = yday, y = ave_q, color = as.factor(water_year)), 
             shape = 19, size = 1.5, alpha = 0.9)

BF_sum <- Baseflowplot1%>%
  filter(!is.na(Baseflowplot1$ave_q))

BF_sum

```


```{r, warning=F, size = 'small', echo=F, fig.width=9, fig.height=4.25}

Baseflowplot <- ggplot(Baseflowplot1 %>% 
                         filter(water_year < 2024), 
                       aes(x = yday, y = Q_m, color = as.factor(water_year))) +
  geom_line() +  
  #scale_color_viridis_d(name = "Year") + 
      scale_color_manual(values=c("#440154", "white", "white")) +
  xlim(91, 234) + 
 # ylim(0, 8.75) + 
  facet_grid(site~., scales = "free") + 
  theme_classic() + 
  theme(legend.position = "bottom")+
  ylab(expression(Streamflow~(m^3~s^-1))) +
  xlab("Day of year") +

  # Vertical lines at ave_q day
  geom_vline(data = Baseflowplot1 %>% filter(water_year < 2024),
             aes(xintercept = yday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) #+
# 
#   # Points for max_pf and ave_q
#   geom_point(aes(x = yday, y = ave_q, color = as.factor(water_year)), 
#              shape = 19, size = 1.5, alpha = 0.9) +
# 
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "BWL" & water_year < 2025), 
#             aes(x = 232, y = 8.7, label = "51 days"),  # Adjusted to max Q_m and max yday for top-right corner
#             color = "#440154", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "BWL" & water_year < 2025), 
#             aes(x = 232, y = 8, label = "30 days"),  # Adjust the position and text as needed
#             color = "#3B528B", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "BWL" & water_year < 2025), 
#             aes(x = 232, y = 7.3, label = "27 days"),  # Adjust the position and text as needed
#             color = "#fad311", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "GBL" & water_year < 2025), 
#             aes(x = 232, y = 0.85, label = "42 days"),  # Adjusted to max Q_m and max yday for top-right corner
#             color = "#440154", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "GBL" & water_year < 2025), 
#             aes(x = 232, y = 0.78, label = "38 days"),  # Adjust the position and text as needed
#             color = "#3B528B", size = 4, hjust = 1, vjust = 1, fontface = "bold") +
#   geom_text(data = Baseflowplot1 %>% 
#               filter(site == "GBL" & water_year < 2025), 
#             aes(x = 232, y = 0.71, label = "29 days"),  # Adjust the position and text as needed
#             color = "#fad311", size = 4, hjust = 1, vjust = 1, fontface = "bold")




Baseflowplot

```  




```{r, warning=F, size = 'small', echo=F, include=FALSE}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Fig2_baseflow_WY21.png", plot = Baseflowplot, width =4.5, height = 4.75, units = "in")

```



#### Calculate nitrogen demand:

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
# Constants
ra <- 0.5  # Autotrophic respiration coefficient (Hall & Tank 2003)
C_Nauto <- 16  # Autotrophic C:N ratio (Stelzer & Lamberti 2001)
C_Nhetero <- 20  # Heterotrophic C:N ratio (Hall & Tank 2003)
HGE <- 0.05  # Heterotrophic Growth Efficiency (Hall & Tank 2003)

# Calculate components of nitrogen demand
covariat_ndemand <- covariat_datq %>%
  group_by(site) %>% 
  mutate(
    # Autotrophic respiration (raGPP)
    raGPP = GPP_mean * ra,
    # Autotrophic assimilation of N
    Auto_N_assim = GPP_mean / C_Nauto,
    # Heterotrophic respiration (Rh)
    Rh = ER_mean - raGPP,
    # Heterotrophic assimilation of N
    Hetero_N_assim = (Rh * HGE) / C_Nhetero,
    # Total nitrogen demand
    Ndemand = Auto_N_assim + Hetero_N_assim, # unit should be g N m-2 d-1
    Ndemand = if_else(Ndemand < 0, 0.0001, Ndemand)) %>%
  dplyr::select(site, date, Ndemand)
```


```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}

### Quality check on supply calculations:
# (1) How many data points do we have for velocity (v) and width (w)?

morph_counts <- covariat_datq %>%
  group_by(site) %>%
  summarise(
    w_m_count = sum(!is.na(w_m)), 
    v_m_count = sum(!is.na(v_m))
  )

morph_counts
```


```{r, warning=F, size = 'small', echo=F, include=F,fig.width=6, fig.height=4}
##### better way to infill width and velocites to get at reach length:
# na.approx, and using Q as "Variables to be used for interpolation as in"


covariat_datqI <- covariat_datq %>%
  group_by(site) %>%
  arrange(date) %>%
  mutate(
    # Fit a linear model for w_m ~ Q_m using non-missing values
    w_m_apr = ifelse(is.na(w_m), predict(lm(w_m ~ depth_m, data = cur_data(), na.action = na.omit), newdata = cur_data()), w_m),
    v_m_apr = ifelse(is.na(v_m), predict(lm(v_m ~ Q_m, data = cur_data(), na.action = na.omit), newdata = cur_data()), v_m)
  ) %>%
  ungroup()

covariat_datqI <- covariat_datqI %>%
  group_by(site) %>%
  arrange(date) %>%  # Ensure chronological order
   mutate( ## 17 NAs
    w_m_apr = ifelse(is.na(w_m_apr),
                  rollapplyr(w_m_apr, width = 15, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE),
                  w_m_apr),
    v_m_apr = ifelse(is.na(v_m_apr),
                  rollapplyr(v_m_apr, width = 15, FUN = function(x) mean(x, na.rm = TRUE), fill = NA, partial = TRUE),
                  v_m_apr))
```

```{r, warning=F, size = 'small', echo=F,include=F, fig.width=9, fig.height=10}

TS_plot_w <- ggplot(covariat_datqI, aes(x = date, y = w_m_apr, color=site, shape =site)) +
  geom_point(size = 2, alpha = 0.5, position = position_dodge(width = 0.3)) +
  geom_point(size = 2, alpha = 0.7) +
  geom_point(aes(x = date, y = w_m), color="black") +
  scale_color_manual(values = siteC_colors) +
  scale_shape_manual(values = c(15, 0, 17, 2)) + theme_classic() +
  facet_grid(site~.)

TS_plot_v <- ggplot(covariat_datqI, aes(x = date, y = v_m_apr, color=site, shape =site)) +
  geom_point(size = 3, alpha = 0.95, position = position_dodge(width = 0.3)) +
  geom_point(size = 2, alpha = 0.7) +
    geom_point(aes(x = date, y = v_m), color="black") +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
    scale_color_manual(values = siteC_colors) +
  theme_classic() +facet_grid(site~.)

n_s_grid <- ggarrange(
  TS_plot_w,
  TS_plot_v,
  labels = c("a", "b"),
  ncol = 1, nrow = 2,
  label.x = 0.95,  # Move label to the right
  label.y = 1,     # Keep label at the top
  hjust = 1,       # Right-align the label
  vjust = 1        # Top-align the label
)

n_s_grid

# Where the black points are the survey measures of width or velocity.


```





#### Calculate nitrogen supply:

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
covariat_nsupply <- covariat_datqI %>%
  left_join(nitrogen_data, by=c("date", "site")) %>%
  dplyr::select(site, catch, date, water_year, reach_length,
                v_m_apr,w_m_apr, Q_m, K600_daily_mean,
                NO3_mgL_dl_sw, NH4_mgL_dl_sw, PO4_ugL_dl_sw, NO3_mgL_i, NH4_mgL_i) %>%
  mutate(
    K600_daily_mean = case_when(
      is.na(K600_daily_mean) & site == "BWL" ~ 22, 
      is.na(K600_daily_mean) & site == "BWU" ~ 16,
      is.na(K600_daily_mean) & site == "GBU" ~ 32,
      is.na(K600_daily_mean) & site == "GBL" ~ 25,
      TRUE ~ K600_daily_mean)) %>%
  mutate(Q_Ls = Q_m * 1000, # flow from csm to Ls
  # calculate reach length in m
  reachL = c((v_m_apr*w_m_apr*8640)/K600_daily_mean), # seconds to days
   # Calculate no3 supply 
  NO3_supply = c(((86400*Q_Ls*(NO3_mgL_dl_sw/1000))/(w_m_apr*reachL))),
  NO3_supply_new = c(((86400*Q_Ls*(NO3_mgL_i/1000))/(w_m_apr*reachL))),

   # Calculate nh3 supply 
  NH4_supply = c(((86400*Q_Ls*(NH4_mgL_dl_sw/1000))/(w_m_apr*reachL))), ## unit should be g N m-2 d-1
 NH4_supply_new = c(((86400*Q_Ls*(NH4_mgL_i/1000))/(w_m_apr*reachL))),
  PO4_supply = c(((86400*Q_Ls*(PO4_ugL_dl_sw/1e-6))/(w_m_apr*reachL))) ## unit should be g N m-2 d-1
   )  %>%
  dplyr::select(site, date, reachL, reach_length, Q_Ls, NO3_supply, 
                NO3_supply_new, NH4_supply, PO4_supply, NH4_supply_new, NO3_mgL_i, NH4_mgL_i, PO4_ugL_dl_sw)

```



```{r, warning=F, size = 'small', echo=F, include=F, fig.width=6, fig.height=4}
ndyn <- covariat_datq %>%
  left_join(covariat_nsupply, by=c("site", "date"))%>%
  left_join(covariat_ndemand, by=c ("site", "date")) 

summary(ndyn)

```


### **Figure 2**
Setting: Showing the delayed onset of baseflow in wetter years.
Color as water year. 

DOY - April 1 = 91
DOY - May 1 = 121
DOY -June 1 = 152
DOY - Sept 1 = 244
DOY - Oct 1 = 274

```{r, warning=F, size = 'small', echo=F, include=FALSE}

names(covariat_datq)

covariat_datq<-covariat_datq%>%
  mutate(yday=yday(date))


overall_pf<- covariat_datq %>%
    filter(yday >= 92, yday <= 153) %>%
  group_by(site, water_year) %>%
  summarise(
    max_pf = max(Q_m, na.rm = TRUE),
    min_pf = min(Q_m, na.rm = TRUE),
    max_pf_yday = yday[which.max(Q_m)],
    min_pf_yday = yday[which.min(Q_m)],
    mean_pf = mean(Q_m, na.rm = TRUE),
    mean_pf_yday = yday[which.min(abs(Q_m - mean_pf))]) %>%
  ungroup()


overall_bf<- covariat_datq %>%
    filter(yday >= 170, yday <= 274) %>%
  group_by(site) %>%
  summarise(
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q = mean(Q_m, na.rm = TRUE),
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]) %>%
  ungroup()


overall_bf_df <- covariat_datq %>%
  left_join(overall_bf, by = c("site"))


steepest_slopes <- overall_bf_df %>%
    filter(yday >= 153, yday <= 274) %>%
  #  arrange(site, water_year, yday) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2023 ~ mean_q + 0.031,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
  mutate(
    mean_q = case_when(
      site == "GBL" & water_year == 2024 ~ mean_q + 0.01,  # Adjust ave_q for GBL in 2021
      TRUE ~ mean_q)) %>%
    mutate(
    Q_m = case_when(
      site == "GBL" & water_year == 2021 ~ Q_m + 0.001,  # Adjust ave_q for GBL in 2021
      TRUE ~ Q_m)) %>%
  group_by(site, water_year) %>%
  summarise(
    ave_q = mean(mean_q, na.rm = TRUE),
    max_q = max(Q_m, na.rm = TRUE),
    min_q = min(Q_m, na.rm = TRUE),
    max_q_yday = yday[which.max(Q_m)],
    min_q_yday = yday[which.min(Q_m)],
    mean_q_yday = yday[which.min(abs(Q_m - mean_q))]
   # max_slope = max(slope, na.rm = TRUE),
  #  max_slope_yday = yday[which(slope == max(slope, na.rm = TRUE))[1]],
  #  min_slope = min(slope, na.rm = TRUE),
  #  min_slope_yday = yday[which(slope == min(slope, na.rm = TRUE))[1]]
  ) %>%
  ungroup() 

```




##### Plots of Ratios of NH4+- N supply to demand, ratios NO3-- N supply to demand

Remaid figure 4: Now with chemistry data aligned to data data frame so
that there are no sample NAs for NO3 and Nh4 when the metabolism model
failed to estimate that parameter on the same day as the chemistry
sample.

Color is now also by water year 2021-2024, and shape is reach location.

```{r, warning=F, size = 'small', echo=F, include=F}
covariat_NS_ND <- ndyn %>%
  mutate(jday = yday(date)) %>%
  dplyr::select(date, jday, Q_m, w_m, site, catch, water_year, 
                Ndemand, NO3_supply_new, NO3_supply, NH4_supply, NH4_supply_new,
                NH4_mgL_i, NO3_mgL_i) %>%
  mutate(
    NO3_SupDem = NO3_supply / Ndemand,
    NH4_SupDem = NH4_supply / Ndemand,
    N_SupDem = (NH4_supply + NO3_supply) / Ndemand
  ) %>%
  distinct(date, site, .keep_all = TRUE)  %>%
    mutate(WY_lab = case_when(
    water_year %in% c(2021, 2022) ~ "dry",
    water_year %in% c(2023) ~ "wet",
    water_year %in% c(2024) ~ "normal",
    TRUE ~ NA_character_  
  )) %>%
    group_by(site, date, catch) %>%
    summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop")

covariat_NS_ND1 <- covariat_NS_ND%>%
  left_join(Baseflowplot1, by=c("site", "date", "catch",
                                "water_year", "Q_m"))
  


NO3_supp_plots <- ggplot(covariat_NS_ND1%>%filter(water_year<2022 & jday<274), aes(y = (NO3_supply_new), x = jday, color = as.factor(water_year), shape=site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
  geom_point( aes(y = log(NH4_supply_new+1), x = jday, color = as.factor(water_year), shape=site), 
                size = 2, alpha = 0.85, position = position_dodge(width = 0.3)) +
    scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
  scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
  ylim(0,1.50) +
   xlab(NULL) +
       scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
      scale_fill_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10)) +
  geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
     ylab(expression(Supply~(g~N~m^-2~d^-1))) +
  theme_classic() +
  scale_shape_manual(values = c(15,0,17,2)) +
    theme(legend.position = "none") +
    geom_vline(data = covariat_NS_ND1 %>% filter(water_year < 2022),
             aes(xintercept = jday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) +
   facet_grid(catch~.) 

# Remove legend

# Define threshold for high NO3 demand (above 3rd quartile)
high_Ndemand_threshold <- 0.075

N_demand_plot1 <- ggplot(covariat_NS_ND1%>%filter(water_year<2022), aes(y = (Ndemand), x = jday, color = as.factor(water_year), shape = site)) +
  geom_point(size = 2, alpha = 0.85, position = position_dodge(width = 0.8)) +
 # scale_color_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
      scale_color_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
      scale_fill_manual(values=c("#440154", "#3B528B", "#4bcc94")) +
  #  scale_fill_viridis_d(option = "viridis", name = "Water year") +  # Use _c for continuous data
  xlab("Day of year") + 
    geom_smooth(aes(group = as.factor(water_year), fill = as.factor(water_year)), 
            se = TRUE, linewidth = 1, method = "gam", formula = y ~ s(x, k = 10), alpha = 0.1)+
    scale_x_continuous(
  limits = c(15, 285),
  breaks = seq(15, 285, by = 30)  
)+
    scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                         labels = scales::label_number()) +
 # geom_hline(yintercept = high_Ndemand_threshold, linetype = "dashed", color = "grey25") +
  scale_shape_manual(values = c(15, 0, 17, 2)) +
  ylab(expression(Demand~(g~N~m^-2~d^-1))) +
  theme_classic() +
  theme(legend.position = "none")  +
    facet_grid(catch~.) +
    geom_vline(data = covariat_NS_ND1 %>% filter(water_year < 2022),
             aes(xintercept = yday[ave_q == ave_q], color = as.factor(water_year)),
             linetype = "dashed", size = 0.5, alpha = 0.9) 


N_demand_plot1

```

### **Figure 5**


```{r, warning=F, size = 'small', echo=F,include=T, fig.width=12, fig.height=6}

n_s_grid <- ggarrange(
  NO3_supp_plots,
  N_demand_plot1,
  labels = c("b", "c"),
    align = c("v"),
  ncol = 1, nrow = 2,
  label.x = 0.91,  # Move label to the right
  label.y = 1,     # Keep label at the top
  hjust = 1,       # Right-align the label
  vjust = 1        # Top-align the label
)



n_s_grid


```

```{r, warning=F, size = 'small', echo=F, include=F, fig.width=9, fig.height=4}
# ggsave("/Users/kellyloria/Documents/Publications/CH1\ biogeochem\ linkages/Figs/Nsupply_grid_CH1_SFS_ppt_21.png", plot = n_s_grid, width = 8.5, height = 7.75, units = "in")

```

### Nitrogen Uptake dynamics from TASCC

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=6, fig.height=4}
n_df <- n_up%>%
  filter(success=="yes") %>%
  mutate(
         Uadd_g_m2_min = c(Uadd_ug_L_min/1000000),
         Uadd_g_m2_D = c(Uadd_g_m2_min*1440)) 

uptake <- n_df %>%
  pivot_wider(
    names_from = method, 
    values_from = Uadd_g_m2_D, 
    names_prefix = "Uadd_g_m2_D_") %>%
  dplyr::select(site, date, Uadd_g_m2_D_NH3, Uadd_g_m2_D_NO3)

summary(uptake)
```

```{r, warning=F, size = 'small', echo=F, include=F,fig.width=6, fig.height=4}

uptake_df <-  covariat_NS_ND%>%
  full_join(uptake)

```

### Linking ecosystem funtion and biogeochemical dynamics:

We found that autotrophic production (and so nitrogen demand) is
suppressed in wetter years. We want to know how the reduced autrophic
production could effect nitrogen retention and transport.

-   To estimate what amount of the nitrogen can be could be transformed
    via uptake in at a given reach?

    -   Fractional uptake efficiencies (FUE) for N?

    -   Might consider the FUE at the lower station as downstream
        export?

-   To estimate what amount of the nitrogen can be could be transformed
    via uptake in a stream

    -   Difference in FUE between reaches for overall stream processing?

-   In wet years, even though supply increases, we would expect
    biological processes like uptake to decrease. So what percent of the
    N supply is either processes or exported in wet verse dry years?

##### Consider the units

-   Areal max uptake - Ut - ug / m2 min -\> we converted to g m2 day

-   Uptake velocity - Vf - m / min

-   Uptake length - Sw - m

-   Supply and demand is in g N m-2 d-1

### Methods for evaluating nitrogen cycling

According to [Covino et al.
2018](https://www.journals.uchicago.edu/doi/abs/10.1086/699202?casa_token=z4FRLNRrBH8AAAAA:g2dwDEk5ct5-NxSqV0w2PILcK9ArSQhBnEeVBmOHiEHU_RMeWBRu4RWAzJG6ilh7wSeD30YDPkyH&casa_token=_yfAVNMW2ZQAAAAA:LH4hXzHGlRGjdsnCKp28IBUCho9RqquzLdge8TkzP3UNGQdvoI8SqtoxxyHw67tIEtm_4uM6CLmY)
:

-   Supply (S): The available nitrogen in the system (in g N/m²/day).

-   Demand (D): The total biological nitrogen demand (in g N/m²/day).

-   Uptake rate (U): The amount of nitrogen removed from the water
    column via assimilation, denitrification, or other biological
    processes.

#### U = (D \* S) / (D + S)

-   When demand (D) is much greater than supply (S): Almost all
    available nitrogen is taken up, so uptake approaches supply (U ≈ S)

-   When supply (S) is much greater than demand (D): Uptake is limited
    by biological demand, and uptake approaches demand (U ≈ D)

```{r, warning=F, size = 'small', echo=F, include=F,fig.width=9, fig.height=3}
uptake_dyn <- uptake_df %>%
  group_by(site, date) %>%
  mutate(
    jday = yday(date),
    U_NO3_calc = c((Ndemand*NO3_supply_new)/(Ndemand+NO3_supply_new)),
    U_NH4_calc = c((Ndemand*NH4_supply_new)/(Ndemand+NH4_supply_new))) %>%
  group_by(site, date) %>%
    summarise(across(everything(), mean, na.rm = TRUE), .groups = "drop") %>%
  mutate(catch = case_when(
    site== "BWL" ~ "BW",
    site== "BWU" ~ "BW",
    site== "GBL" ~ "GB",
    site== "GBU" ~ "GB",
    TRUE ~ NA_character_  
  ))
  

summary(uptake_dyn)

hist1<- ggplot(uptake_dyn, aes(x = U_NO3_calc)) +
  geom_histogram(aes(fill = site,  color = site),  position = "dodge", alpha = 0.7, bins = 10) +  scale_color_manual(values = site_colors) + scale_fill_manual(values = site_colors) + theme_minimal()
hist1

hist2<- ggplot(uptake_dyn, aes(x = U_NH4_calc)) +
  geom_histogram(aes(fill = site,  color = site),  position = "dodge", alpha = 0.7, bins = 10) +  scale_color_manual(values = site_colors) + scale_fill_manual(values = site_colors) + theme_minimal()
hist2


n_df_temp <- uptake_dyn %>%
  group_by(site) %>%
  summarise(U_NO3_calc= c(mean(U_NO3_calc, na.rm=T)*1000),
            U_NH4_calc= c(mean(U_NH4_calc, na.rm=T)*1000), 
            Q_ls=c(mean(Q_m, na.rm=T)*1000))
#   
# n_df_temp$Uadd_mg_m2_D <- (n_df_temp$Uadd_g_m2_D*1000)


```

```{r, warning=F, size = 'small', echo=T, include=T,fig.width=9, fig.height=6}

uptake_dyn <- uptake_dyn %>%
  mutate(
    S_D_NO3 = NO3_supply_new / Ndemand,
    S_D_NH4 = NH4_supply_new / Ndemand
  )

summary(uptake_dyn$S_D_NO3)
summary(uptake_dyn$S_D_NH4)
quantile(uptake_dyn$S_D_NO3, probs = c(0.1, 0.9), na.rm = TRUE)
quantile(uptake_dyn$S_D_NH4, probs = c(0.1, 0.9), na.rm = TRUE)

```

```{r, warning=F, size = 'small', echo=F, include=F,fig.width=9, fig.height=6}
hist1<- ggplot(uptake_dyn%>%filter(water_year<2024), aes(x = log(S_D_NO3+1))) +
 # geom_histogram(aes(fill = site,  color = site),  position = "dodge", alpha = 0.7, bins = 5) +  scale_color_manual(values = site_colors) + 
  geom_density(alpha = 0.7, aes(fill = site,  color = site)) +
  scale_color_manual(values = site_colors)  +
  scale_fill_manual(values = site_colors) + theme_minimal() +
  xlab(expression(log(NO[3]~Supply/N~Demand)))  
hist1

hist2<- ggplot(uptake_dyn%>%filter(water_year<2024), aes(x = log(S_D_NH4+1))) +
 # geom_histogram(aes(fill = site,  color = site),  position = "dodge", alpha = 0.7, bins = 5) +  scale_color_manual(values = site_colors) + 
  geom_density(alpha = 0.7, aes(fill = site,  color = site)) +
  scale_color_manual(values = site_colors)  +
  scale_fill_manual(values = site_colors) + theme_minimal() +
  xlab(expression(log(NH[4]~Supply/N~Demand))) 
hist2

```

```{r, warning=F, size = 'small', echo=F,include=F, fig.width=7, fig.height=6}

Ndyn_grid <- ggarrange(
  hist1,
  hist2,
  ncol = 1, nrow = 2,
    labels = c("a", "b"),
  common.legend = TRUE, 
   widths = c(1.75, 2.25),
  legend = "bottom")

Ndyn_grid
```

### Supply and demand dynamics tend to vary with catchment and reach:

Looks like N supply is greater than demand at GB -- indicating that
uptake (U) should be low and possibly limited by low biologic demand. So
we expect more nitrogen to be transported in GB.

Looks Like N demand is greater than supply at BW -- U should be high and
most available nitrogen should be retained.

### **Figure 6**


```{r, warning=F, size = 'small', echo=F, include=T,fig.width=9, fig.height=6}

no3_S_plot <- ggplot(uptake_dyn, aes(x = (NO3_supply_new), y = (U_NO3_calc), color = site, shape = site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.1)) +
  #scale_color_viridis_c(option = "viridis", name = "Day of year") + 
      scale_color_manual(values = siteC_colors) +
    scale_x_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                         labels = scales::label_number()) + 
      scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                           labels = scales::label_number()) + 
  # scale_y_continuous(trans = "log") +#, breaks = c(0.01, 0.03, 0.1, 0.3, 1, 3, 10)) 
  #scale_x_continuous(trans = "log")  +
  geom_abline(slope = (1), intercept = 0, color = "black") +  # 1:1 line
  scale_shape_manual(values = c(15, 0, 17, 2)) +
 #xlim(0, 1) +
  theme_classic() + xlab(NULL)+ 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  #xlab(expression(log(N~supply)~(g~m^-2~d^-1))) +
  ylab(expression(log(NO[3]~U[t])~(g~m^-2~d^-1)))



no3_D_plot <-ggplot(uptake_dyn, aes(x = (Ndemand), y = (U_NO3_calc), color = site, shape = site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.1)) +
 # scale_color_viridis_c(option = "viridis", name = "Day of year") + 
    scale_color_manual(values = siteC_colors) +
      scale_x_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                           labels = scales::label_number()) + 
      scale_y_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                           labels = scales::label_number()) + 
   # xlim(0, 0.45) +
    geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
  scale_shape_manual(values = c(15, 0, 17, 2)) +theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
 # xlab(expression(log(N~demand)~(g~m^-2~d^-1))) +
  xlab(NULL)+ylab(NULL)
   #ylab(expression(log(NO[3]~U[calc])~(g~m^-2~d^-1)))


nh4_S_plot <-ggplot(uptake_dyn, aes(x = NH4_supply_new, y = U_NH4_calc,color = site, shape = site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.1)) +
  #scale_color_viridis_c(option = "viridis", name = "Day of year") +
     # xlim(0, 0.150) +
        scale_x_continuous(trans = "log", 
                           breaks = scales::log_breaks(base = 10), 
                             labels = scales::label_number(),
                                 limits = c(1e-4, 10),) + 
      scale_y_continuous(trans = "log", 
                         breaks = scales::log_breaks(base = 10),
                         limits = c(1e-4, 1e-1),
                           labels = scales::label_number()
                         ) + 
    geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
      scale_color_manual(values = siteC_colors) +
  scale_shape_manual(values = c(15, 0, 17, 2)) +theme_classic() +
  xlab(expression(log(N~supply)~(g~m^-2~d^-1))) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
   ylab(expression(log(NH[4]~U[t])~(g~m^-2~d^-1)))


nh4_D_plot <-ggplot(uptake_dyn, aes(x = (Ndemand), y = (U_NH4_calc), color = site, shape = site)) +
  geom_point(size = 3, alpha = 0.85, position = position_dodge(width = 0.1)) +
  #scale_color_viridis_c(option = "viridis", name = "Day of year") + 
      scale_color_manual(values = siteC_colors) + 
          scale_x_continuous(trans = "log", breaks = scales::log_breaks(base = 10),
                               labels = scales::label_number()) + 
      scale_y_continuous(trans = "log", 
                         breaks = scales::log_breaks(base = 10), 
                         limits = c(1e-4, 1e-1),
                           labels = scales::label_number()) + 
    geom_abline(slope = 1, intercept = 0, color = "black") +  # 1:1 line
  scale_shape_manual(values = c(15, 0, 17, 2)) +theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  xlab(expression(log(N~demand)~(g~m^-2~d^-1))) + ylab(NULL)
    #  xlim(0, 0.45) +
  # ylab(expression(log(NH[4]~U[calc])~(g~m^-2~d^-1)))

Ndyn_grid <- ggarrange(
  no3_S_plot,
  no3_D_plot,
  nh4_S_plot,
  nh4_D_plot,
  align=c("v"),
  labels = c("a", "b", "c", "d"),
  ncol = 2, nrow = 2,
  common.legend = TRUE, 
  label.x = 0.3, 
  label.y = 1,    
  hjust = 1,  
  vjust = 1,  
  widths = c(0.49, 0.48),
  legend = "bottom")

Ndyn_grid


```


