---
title: 'Streammetabolizer output for GBL'
author: "Kelly Loria"
date: "2025-01-31"
output:
   html_document:
    theme: united
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 12px;}
code.r{font-size: 8px;}
pre {font-size: 10px}
</style>
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

```{r, echo = F, message = F}
### Packages
setwd("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/")

library(tidyverse)
library(lubridate)
library(dataRetrieval)
library(viridis)
library(cowplot)
library(zoo)
library(readxl)
library(ggpubr)
library(streamMetabolizer)
```

### Background
This is a small, turbulent, shaded stream with a fair amount of fine particulate organic matter. 

#### Identifying reasonable modeled estimates of K600
We then ran normal pooling on k600 in stream metabolizer model: `b_Kb_oipi_tr_plrckm.stan` to get modeled K600 to see if we could resolve the negative correlation between ER and K600.
Priors on `K600_lnQ_nodes_meanlog` were set as 5 bins based on `rnorm(1000, mean = logQ_mean, sd = logQ_sd)` centered around the mean and logQ values 1-2 sd away from the mean. 

##### Load GBL versions of metab data from different model segments with normal prior for gas exchange. 
We chose segments of time where we believe GPP occurred and was greater than 0. 
These chunks of time are from a previous model where we binned flow and incorporated measured and estimated K600 priors from gas exchange measurements a the reach. 

```{r, warning=F, size = 'small', echo=FALSE}
met.May23 <- read_csv("./GBL_May23_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
  mutate(lab = "May23")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  filter(K600_daily_Rhat<1.1)

fits.May23 <- read_csv("./GBL_May23_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
  mutate(lab = "May23")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.May21 <- read_csv("./GBL_May21_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
  mutate(lab = "May21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.May21 <- read_csv("./GBL_May21_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
    mutate(lab = "May21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100)) 

met.Mar24 <- read_csv("./GBL_Mar24_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
    mutate(lab = "Mar24")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar24 <- read_csv("./GBL_Mar24_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Mar24")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.Mar22 <- read_csv("./GBL_Mar22_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
     mutate(lab = "Mar22")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar22 <- read_csv("./GBL_Mar22_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
       mutate(lab = "Mar22")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100)) 


met.Mar21 <- read_csv("./GBL_Mar21_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
       mutate(lab = "Mar21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean) %>%
  filter(K600_daily_Rhat<1.05)

fits.Mar21 <- read_csv("./GBL_Mar21_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
       mutate(lab = "Mar21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))


met.Feb24 <- read_csv("./GBL_Feb24_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
       mutate(lab = "Feb24")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Feb24 <- read_csv("./GBL_Feb24_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
   mutate(lab = "Feb24")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100),
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))  


met.Apr22 <- read_csv("./GBL_Apr22_K_2025-01-30_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Apr22")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.Apr22 <- read_csv("./GBL_Apr22_K_2025-01-30_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Apr22")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.full21 <- read_csv("./GBL_21_newbin_2025-02-02_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Full_21")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  %>%
  filter(K600_daily_Rhat<1.05)

fits.full21 <- read_csv("./GBL_21_newbin_2025-02-02_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Full_21")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))


met.full <- read_csv("./GBL_Full_2025-02-02_daily.csv") %>%
  mutate(year = year(date))%>%
  mutate(jday = yday(date))%>%
  mutate(mon = month(date))%>%
   mutate(lab = "Full")%>%
  mutate(NEP_daily_mean = GPP_daily_mean + ER_daily_mean)  

fits.full <- read_csv("./GBL_Full_2025-02-02_errors.csv_mod_and_obs_DO.csv") %>%
  mutate(week = week(solar.time))%>%
  mutate(jday = yday(solar.time)) %>%
     mutate(lab = "Full")%>%
  mutate(
    DO_pr_sat = c((DO.obs/DO.sat) * 100), 
    DO_pr_sat_mod = c((DO.mod/DO.sat) * 100))

met.df <- rbind(met.May23, 
                met.May21, 
                met.Mar24,
                met.Mar21,
                met.Apr22,
                met.Feb24)
met.df$source <- ("normal_mod")
fits.df <- rbind(fits.May23, fits.May21, fits.Mar21,
                 fits.Mar24, fits.Apr22, met.Feb24)


input_dat <- readRDS("./25_GBL_mod_dat.rds") 

measured_K <- read.csv("/Users/kellyloria/Documents/Publications/2024_stream_metab_output/GBL_k600/ArN_NEON_K_GBL_notpooled_241204.csv")%>%
  mutate(date= as.Date(date, format=c("%m/%d/%y"))) 
measured_K$source <- ("measured")

```


```{r, warning=F, size = 'small', include=FALSE, echo=FALSE, fig.width=10, fig.height=12}
met.df %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~lab, scales="free", nrow = 4)+
  theme_bw()
```

### Metabolism segements  with CIs

Where GPP is in blue and ER is in orange.

#### 2021 re-run with better binning 

`K600_lnQ_nodes_meanlog = log(25)` Where 25 was the mean value from observed measurements and normal pooled modeled, 
`K600_lnQ_nodes_sdlog = 1.1` was 1.1, 
`bayes_specs_new$K600_lnQ_nodes_centers <- log_bins` was from `prior_samples <- rnorm(1000, mean = logQ_mean, sd = logQ_sd)`
based on mean discharge for all flows > 0.99 cms to cut off some extreme values.


```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=12}
met.full %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_line(col = "cyan4", alpha = 0.8)+
  geom_ribbon(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), 
              alpha = 0.4, fill = "cyan4")+
  geom_line(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_ribbon(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), 
              alpha = 0.4, fill = "chocolate3")+ 
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~year, scales="free", nrow = 4)+
  theme_bw()
```

## Checking model convergence
Make sure the chains converged; all r-hat values were well below 1.05 (the red line)
for GPP, ER, and K600. 
The blue lines are the mean for each parameter. 

** Not all all segments had great convergence. 

```{r, , echo=FALSE, warning=F, include=FALSE, size = 'small', fig.width=9, fig.height = 2.5}
k600.rhat <- mean(na.omit(met.df$K600_daily_Rhat))
GPP.rhat <- mean(na.omit(met.df$GPP_Rhat))
ER.rhat <- mean(na.omit(met.df$ER_Rhat))

# Create a dataframe for the vertical lines
rhat_lines <- data.frame(
  Param = c("K600_daily_Rhat", "GPP_daily_Rhat", "ER_daily_Rhat"),
  yintercept = c(k600.rhat, GPP.rhat, ER.rhat)
)

met.df %>%
  select(date, GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  pivot_longer(2:4, values_to = "Rhat", names_to = "Param") %>%
  ggplot(aes(x = date, y = Rhat), col=lab) +
  geom_hline(yintercept = 1.05, col = "red", lty = 2) + 
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(data = rhat_lines, aes(yintercept = yintercept), col = "blue", lty = 2) + 
  theme_classic() +
  facet_wrap(~Param, ncol = 3)
```

### 2021 re-run rhats on GPP and ER
K600 was already filtered (< 1.05).

```{r, , echo=T, warning=F, size = 'small', fig.width=9, fig.height = 2.5}
k600.rhat <- mean(na.omit(met.full$K600_daily_Rhat))
k600.rhat
GPP.rhat <- mean(na.omit(met.full$GPP_Rhat))
GPP.rhat
ER.rhat <- mean(na.omit(met.full$ER_Rhat))
ER.rhat

# Create a dataframe for the vertical lines
rhat_lines <- data.frame(
  Param = c("K600_daily_Rhat", "GPP_daily_Rhat", "ER_daily_Rhat"),
  yintercept = c(k600.rhat, GPP.rhat, ER.rhat)
)

met.full %>%
  select(date, GPP_daily_Rhat, ER_daily_Rhat, K600_daily_Rhat) %>%
  pivot_longer(2:4, values_to = "Rhat", names_to = "Param") %>%
  ggplot(aes(x = date, y = Rhat), col=lab) +
  geom_hline(yintercept = 1.05, col = "red", lty = 2) + 
  geom_point(size = 0.5, alpha = 0.5) +
  geom_hline(data = rhat_lines, aes(yintercept = yintercept), col = "blue", lty = 2) + 
  theme_classic() +
  facet_wrap(~Param, ncol = 3)
```



```{r, warning=F, , echo=FALSE, size = 'small', fig.width=9, fig.height = 18}
fits.day <- fits.full%>% 
group_by(date, lab) %>%
 dplyr:: summarize(rmse = sqrt(mean((DO.mod-DO.obs)^2)),
            sd = sd(DO.obs),
            min = min(DO.obs, na.rm = T),
            max = max(DO.obs, na.rm = T),
            range = range(max-min, na.rm = T),
            nrmse = rmse/range,
            minT = min(temp.water, na.rm = T),
            maxT = max(temp.water, na.rm = T),
            rangeT= range(maxT-minT, na.rm = T),
            minQ= min(discharge, na.rm = T), 
            maxQ= max(discharge, na.rm = T),
            rangeQ= range(maxQ-minQ, na.rm = T)
            )

summary(fits.day)
```

## Looking at the K600 and ER relationships from segements with normal K600 pooling

Here is the normal pooling run configuration:
**bayes_specs_new**  

`bayes_name_new <- mm_name(type = 'bayes', pool_K600 = "normal", err_obs_iid = TRUE, err_proc_iid = TRUE, ode_method = "trapezoid", deficit_src = 'DO_mod', engine = 'stan')`

 Where each "lab" segment was run as individual model with the `K600_lnQ_nodes_meanlog` adjusted to match streamline in cms during that time. 

### K-ER, K-Q, general K600 trends
There is a strong negative correlation between ER and K600 (-0.96). 
GPP and K600 are slightly less correlated (0.288).

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=8, fig.height = 5}

ker <- ggplot(met.df, aes(x = K600_daily_mean, y = ER_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+
  theme_bw()+ geom_smooth(method="lm", se=F, color="grey50") +
  scale_color_viridis_d(option = "inferno") +  theme(legend.position = "null")


KER_cor <- round(cor(met.df$ER_daily_mean, met.df$K600_daily_mean, use = "complete.obs"),3)
print(KER_cor)

KGPP_cor <-round(cor(met.df$GPP_daily_mean, met.df$K600_daily_mean, use = "complete.obs"),3)
print(KGPP_cor)

kgpp <- ggplot(met.df, aes(y = K600_daily_mean, x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")

kyear <- met.df %>%
ggplot(aes(x = as.factor(year), y = K600_daily_mean, fill = as.factor(lab), col = as.factor(lab)))+
    geom_hline(yintercept = 10, lty = 2)+
  geom_boxplot(alpha = 0.8)+
  theme_bw()+
  scale_fill_viridis_d(option ="inferno") +  # Use _d for discrete
  scale_color_viridis_d(option = "inferno") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mod.env.ag <- input_dat%>%
  mutate(date= as.Date(solar.time))%>%
  group_by(date) %>%
  dplyr:: summarise(
    DO.obs = mean(DO.obs, na.rm=T),
    DO.sat = mean(DO.sat, na.rm=T),
    depth = mean(depth, na.rm=T),
    temp.water = mean(temp.water, na.rm=T),
    light = mean(light, na.rm=T),
    discharge =  mean(discharge, na.rm=T))


k.q <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = lab))+
  geom_point( size = 2)+ geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "inferno") 

mean_k_mod <- mean(met.df$K600_daily_mean)
mean_k_mod
mean_k_measure <- mean(measured_K$K600)
mean_k_measure

possibleK <- (mean_k_mod + mean_k_measure)/2
possibleK

p1 <- plot_grid(ker, kgpp, k.q, rel_widths =c(0.3, 0.3, 0.4), ncol = 3)
plot_grid(p1, kyear, ncol = 1)

```

### Compared modeled and measured K600

Plots for (1) measured v modeled K600 and flow and (2) logK600 and log(flow+1).

Could be one poor measurement at the highest flow for measured gas exchange.
But in general the modeled K600 does seem similar to the measured, which is kind of nice to see how robust the `pool_K600 = "normal"` is getting at K600.

```{r, warning=F, size = 'small', echo=FALSE, fig.align = "center", fig.width=9, fig.height = 2.5}

k.q <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  full_join(measured_K, by = c("date", "source",
                                 "K600_daily_mean"= "K600",
                                "discharge"= "Q_cms")) %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = source, shape=source))+
  geom_point( size = 2) +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "viridis") 


k.q_log <- met.df %>%
  left_join(mod.env.ag, by = "date") %>%
  full_join(measured_K, by = c("date", "source",
                                 "K600_daily_mean"= "K600",
                                "discharge"= "Q_cms")) %>%
  ggplot(aes(x = log(discharge+1), y = log(K600_daily_mean), col = source, shape=source))+
  geom_point( size = 2) +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "viridis") 



plot_grid(k.q, k.q_log, ncol = 2)

```


### 2021 re-run test with binned K600 
On `met.full21_clean` filtered for days with `GPP_daily_Rhat<1.06` and `ER_daily_Rhat<1.06`.

```{r, warning=F, size = 'small', echo=T, fig.align = "center", fig.width=8, fig.height = 5}

met.full21_clean <- met.full %>%
  filter(GPP_daily_Rhat<1.05)%>%
  filter(GPP_97.5pct>0)%>%
  filter(ER_daily_Rhat<1.05) %>%
  filter(ER_2.5pct<0)%>%
  filter(K600_daily_Rhat<1.05)

ker <- ggplot(met.full21_clean, aes(x = K600_daily_mean, y = ER_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+
  theme_bw()+ geom_smooth(method="lm", se=F, color="grey50") +
  scale_color_viridis_d(option = "inferno") +  theme(legend.position = "null")


KER_cor <- round(cor(met.full21_clean$ER_daily_mean, met.full21_clean$K600_daily_mean, use = "complete.obs"),3)
print(KER_cor)

KGPP_cor <-round(cor(met.full21_clean$GPP_daily_mean, met.full21_clean$K600_daily_mean, use = "complete.obs"),3)
print(KGPP_cor)

kgpp <- ggplot(met.full21_clean, aes(y = K600_daily_mean, x = GPP_daily_mean, col = as.factor(lab)))+
  geom_point(size = 2)+  geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  scale_color_viridis_d(option = "inferno") +
  theme(legend.position = "null")

kyear <- met.full21_clean %>%
ggplot(aes(x = as.factor(year), y = K600_daily_mean, fill = as.factor(lab), col = as.factor(lab)))+
    geom_hline(yintercept = 10, lty = 2)+
  geom_boxplot(alpha = 0.8)+
  theme_bw()+
  scale_fill_viridis_d(option ="inferno") +  # Use _d for discrete
  scale_color_viridis_d(option = "inferno") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mod.env.ag <- input_dat%>%
  mutate(date= as.Date(solar.time))%>%
  group_by(date) %>%
  dplyr:: summarise(
    DO.obs = mean(DO.obs, na.rm=T),
    DO.sat = mean(DO.sat, na.rm=T),
    depth = mean(depth, na.rm=T),
    temp.water = mean(temp.water, na.rm=T),
    light = mean(light, na.rm=T),
    discharge =  mean(discharge, na.rm=T))


k.q <- met.full21_clean %>%
  left_join(mod.env.ag, by = "date") %>%
  ggplot(aes(x = discharge, y = K600_daily_mean, col = lab))+
  geom_point( size = 2)+ geom_smooth(method="lm", se=F, color="grey50", lty="dashed") +
  theme_bw()+
  #scale_x_log10(limits = c(20, 800))+
   scale_color_viridis_d(option = "inferno") 



mean_k_mod <- mean(met.full21_clean$K600_daily_mean)
mean_k_mod
mean_k_measure <- mean(measured_K$K600)
mean_k_measure


p1 <- plot_grid(ker, kgpp, k.q, rel_widths =c(0.3, 0.3, 0.4), ncol = 3)
plot_grid(p1, kyear, ncol = 1)

```

## Final thoughts.
The direction of the K600 ~ flow relationship looks more logical, where K600 increases with flow.
ER and K600 are still very negatively correlated (-0.97) but it might just mean that ER is not an identifiable parameter in this reach. 

The mean modeled K600 is lower than expected, 5.50 when it should be closer to 25. We'll see how that changes on the final model run with longer time series. 

### Final cleaned model output:

```{r, warning=F, size = 'small', echo=FALSE, fig.width=10, fig.height=12}
met.full21_clean %>% 
ggplot(aes(x = jday, y = GPP_daily_mean))+
  geom_abline(slope = 0, intercept = 0, width = 1, alpha = 0.6)+
  geom_point(col = "cyan4", alpha = 0.8)+
   geom_errorbar(aes(ymin = GPP_daily_2.5pct, ymax = GPP_daily_97.5pct), width = 0.2, color = "cyan4") +
  geom_point(aes(x = jday, y = ER_daily_mean),
            col = "chocolate3", alpha = 0.8)+
  geom_errorbar(aes(ymin = ER_daily_2.5pct, ymax = ER_daily_97.5pct), width = 0.2, color = "chocolate3") +
  geom_point(aes(x = jday, y = NEP_daily_mean),
             alpha = 0.5, size = 0.2)+
  facet_wrap(~year, scales="free", nrow = 4)+
  theme_bw()
```